### 2018-08-30 random_vorticity.tex
Getting this going

 \documentclass[a4paper,11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{epstopdf}
\epstopdfsetup{update}

\newcommand{\ba}{\begin{array}}
\newcommand{\ea}{\end{array}}

\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\newcommand{\bc}{\begin{center}}
\newcommand{\ec}{\end{center}}

\newcommand{\ds}{\displaystyle}

\newcommand{\bt}{\begin{tabular}}
\newcommand{\et}{\end{tabular}}

\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}

\newcommand{\bd}{\begin{description}}
\newcommand{\ed}{\end{description}}

\newcommand{\bp}{\begin{pmatrix}}
\newcommand{\ep}{\end{pmatrix}}

\newcommand{\pd}{\partial}
\newcommand{\sech}{\mbox{sech}}

\newcommand{\cf}{{\it cf.}~}

\newcommand{\ltwo}{L_{2}(\mathbb{R}^{2})}
\newcommand{\smooth}{C^{\infty}_{0}(\mathbb{R}^{2})}

\newcommand{\br}{{\bf r}}
\newcommand{\bk}{{\bf k}}
\newcommand{\bv}{{\bf v}}

\newcommand{\gnorm}[1]{\left|\left| #1\right|\right|}
\newcommand{\ipro}[2]{\left<#1,#2 \right>}

%\setlength{\topmargin}{-40pt} \setlength{\oddsidemargin}{0pt}
%\setlength{\evensidemargin}{0pt} \setlength{\textwidth}{460pt}
%\setlength{\textheight}{680pt}
\title{Evolution of Narrow-Band Spectra in Deep-Water Constant Vorticity Flows}
\author{Christopher W. Curtis and Mackensie Murphy}
\date{}
\begin{document}
\maketitle
\section*{Introduction}
To describe the evolution of a nearly linear wave-train in a constant-vorticity-shear current running over infinitely deep water, we use the form of the nonlinear Schr\"{o}dinger equation derived in \cite{curtis8}, though see also \cite{thomas2012nonlinear}, given by 
\[
i\pd_{\tau}\eta_{1} + \alpha_{d}\pd_{\xi}^{2}\eta_{1} + \alpha_{nl}\left|\eta_{1} \right|^{2}\eta_{1} = 0, 
\]
where $\tau = \epsilon^{2}t$, $\xi = \epsilon(x+c_{g}t)$, and 
\[
c_{g} = \frac{1+3\tilde{\sigma}k_{0}^{2}}{2s\Omega - \omega}, ~ s = \mbox{sgn}(k_{0}),
\]
 \begin{align*}
\alpha_{d}(k_{0},\omega) = & \frac{(c^2_{g} - 3|k_{0}|\tilde{\sigma})}{2\Omega-s\omega},\\
\alpha_{nl}(k_{0},\omega) = & \frac{k_{0}\left( sk_{0}^{3}\left(8 + \tilde{\sigma}k_{0}^{2} + 2(\tilde{\sigma}k_{0}^{2})^{2}\right) + \omega \alpha_{v}\right)}{\left(2s\Omega -\omega\right)(1+c_{g}\omega)\left(4\Omega^2-s(2k_{0}(1+4\tilde{\sigma}k_{0}^{2})+2\omega\Omega)\right)},
\end{align*}
where $\Omega$ denotes the dispersion relationship and $\omega$ is the non-dimensionalized magnitude of the vorticity of the flow, and $\tilde{\sigma}$ is the surface tension.  In a now seminal paper, \cite{alber} studies families of profiles with relatively narrow banded spectra around $k_{0}$ and modulational instability (MI) is either manifested or supressed.  The results in \cite{alber} were confirmed numerically in \cite{dysthe2,dysthe3} by examing the mean properties associated with ensembles of initial conditions.  This was done by starting from the $2L$ periodic-initial condition
\[
\eta_{1}(\xi,0) = \epsilon_{rms}\sqrt{\frac{\delta \tilde{k}}{\sigma\sqrt{\pi}}} \sum_{k=-K+1}^{K}e^{-\tilde{k}^{2}/2\sigma^{2}}e^{i\theta_{k}}e^{i\tilde{k}\xi}, ~ \tilde{k} = \frac{\pi k}{L}
\]
where $\delta \tilde{k}=\pi/L$ and the phases $\theta_{k}$ are randomly chosen uniformly between $0$ and $2\pi$.  We can then readily show that, for $L\gg 1$, 
\[
\epsilon_{rms} \approx \overline{\left|\eta_{1}\right|^{2}}^{1/2},
\]
where $\overline{()}$ denotes the ensemble average.  In our coordinates then, the work in \cite{alber} and later confirmed in \cite{dysthe2} shows that MI is suppressed for spectral widths $\sigma$ such that 
\[
\sigma \geq \epsilon_{rms}\sqrt{\frac{2\alpha_{nl}}{\alpha_{d}}}
\]
However, all of these results are for the case of zero vorticity and surface tension.

Thus, in this paper we establish the validity of the above bound on the spectral width for the appearance or suppression of MI.  We likewise examine the impact that vorticity plays in controlling the effective width for the onset of MI.  Moving beyond just examining the properties of the NLS equation with vorticity, we look at the statistical properties of solutions to a higher-order model, the vor-Dysthe equation derived in \cite{curtis8}.

\bibliography{deep_water_shear_Feb2}
\bibliographystyle{unsrt}
\end{document}

### 2018-08-30 nls_Dysthe_comparison.m

@@ -1,0 +1,67 @@
+function nls_Dysthe_comparison(Llx,K,ep,rms,tf,dt,om,sig,width,Nens)
+    
+    % [-Llx,Llx] is size of simulation
+    % K is number of modes in simulation
+    % ep is magnitude of perturbation in Dysthe equation
+    % tf is length of time to let simulation run
+    % dt is time step in RK4 solver for time simulation
+    % k0 is wave number of carrier wave 
+    % om is magnitude of vorticity 
+    % sig is surface tension
+    % Nens is number of ensemble members
+
+    disp('Spatial mesh size is')
+    disp(Llx/K)
+    
+    nmax = round(tf/dt); % Step size for time integrator and number of time steps
+    k0 = pi*ep/Llx*floor(Llx/(pi*ep));
+    disp('Carrier wave number is')
+    disp(k0)
+    
+    [Om,cg,ad,anl] = param_maker(k0,om,sig);    
+    
+    KT = 2*K;
+    
+    Kc = floor(KT/3);
+    Kuc = KT - Kc + 1;
+    Kc = Kc + 1;    
+    
+    Kvec = [ 0:K -K+1:-1 ]';
+    Kmesh = pi/Llx*Kvec;
+        
+    Xmesh = -Llx:Llx/K:Llx-Llx/K;
+    
+    uints = zeros(KT,Nens);
+    nls_sol = zeros(KT,Nens);
+    dysthe_sol = zeros(KT,Nens);
+    
+    rvec = exp(-(Kmesh).^2/(2*width^2));
+    
+    for jj=1:Nens
+        arand = 2*pi*1i*rand(KT,1);
+        uints(:,jj) = KT*rms*sqrt(sqrt(pi)/(Llx*width))*rvec.*exp(arand);    
+        uints(Kc:Kuc,jj) = 0;
+        %plot(Xmesh,abs(ifft(uints(:,jj))),'k','LineWidth',2)
+        %pause
+    end
+    
+    aval = rms*sqrt(sqrt(pi)/(Llx*width)*sum(rvec.^2));
+    disp('Root mean square NLS amplitude')
+    disp(aval)
+    disp('Minimal Stable Envelope Width')
+    disp(4*aval)
+    
+    parfor jj=1:Nens
[…]

### 2018-08-30 nls_solver.m

@@ -1,0 +1,50 @@
+function w = nls_solver(K,Llx,nmax,ad,anl,dt,uint)
+
+    % solves u_t = dk2Om*u_xx + a3*|u|^2*u 
+
+    % Llx is domain size, i.e. solve on domain [-Llx,Llx]
+    % sig is -1 for bright/focusing, +1 for dark/defocusing
+    % tf is total simulation time
+    % mu is magnitude of damping on envelope of dark initial condition
+   
+        Kvec = [ 0:K -K+1:-1 ]';
+        Kmesh = pi/Llx*Kvec;        
+        Dx = 1i*Kmesh; % Fourier transform of first derivative
+        Dx2 = Dx.^2; % Fourier transform of second derivative                       
+        KT = 2*K;
+        Kc = floor(KT/3);
+        Kuc = KT - Kc + 1;
+        Kc = Kc + 1;
+        
+        w = uint;             
+        
+    % Begin setup of RK4 method
+
+        Lap = 1i*dt*ad*Dx2;
+        E = exp(Lap/2);
+        E2 = E.^2;
+                
+    % Solve NLS equation in time
+    
+        for nn=1:nmax            
+            wp = ifft(w);
+            a = 1i*dt*anl*fft(wp.^2.*conj(wp));
+            a(Kc:Kuc) = 0;
+            
+            ap = ifft(E.*(w+a/2));
+            b = 1i*dt*anl*fft(ap.^2.*conj(ap));
+            b(Kc:Kuc) = 0;
+            
+            bp = ifft(E.*w + b/2);
+            c = 1i*dt*anl*fft(bp.^2.*conj(bp));
+            c(Kc:Kuc) = 0;
+            
+            cp = ifft(E2.*w + E.*c);
+            d = 1i*dt*anl*fft(cp.^2.*conj(cp));
+            d(Kc:Kuc) = 0;
+            
+            w = E2.*w + (E2.*a + 2*E.*(b+c) + d)/6;                                        
+            w(Kc:Kuc) = 0;
+        end
+        
+end

### 2018-08-30 nonlinearity.m

@@ -1,0 +1,29 @@
+function rhs = nonlinearity(eta1f,k0,cg,om,Om,ep,sig,ad,anl,Dx,L0i)
+
+s = sign(k0);
+
+n00 = om*(2*s*Om-om)/(1+om*cg);
+
+eta1p = ifft(eta1f);
+dxeta1p = ifft(Dx.*eta1f);
+meta1p = real(eta1p.*conj(eta1p));
+Hdxmeta1p = real(ifft(-abs(Dx).*fft(meta1p)));
+
+eta0p = n00*meta1p;
+dxeta0p = real(ifft(Dx.*fft(eta0p)));
+Hdxeta0p = real(ifft(-abs(Dx).*fft(eta0p)));
+eta11p = cg*(2*s*Om-om)/(1+om*cg)^2*Hdxmeta1p + 1i*om/(1+om*cg)*(ad*om*(2*s*Om-om)/(1+om*cg) + s*cg)*(eta1p.*conj(dxeta1p)-conj(eta1p).*dxeta1p);
+
+ellz = -k0*(om^2+2*Om^2-4*s*om*Om)/(2*(k0+om*Om-2*s*Om^2+4*sig*k0^3));
+ello = (2*cg*k0*(Om-s*om)-2*s*om*Om+Om^2+om^2/2 + ellz*(1+om*cg+12*k0^2*sig-4*s*cg*Om))/(k0+om*Om-2*s*Om^2+4*sig*k0^3);
+
+a1 = (k0*(4*cg*s*k0*Om-9*sig*k0^3+6*s*Om^2-3*s*om^2) + ellz*(2*om^2+4*Om^2-8*s*om*Om+4*k0*cg*(Om-s*om)) + ello*k0*(4*s*om*Om-2*Om^2-om^2))/(2*s*Om-om);
+a2 = (k0*(-3/2*k0^3*sig+2*s*Om^2-s*om^2) + ellz*(om^2+2*Om^2-4*s*om*Om))/(2*s*Om-om);
+a3 = k0*(2*s*Om-om);
+
+fvec = -1i*k0*om*eta1p.*eta11p + a1*meta1p.*dxeta1p + a2*eta1p.^2.*conj(dxeta1p) + 1i*a3*eta1p.*Hdxmeta1p ... 
+       + om*(om-2*s*Om-s*k0*cg)/(2*s*Om-om)*eta1p.*dxeta0p ...
[…]

### 2018-08-30 vor_Dysthe_solver.m

@@ -1,0 +1,51 @@
+function w = vor_Dysthe_solver(K,Llx,nmax,ad,anl,cg,k0,Om,om,sig,ep,dt,uint)
+
+    % solves u_t = dk2Om*u_xx + a3*|u|^2*u 
+
+    % Llx is domain size, i.e. solve on domain [-Llx,Llx]
+    % sig is -1 for bright/focusing, +1 for dark/defocusing
+    % tf is total simulation time
+    % mu is magnitude of damping on envelope of dark initial condition
+    Kvec = [ 0:K -K+1:-1 ]';
+    Kmesh = pi/Llx*Kvec;    
+    Dx = 1i*Kmesh;
+    Dx2 = Dx.^2;
+    s = sign(k0);
+    %L0i = (1 + 2*1i*ep*s*cg/(om-2*s*Om)*Dx).^(-1);
+    L0i = (1 - 2*1i*ep*s*cg/(om-2*s*Om)*Dx);
+    %Lap = dt*L0i.*(1i*ad + ep*sig/(om-2*s*Om)*Dx).*Dx2;
+    Lap = dt*(L0i*1i*ad+ ep*sig/(om-2*s*Om)*Dx).*Dx2;
+    KT = 2*K;
+    Kc = floor(KT/3);
+    Kuc = KT - Kc + 1;
+    Kc = Kc + 1;
+    
+    % Bright/Focusing initial conditions       
+    w = uint; 
+        
+    % Begin setup of RK4 method
+    E = exp(Lap/2);
+    E2 = E.^2;
+                
+    % Solve NLS equation in time    
+    for nn=1:nmax            
+        a = dt*nonlinearity(w,k0,cg,om,Om,ep,sig,ad,anl,Dx,L0i);
+        af = E.*(w+a/2);
+        af(Kc:Kuc) = 0;
+        
+        b = dt*nonlinearity(af,k0,cg,om,Om,ep,sig,ad,anl,Dx,L0i);
+        bf = E.*w + b/2;
+        bf(Kc:Kuc) = 0;
+        
+        c = dt*nonlinearity(bf,k0,cg,om,Om,ep,sig,ad,anl,Dx,L0i);
+        cf = E2.*w + E.*c;
+        cf(Kc:Kuc) = 0;
+        
+        d = dt*nonlinearity(cf,k0,cg,om,Om,ep,sig,ad,anl,Dx,L0i);
+        d(Kc:Kuc) = 0;
[…]

### 2018-09-20 random_vorticity.tex
Code Update

@@
 \section*{Introduction}
+
+Building off the now seminal work in \cite{alber}, there is a relatively wide range of literature which shows that nonlinear instabilities, in particular the modulational instability (MI), are responsible for significant modifications to the statistical properties of water waves; see \cite{dysthe2,dysthe3,onorato,thomas2012nonlinear} among others.
+These results have potentially significant impact on the understanding of rogue wave formation among other oceanographic phenomena.
+However, aside from
+
 To describe the evolution of a nearly linear wave-train in a constant-vorticity-shear current running over infinitely deep water, we use the form of the nonlinear Schr\"{o}dinger equation derived in \cite{curtis8}, though see also \cite{thomas2012nonlinear}, given by
@@
 where $\tau = \epsilon^{2}t$, $\xi = \epsilon(x+c_{g}t)$, and 
-\[
-c_{g} = \frac{1+3\tilde{\sigma}k_{0}^{2}}{2s\Omega - \omega}, ~ s = \mbox{sgn}(k_{0}),
-\]
 \begin{align*}
+c_{g} = & \frac{1+3\tilde{\sigma}k_{0}^{2}}{2s\Omega - \omega}\\
 \alpha_{d}(k_{0},\omega) = & \frac{(c^2_{g} - 3|k_{0}|\tilde{\sigma})}{2\Omega-s\omega},\\
@@
 \end{align*}
-where $\Omega$ denotes the dispersion relationship and $\omega$ is the non-dimensionalized magnitude of the vorticity of the flow, and $\tilde{\sigma}$ is the surface tension.
+where $s = \mbox{sgn}(k_{0})$, $\Omega$ denotes the dispersion relationship, and $\omega$ is the non-dimensionalized magnitude of the vorticity of the flow, and $\tilde{\sigma}$ is the surface tension.
 In a now seminal paper, \cite{alber} studies families of profiles with relatively narrow banded spectra around $k_{0}$ and modulational instability (MI) is either manifested or supressed.
@@
 \]
-However, all of these results are for the case of zero vorticity and surface tension.
+This work has helped motivate defining this parameter as the Benjamin--Feir Index (BFI), which has been shown to be a critical parameter to understanding the statstical properties of nonlinear free-surface flows; \cite{onorato, thomas2012nonlinear}.
 
-Thus, in this paper we establish the validity of the above bound on the spectral width for the appearance or suppression of MI.
+In part then, in this paper we establish the validity of the above bound on the spectral width for the appearance or suppression of MI.
 We likewise examine the impact that vorticity plays in controlling the effective width for the onset of MI.
 Moving beyond just examining the properties of the NLS equation with vorticity, we look at the statistical properties of solutions to a higher-order model, the vor-Dysthe equation derived in \cite{curtis8}.
+As shown in \cite{onorato}, the higher-order terms associated with the Dysthe equation can have significant impacts on the statistics of the waves, so it is a nontrivial question to examine how the vorticity will interact with these higher-order nonlinearities.
 

### 2018-09-20 nls_Dysthe_comparison.m

@@ -1,4 +1,4 @@
-function nls_Dysthe_comparison(Llx,K,ep,rms,tf,dt,om,sig,width,Nens)
+function nls_Dysthe_comparison(Llx,K,ep,tf,dt,om,sig,width,Nens)
     
     % [-Llx,Llx] is size of simulation
     % K is number of modes in simulation
@@ -10,11 +10,13 @@ function nls_Dysthe_comparison(Llx,K,ep,rms,tf,dt,om,sig,width,Nens)
     % sig is surface tension
     % Nens is number of ensemble members
 
+    Llx = ep*Llx;
+    
     disp('Spatial mesh size is')
     disp(Llx/K)
     
     nmax = round(tf/dt); % Step size for time integrator and number of time steps
-    k0 = pi*ep/Llx*floor(Llx/(pi*ep));
+    k0 = 1;
     disp('Carrier wave number is')
     disp(k0)
     
@@ -39,13 +41,13 @@ function nls_Dysthe_comparison(Llx,K,ep,rms,tf,dt,om,sig,width,Nens)
     
     for jj=1:Nens
         arand = 2*pi*1i*rand(KT,1);
-        uints(:,jj) = KT*rms*sqrt(sqrt(pi)/(Llx*width))*rvec.*exp(arand);    
+        uints(:,jj) = KT*sqrt(sqrt(pi)/(Llx*width))*rvec.*exp(arand);    
         uints(Kc:Kuc,jj) = 0;
         %plot(Xmesh,abs(ifft(uints(:,jj))),'k','LineWidth',2)
         %pause
     end
     
-    aval = rms*sqrt(sqrt(pi)/(Llx*width)*sum(rvec.^2));
+    aval = sqrt(sqrt(pi)/(Llx*width)*sum(rvec.^2));
     disp('Root mean square NLS amplitude')
     disp(aval)
     disp('Minimal Stable Envelope Width')

### 2018-09-21 vor_Dysthe_solver.m
Code Update

@@ -29,19 +29,19 @@ function w = vor_Dysthe_solver(K,Llx,nmax,ad,anl,cg,k0,Om,om,sig,ep,dt,uint)
                 
     % Solve NLS equation in time    
     for nn=1:nmax            
-        a = dt*nonlinearity(w,k0,cg,om,Om,ep,sig,ad,anl,Dx,L0i);
+        a = dt*nonlinearity(w,k0,cg,om,Om,ep,sig,ad,anl,Dx);
         af = E.*(w+a/2);
         af(Kc:Kuc) = 0;
         
-        b = dt*nonlinearity(af,k0,cg,om,Om,ep,sig,ad,anl,Dx,L0i);
+        b = dt*nonlinearity(af,k0,cg,om,Om,ep,sig,ad,anl,Dx);
         bf = E.*w + b/2;
         bf(Kc:Kuc) = 0;
         
-        c = dt*nonlinearity(bf,k0,cg,om,Om,ep,sig,ad,anl,Dx,L0i);
+        c = dt*nonlinearity(bf,k0,cg,om,Om,ep,sig,ad,anl,Dx);
         cf = E2.*w + E.*c;
         cf(Kc:Kuc) = 0;
         
-        d = dt*nonlinearity(cf,k0,cg,om,Om,ep,sig,ad,anl,Dx,L0i);
+        d = dt*nonlinearity(cf,k0,cg,om,Om,ep,sig,ad,anl,Dx);
         d(Kc:Kuc) = 0;
         
         w = E2.*w + (E2.*a + 2*E.*(b+c) + d)/6;    

### 2018-09-22 nls_Dysthe_comparison.m
Update to T&D scalings

@@ -39,12 +39,13 @@ function nls_Dysthe_comparison(Llx,K,ep,tf,dt,om,sig,width,Nens)
     
     rvec = exp(-(Kmesh).^2/(2*width^2));
     kvec = pi/Llx*(-K+1:K)';
-    deadinds = abs(ifftshift(kvec)) > 3/sqrt(2)*width;
+    %deadinds = abs(ifftshift(kvec)) > sqrt(2)*width;
     
     for jj=1:Nens
         arand = 2*pi*1i*rand(KT,1);
         uints(:,jj) = KT*sqrt(sqrt(pi)/(Llx*width))*rvec.*exp(arand);    
-        uints(deadinds,jj) = 0;
+        %uints(deadinds,jj) = 0;
+        uints(Kc:Kuc,jj) = 0;
         %plot(Xmesh,abs(ifft(uints(:,jj))),'k','LineWidth',2)
         %pause
     end   
@@ -60,11 +61,16 @@ function nls_Dysthe_comparison(Llx,K,ep,tf,dt,om,sig,width,Nens)
         dysthe_sol(:,jj) = vor_Dysthe_solver(K,Llx,nmax,ad,anl,cg,k0,Om,om,sig,ep,dt,uints(:,jj));
     end
     
-    nls_spec_mean = fftshift(abs(mean(nls_sol.*conj(nls_sol),2))/KT^2);
-    dysthe_spec_mean = fftshift(abs(mean(dysthe_sol.*conj(dysthe_sol),2))/KT^2);
-    mean_int = fftshift(abs(mean(uints.*conj(uints),2))/KT^2);
+    nls_spec_mean = ep^2*fftshift(abs(mean(nls_sol.*conj(nls_sol),2))/KT^2);
+    dysthe_spec_mean = ep^2*fftshift(abs(mean(dysthe_sol.*conj(dysthe_sol),2))/KT^2);
+    mean_int = ep^2*fftshift(abs(mean(uints.*conj(uints),2))/KT^2);
     
     
-    plot(kvec,mean_int,'k-',kvec,nls_spec_mean,'k--',kvec,dysthe_spec_mean,'k-.','LineWidth',2)
+    plot(ep*kvec+k0,mean_int,'k-',ep*kvec+k0,nls_spec_mean,'k--',ep*kvec+k0,dysthe_spec_mean,'k-.','LineWidth',2)
+    h = set(gca,'FontSize',30);
+    set(h,'Interpreter','LaTeX')
[…]

### 2018-09-22 nls_Dysthe_comparison.m
Code Update

@@ -70,7 +70,7 @@ function nls_Dysthe_comparison(Llx,K,ep,tf,dt,om,sig,width,Nens)
     h = set(gca,'FontSize',30);
     set(h,'Interpreter','LaTeX')
     xlabel('$k$','Interpreter','LaTeX','FontSize',30)
-    %ylabel('$\left<\left|\hat{\eta}(k,t_{f})\right|^{2}\right>$','Interpreter','LaTeX','FontSize',30)
+    ylabel('$\left<\left|\hat{\eta}(k,t_{f})\right|^{2}\right>$','Interpreter','LaTeX','FontSize',30)
     legend({'$Initial$','$NLS$','$Dysthe$'},'Interpreter','LaTeX','FontSize',30)
 end
 

### 2019-03-08 Writeup/random_vorticity.tex
Mega Draft Update

 \documentclass[a4paper,11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{epstopdf}
\epstopdfsetup{update}

\newcommand{\ba}{\begin{array}}
\newcommand{\ea}{\end{array}}

\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\newcommand{\bc}{\begin{center}}
\newcommand{\ec}{\end{center}}

\newcommand{\ds}{\displaystyle}

\newcommand{\bt}{\begin{tabular}}
\newcommand{\et}{\end{tabular}}

\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}

\newcommand{\bd}{\begin{description}}
\newcommand{\ed}{\end{description}}

\newcommand{\bp}{\begin{pmatrix}}
\newcommand{\ep}{\end{pmatrix}}

\newcommand{\pd}{\partial}
\newcommand{\sech}{\mbox{sech}}

\newcommand{\cf}{{\it cf.}~}

\newcommand{\ltwo}{L_{2}(\mathbb{R}^{2})}
\newcommand{\smooth}{C^{\infty}_{0}(\mathbb{R}^{2})}

\newcommand{\br}{{\bf r}}
\newcommand{\bk}{{\bf k}}
\newcommand{\bv}{{\bf v}}

\newcommand{\gnorm}[1]{\left|\left| #1\right|\right|}
\newcommand{\ipro}[2]{\left<#1,#2 \right>}

\title{Evolution of Narrow-Band Spectra in Deep-Water Constant Vorticity Flows}
\author{Christopher W. Curtis and Mackensie Murphy}
\date{}
\begin{document}
\maketitle
\section*{Introduction}
There is now a relatively wide range of literature which shows that nonlinear instabilities, in particular the modulational instability (MI), are responsible for significant modifications to the statistical properties of water waves; see \cite{dysthe2,dysthe3,onorato,thomas2012nonlinear} among others.   These results have significant impact on the understanding of deep-water rogue wave formation among other oceanographic phenomena.  However, much of our understanding around MI relies on highly idealized assumptions which cannot be expected to hold in natural settings.  In particular, most understandings of MI rely on looking at the stability of perturbations to carrier waves at essentially a fixed wave number.  To address this shortcoming, in a now seminal paper, \cite{alber} analytically studies families of perturbations of wave packets with narrow, but non-zero, spectral width around a central wave number.  An analytic critierion determining when modulational instability (MI) is either manifested or supressed depending on this spectral width is derived.  The results in \cite{alber} were confirmed numerically in \cite{dysthe2,dysthe3} by examing the mean properties associated with ensembles of initial conditions.

However, many physical effects were ignored in the derivation and verification of the results described above. In particular, vorticity was not taken into account, thereby ignoring a central mechanism for the transfer of energy across length scales in oceanic flows.  Thus, building on the work in \cite{thomas2012nonlinear} and \cite{curtis8}, we numerically explore the analytic stability condition derived in \cite{alber}, focusing in particular on the impact that vorticity plays in controlling the effective width for the onset of MI.  Moving beyond just examining the properties of the NLS equation with vorticity, we look at the statistical properties of solutions to a higher-order model, the vor-Dysthe equation derived in \cite{curtis8}.  As shown in \cite{onorato}, the higher-order terms associated with the Dysthe equation can have significant impacts on the statistics of the waves, so it is a nontrivial question to examine how the vorticity will interact with these higher-order nonlinearities.

\section*{Modeling and Stability Theory}
In non-dimensional variables, we model the evolution of a relatively small free-fluid surface over infinitely-deep water $z=\epsilon \eta(x,t)$ via the expansion 
\begin{multline*}
\eta(x,t) = \eta_{1}(\xi,\tau)e^{i\theta(x,t)} + \eta^{\ast}_{1}(\xi,\tau)e^{-i\theta(x,t)}\\
+ \epsilon\left(\eta_{0}(\xi,\tau) + \eta_{2}(\xi,\tau)e^{2i\theta(x,t)} + \eta^{\ast}_{2}(\xi,\tau)e^{-2i\theta(x,t)}\right) + \mathcal{O}\left(\epsilon^{2} \right)
\end{multline*}
where $\tau = \epsilon^{2}t$, $\xi = \epsilon(x+c_{g}t)$, and $\theta(x,t) = k_{0}x + \Omega(k_{0},\omega)t$, where the linear dispersion relationship $\Omega(k_{0},\omega)$ is given by
\[
\Omega(k_{0},\omega) = \frac{1}{2}\left(s\omega \pm \sqrt{\omega^{2} + 4|k_{0}|\left(1+\tilde{\sigma}k_{0}^{2}\right)} \right).
\]
Here, $s = \mbox{sgn}(k_{0})$, $\omega$ is the non-dimensionalized magnitude of the vorticity of the flow, and $\tilde{\sigma}$ is the surface tension.

To describe the evolution of the slowly evolving envelope $\eta_{1}(\xi,\tau)$, we use two asymptotically accurate models.  The first model is the the nonlinear Schr\"{o}dinger equation (NLSE) derived in \cite{curtis8}, though see also \cite{thomas2012nonlinear}, given by 
\[
i\pd_{\tau}\eta_{1} + \alpha_{d}\pd_{\xi}^{2}\eta_{1} + \alpha_{nl}\left|\eta_{1} \right|^{2}\eta_{1} = 0, 
\]
where 
\begin{align*}
c_{g} = & \frac{1+3\tilde{\sigma}k_{0}^{2}}{2s\Omega - \omega},\\
\alpha_{d}(k_{0},\omega) = & \frac{(c^2_{g} - 3|k_{0}|\tilde{\sigma})}{2\Omega-s\omega},\\
\alpha_{nl}(k_{0},\omega) = & \frac{k_{0}\left( sk_{0}^{3}\left(8 + \tilde{\sigma}k_{0}^{2} + 2(\tilde{\sigma}k_{0}^{2})^{2}\right) + \omega \alpha_{v}\right)}{\left(2s\Omega -\omega\right)(1+c_{g}\omega)\left(4\Omega^2-s(2k_{0}(1+4\tilde{\sigma}k_{0}^{2})+2\omega\Omega)\right)}.
\end{align*}

The second model we study in this paper is a higher-order generalization of the NLSE which we call the vor-Dysthe Equation (VDE) which was derived in \cite{curtis8}.  In this paper, we use it in the form 
\[
i\pd_{\tau}\eta_{1} + \alpha_{d}\pd_{\xi}^{2}\eta_{1} + \alpha_{nl}\left|\eta_{1} \right|^{2}\eta_{1} - \epsilon\frac{i(2sc_{g}\alpha_{d}+\tilde{\sigma})}{\omega-2s\Omega}\pd^{3}_{\xi}\eta_{1} + i\epsilon\mathcal{N}(\eta_{1}) = 0, 
\]
where
\[
\mathcal{N}(\eta_{1}) = \tilde{\alpha}_{1}\left|\eta_{1}\right|^{2}\pd_{\xi}\eta_{1} + \tilde{\alpha}_{2} \eta_{1}^{2}\pd_{\xi}\eta^{\ast}_{1} + i\tilde{\alpha}_{3}\eta_{1}\mathcal{H}\pd_{\xi}\left| \eta_{1}\right|^{2} + \tilde{\alpha}_{4}\pd_{\xi}\left| \eta_{1}\right|^{2}\eta_{1},
\]
and $\mathcal{H}$ is the Hilbert transform.  The terms $\tilde{\alpha}_{j}$ are straightforward to derive with the aid of a program like SAGE, but they are lengthy and not illuminative, and thus they are ommitted for the sake of brevity.

To generate simulations, we start both the NLSE and VDE from the same collection of $2L$ periodic-initial conditions
\[
\eta_{1}(\xi,0) = \epsilon_{rms}\sqrt{\frac{\delta \tilde{k}}{\sigma\sqrt{\pi}}} \sum_{k=-K+1}^{K}e^{-\tilde{k}^{2}/2\sigma^{2}}e^{i\theta_{k}}e^{i\tilde{k}\xi}, ~ \tilde{k} = \frac{\pi k}{L},
\]
where $\delta \tilde{k}=\pi/L$ and the phases $\theta_{k}$ are randomly chosen uniformly between $0$ and $2\pi$.  We can then readily show that, for $L\gg 1$, 
\[
\epsilon_{rms} \approx \overline{\left|\eta_{1}\right|^{2}}^{1/2},
\]
where $\overline{()}$ denotes the ensemble average.  In our coordinates then, the work in \cite{alber} and later confirmed in \cite{dysthe2} shows that for the NLSE the MI is suppressed for spectral widths $\sigma$ such that 
\[
\sigma \geq \epsilon_{rms}\sqrt{\frac{2\alpha_{nl}}{\alpha_{d}}}
\]
This work has helped motivate defining this parameter as the Benjamin--Feir Index (BFI), which has been shown to be a critical parameter to understanding the statstical properties of nonlinear free-surface flows; see \cite{onorato, thomas2012nonlinear}.
\section*{Results}
To understand the relationship between the vorticity $\omega$ and the initial spectral width $\sigma$ of our initial conditions, we generate ensembles of initial conditions and then run these different initial conditions up to $\tau=1$, which corresponds to $t_{f} = 1/\epsilon^{2}$ in the original fast variables.  We then plot the averaged spectral density $S(k,t_{f};\sigma)$ where 
\[
S\left(k,t_{f};\sigma\right) = \overline{\left( \left|\frac{1}{\epsilon}\hat{\eta}_{1}\left(\frac{k-k_{0}}{\epsilon},1;\sigma \right)\right|^{2}\right)}.
\]
which corresponds to the autocorrelation of the surface profile with respect to space.  Note we have emphasized the dependence on the initial spectral width $\sigma$ for the sake of clarity.  Following the approach in \cite{dysthe2,dysthe3}, by comparing the initial spectral density to that at later times, we can see directly whether MI manifests as significant distortions of the statistics, or if through supression the distributions remain relatively close to one another.

Throughout our results, we have used a psuedo-spectral method in space coupled with a 4th-order Runge-Kutta scheme using integrating factors in time.  De-aliasing is implemented via Orszag's `2/3-rule'.  Averaging was done with $N_{e}=3200$ ensemble members, thereby ensuring relative averaging errors on the order of about $1\%$.  Throughout this section, we present comparisons of the distributions for particular choices of $\sigma$ while also comparing the mean and standard deviations of the respective distributions so as to get a more quantitative understanding of the differences between the different distributions.

\subsection*{Strong-Counter Shear: $\omega = -1$}

Motivated by the results in \cite{curtis8}, we anticipate that $\omega=-1$ presents the most case which would allow for the vorticity to have the strongest impact on the BFI condition, which for the parameter choices requires that $\sigma > 11.9$.  This is borne out by numerical experiment as seen in Figures \ref{fig:meanstdomn1} and \ref{fig:pspecomn1}.  In this case, we see that for the NLSE the BFI condition largely holds.  In particular, we see that while the mean never changes in any significant way from that of the initial distribution, increasing the width $\sigma$ causes the standard deviation of the distribution determined via evolution of the NLSE to collapse onto that of the initial, thereby showing that wider spreads in initial conditions are overall less susceptible to MI.

\begin{figure}[!ht]
\centering
\begin{tabular}{c}
\includegraphics[width=.85\textwidth]{omega_n1_ep0pt1_dt5en4}\\
(a) $\epsilon=.1$ \\
\includegraphics[width=.85\textwidth]{omega_n1_ep0pt04}\\
(b) $\epsilon=.04$
\end{tabular}
\caption{For $\omega=-1$, comparison of the means and standard deviations between the initial averaged spectral density and those generated by the NLSE and VDE.  This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).  The vertical line denotes the spectral width corresponding to the BFI condition.}
\label{fig:meanstdomn1}
\end{figure}

\begin{figure}[!ht]
\centering
\begin{tabular}{c}
\includegraphics[width=.65\textwidth]{omega_n1_ep0pt1_dt5en4_width12} \\
(a) $\epsilon=.1$\\
\includegraphics[width=.65\textwidth]{omega_n1_ep0pt04_width12}\\
(b) $\epsilon=.04$ 
\end{tabular}
\caption{For $\omega=-1$, comparison of the averaged spectral densities between the initial averaged spectral density and those generated by the NLSE and VDE.  This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).}
\label{fig:pspecomn1}
\end{figure}

However, the results for the VDE are markedly different.  As seen in Figure \ref{fig:pspecomn1}, where $\sigma=12$ so that we are just past the BFI threshold, the most peculiar feature of the VDE is the bimodality of the distribution.  There is clearly a frequency downshift of the central initial peak, but a secondary peak at high-wave number forms as well.  This formation is less pronounced with decreasing $\epsilon$, but is clearly there nevertheless.  This represents a kind of resonance condition for which no ready explanation exists. This feature is likewise responsible for the strong skewing of the means and standard deviations seen in Figure \ref{fig:meanstdomn1}.

\subsection*{Vorticity-Free Flow: $\omega = 0$}

Removing vorticity, we see that we essentially recreate the standard BFI, with BFI threshold $\sigma > 5.7$, results for the NLSE.  Again though, the VDE produces markedly different statistical results from the NLS, though to nowhere near as drastic an extent as when $\omega=-1$.  In particular, the bimodality and strong deformation of the profile is notably ameliorated for $\epsilon=.04$; see Figures \ref{fig:meanstdom0} and \ref{fig:pspecom0}.  Thus in some sense, though not entirely, MI has more or less similar characteristics for both the NLSE and VDE.

\begin{figure}[!ht]
\centering
\begin{tabular}{c}
\includegraphics[width=.85\textwidth]{omega0_ep0pt1}\\
(a) $\epsilon=.1$ \\
\includegraphics[width=.85\textwidth]{omega0_ep0pt04}\\
(b) $\epsilon=.04$
\end{tabular}
\caption{For $\omega=0$, comparison of the means and standard deviations between the initial averaged spectral density and those generated by the NLSE and VDE.  This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).  The vertical line denotes the spectral width corresponding to the BFI condition.}
\label{fig:meanstdom0}
\end{figure}

\begin{figure}[!ht]
\centering
\begin{tabular}{c}
\includegraphics[width=.65\textwidth]{omega0_ep0pt1_width6} \\
(a) $\epsilon=.1$\\
\includegraphics[width=.65\textwidth]{omega0_ep0pt04_width6}\\
(b) $\epsilon=.04$ 
\end{tabular}
\caption{For $\omega=0$, comparison of the averaged spectral densities between the initial averaged spectral density and those generated by the NLSE and VDE.  This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).}
\label{fig:pspecom0}
\end{figure}

\subsection*{Strong-Co Shear: $\omega = 1$}

Finally, by letting $\omega=1$, we further reduce the BFI threshold to $\sigma > 2.7$, opening up a yet wider range of stable initial conditions.  Likewise, the VDE dynamics are far more in line with those seen from the NLSE, even in the case that $\epsilon=.1$; see Figures \ref{fig:meanstdom1} and \ref{fig:pspecom1}.  Therefore, we see how vorticity propagating in the correct direction can vastly reduce the impact of MI, even in a far more nonlinear model like the VDE.

\begin{figure}[!ht]
\centering
\begin{tabular}{c}
\includegraphics[width=.85\textwidth]{omega1_ep0pt1}\\
(a) $\epsilon=.1$ \\
\includegraphics[width=.85\textwidth]{omega1_ep0pt04}\\
(b) $\epsilon=.04$
\end{tabular}
\caption{For $\omega=1$, comparison of the means and standard deviations between the initial averaged spectral density and those generated by the NLSE and VDE.  This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).  The vertical line denotes the spectral width corresponding to the BFI condition.}
\label{fig:meanstdom1}
\end{figure}

\begin{figure}[!ht]
\centering
\begin{tabular}{c}
\includegraphics[width=.65\textwidth]{omega1_ep0pt1_width3} \\
(a) $\epsilon=.1$\\
\includegraphics[width=.65\textwidth]{omega1_ep0pt04_width3}\\
(b) $\epsilon=.04$ 
\end{tabular}
\caption{For $\omega=1$, comparison of the averaged spectral densities between the initial averaged spectral density and those generated by the NLSE and VDE.  This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).}
\label{fig:pspecom1}
\end{figure}

\section*{Conclusion}
\bibliography{deep_water_shear_Feb2}
\bibliographystyle{unsrt}
\end{document}

### 2019-03-20 Writeup/random_vorticity.tex
Added variables to Results

@@
 To understand the relationship between the vorticity $\omega$ and the initial spectral width $\sigma$ of our initial conditions, we generate ensembles of initial conditions and then run these different initial conditions up to $\tau=1$, which corresponds to $t_{f} = 1/\epsilon^{2}$ in the original fast variables.
+The domain is $[-50\pi,50\pi]$.
+The number of modes is $K = 256$.
+The surface tension is $\tilde{\sigma} = 10^{-5}$.
 We then plot the averaged spectral density $S(k,t_{f};\sigma)$ where
@@
 
-Throughout our results, we have used a psuedo-spectral method in space coupled with a 4th-order Runge-Kutta scheme using integrating factors in time.
+Throughout our results, we have used a psuedo-spectral method in space coupled with a 4th-order Runge-Kutta scheme using integrating factors in time, with a time step of $dt = 2.5 \times 10^{-3}$.
 De-aliasing is implemented via Orszag's `2/3-rule'.

### 2019-04-08 afm_dno_solver.m
Massive Additions

@@ -1,0 +1,70 @@
+function etanls = afm_dno_solver(K,k0,ep,Llx,sig,om,Om,tf,dt,uintsac)
+    
+    uintphys = ifft(uintsac);
+    etan = ep*fft(2*real(uintphys));
+    %etan(1) = 0;
+    Qn = ep*Om*fft(-2*real(uintphys));
+    Qn(1) = 0; %Enforce zero average in the surface velocity
+    
+    KT = 2*K;
+    % Find the wave numbers to implement the 2/3 de-aliasing throughout
+    
+    Kmesh = pi/Llx*[0:K -K+1:-1]';
+    Xmesh = (-Llx:Llx/K:Llx-Llx/K)';
+        
+    nmax = round(tf/dt);
+    
+    G0 = -1i*sign(Kmesh);
+    L1 = -1i*Kmesh.*(1+sig.*Kmesh.^2);
+    dti = 3*dt/4;
+    
+    Linvd = (1 + dti*om*G0 - dti^2*G0.*L1).^(-1);
+    Linv11 = (1 + om*dti*G0).*Linvd;
+    Linv12 = dti*G0.*Linvd;
+    Linv21 = dti*L1.*Linvd;
+    
+    etanm1 = etan;
+    Qnm1 = Qn;
+    
+    nln = dno_nonlinearity(K,etan,Qn,om,ep,sig,Kmesh);
+    
+    nlnm1 = nln;
+    nlnm2 = nlnm1;
+    nlnm3 = nlnm2;
+    
+    for jj=1:nmax
+        
+        nln = dno_nonlinearity(K,etan,Qn,om,ep,sig,Kmesh);
+        
+        nlvecn = 55/24*nln(1:KT) - 59/24*nlnm1(1:KT) + 37/24*nlnm2(1:KT) - 3/8*nlnm3(1:KT);    
+        nlvecq = 55/24*nln(KT+1:2*KT) - 59/24*nlnm1(KT+1:2*KT) + 37/24*nlnm2(KT+1:2*KT) - 3/8*nlnm3(KT+1:2*KT);
+        
+        etatot = etan + etanm1/3 + dt*nlvecn;
+        qtot = Qn + Qnm1/3 + dt*nlvecq;
+        
+        eta1 = Linv11.*etatot;
+        eta2 = Linv12.*qtot;
+        
+        Q1 = Linv21.*etatot;
+        Q2 = Linvd.*qtot;
+        
+        etanp1 = -etanm1/3 + eta1 + eta2;
[…]

### 2019-04-08 dno_nonlinearity.m

@@ -1,0 +1,40 @@
+function rhs = dno_nonlinearity(K,eta,Q,om,ep,sig,Kmesh)
+
+KT = 2*K;
+% Find the wave numbers to implement the 2/3 de-aliasing throughout
+Kc = floor(2*K/3);
+Kuc = KT-Kc+1;
+Kc = Kc+1;
+
+Dx = 1i*Kmesh;
+Dx2 = Dx.^2;
+Dx3 = Dx2.*Dx;
+Hop = 1i*sign(Kmesh);
+
+eta(Kc:Kuc)=0;
+Q(Kc:Kuc)=0;
+
+etax = real(ifft(Dx.*eta)); 
+etaxp2 = etax.*etax;
+etaxx = real(ifft(Dx2.*eta)); 
+etap = real(ifft(eta));
+etap2 = etap.^2;
+etap3 = etap2.*etap;
+
+Qp = real(ifft(Q));
+Qpn = Qp + om*etap;
+Qpn2 = Qpn.^2;
+
+G0 = -real(ifft(Hop.*Q));
+G1 = real(ifft( Dx.*( Hop.*fft(etap.*G0) - fft(om/2*etap.^2 + etap.*Qp) )));
+G2 = real(ifft( Dx.*Hop.*fft( etap.*G1 ) + Dx2.*( fft( etap2.*G0/2 ) + Hop.*fft(om/3*etap3 + etap2.*Qp/2) )));
+G3 = real(ifft( Dx.*Hop.*fft( etap.*G2 ) + Dx2.*fft(etap2.*G1/2) + Dx3.*( Hop.*fft( -etap3.*G0/6 )+ fft(om/8*etap3.*etap + etap3.*Qp/6) )));
+rhs1 = fft( ep*G1 + ep^2*G2 + ep^3*G3 );
+rhs2 = -om*rhs1 + Dx.*fft( ep/2*( -Qpn2 + G0.^2 + 2*ep*G0.*G1 + ep^2*(2*G0.*G2 + G1.^2) ) ...
+                         + ep^2*( etax.*Qpn.*(G0 + ep*G1) - 3/2*sig*etaxp2.*etaxx ) ...
+                         + ep^3*etaxp2.*( Qpn2 - G0.^2 )/2 );
+
+rhs1(Kc:Kuc) = 0;
+rhs2(Kc:Kuc) = 0;
+
+rhs = [rhs1;rhs2];

### 2019-04-08 full_simul_tester.m

@@ -1,0 +1,30 @@
+function full_simul_tester(K,ep,Llx,sig,om,tf,dt,width,Nens)
+
+    k0 = 10;
+    [Om,cg,ad,anl] = param_maker(k0,om,sig);    
+    Xmesh = -Llx:Llx/K:Llx-Llx/K;
+    KT = 2*K;
+    
+    Kc = floor(KT/3);
+    Kuc = KT - Kc + 1;
+    Kc = Kc + 1;    
+    
+    Kvec = [ 0:K -K+1:-1 ]';
+    Kmesh = pi/Llx*Kvec;
+    
+    rvecac = exp(-(Kmesh-k0).^2/(2*width^2*ep^2));
+    uintsac = zeros(KT,Nens);
+    
+    for jj=1:Nens
+        arand = exp(2*pi*1i*rand(KT,1));
+        uintsac(:,jj) = KT*sqrt(sqrt(pi)/(Llx*width*ep))*rvecac.*arand;   
+        uintsac(Kc:Kuc,jj) = 0;
+    end
+    
+    parfor jj=1:Nens
+        act_sol(:,jj) = afm_dno_solver(K,k0,ep,Llx,sig,om,Om,tf,dt,uintsac)
+    end
+    
+    figure(1)
+    plot(fftshift(Kmesh),ep^2*fftshift(abs(uintsac(:,1).*conj(uintsac(:,1)))),'k-',fftshift(Kmesh)+k0,fftshift(abs(act_sol(:,1).*conj(act_sol(:,1)))),'k--','LineWidth',2)
+    

### 2019-04-08 nls_Dysthe_comparison.m

@@ -9,7 +9,7 @@ function nls_Dysthe_comparison(Llx,K,ep,tf,dt,om,sig,width,Nens)
     % om is magnitude of vorticity 
     % sig is surface tension
     % Nens is number of ensemble members
-
+    Llxf = Llx;
     Llx = ep*Llx;
     
     disp('Spatial mesh size is')
@@ -17,6 +17,7 @@ function nls_Dysthe_comparison(Llx,K,ep,tf,dt,om,sig,width,Nens)
     
     nmax = round(tf/dt); % Step size for time integrator and number of time steps
     k0 = 1;
+    
     disp('Carrier wave number is')
     disp(k0)
     
@@ -30,24 +31,27 @@ function nls_Dysthe_comparison(Llx,K,ep,tf,dt,om,sig,width,Nens)
     
     Kvec = [ 0:K -K+1:-1 ]';
     Kmesh = pi/Llx*Kvec;
+    Kmeshf = pi/Llxf*Kvec;
         
     Xmesh = -Llx:Llx/K:Llx-Llx/K;
     
     uints = zeros(KT,Nens);
+    uintsac = zeros(KT,Nens);
     nls_sol = zeros(KT,Nens);
     dysthe_sol = zeros(KT,Nens);
+    act_sol = zeros(KT,Nens);
     
     rvec = exp(-(Kmesh).^2/(2*width^2));
+    rvecac = exp(-(Kmeshf-k0).^2/(2*width^2*ep^2));
     kvec = pi/Llx*(-K+1:K)';
     %deadinds = abs(ifftshift(kvec)) > sqrt(2)*width;
     
     for jj=1:Nens
-        arand = 2*pi*1i*rand(KT,1);
-        uints(:,jj) = KT*sqrt(sqrt(pi)/(Llx*width))*rvec.*exp(arand);    
-        %uints(deadinds,jj) = 0;
+        arand = exp(2*pi*1i*rand(KT,1));
+        uints(:,jj) = KT*sqrt(sqrt(pi)/(Llx*width))*rvec.*arand;   
+        uintsac(:,jj) = KT*sqrt(sqrt(pi)/(Llxf*width*ep))*rvecac.*arand;   
         uints(Kc:Kuc,jj) = 0;
-        %plot(Xmesh,abs(ifft(uints(:,jj))),'k','LineWidth',2)
-        %pause
[…]

### 2019-04-10 Writeup/random_vorticity.tex
Updates

@@
 \section*{Modeling and Stability Theory}
+We examine the unsteady nonlinear wave propagation over a constant shear current.  To do this, we assume the fluid velocity has the form
+\[
+{\bf u} = u({\bf x},t)\hat{{\bf x}} + w({\bf x},t)\hat{{\bf z}} = \omega z\hat{{\bf x}} + \nabla \phi,
+\]
+where $\phi$ is a harmonic function.
+We restrict fluid motion to the $(x,z)$-plane, thereby ignoring transverse variations in the $y$ dimension.
+Following standard arguments, e.g.~see \cite{ashton}, the dynamics of the fluid can be determined by solving the free boundary value problem
+\begin{align}
+\Delta \phi = 0, & & -\infty< z < \eta(x,t), \nonumber \\
+\eta_{t} +\left(\omega\eta+\phi_{x}\right)\eta_{x} - \phi_{z} = 0, & & z=\eta(x,t), \label{kine}\\
+\phi_{t} + \omega\pd_{x}^{-1}\eta_{t} + \frac{1}{2} \left|{\bf
+    u}\right|^{2} + g\eta -
+\frac{\sigma}{\rho}\pd_{x}\frac{\eta_{x}}{\sqrt{1+\eta_{x}^{2}}}= 0, &
+& z=\eta(x,t), \label{bernou}\\
+\lim_{z\rightarrow -\infty} \phi_{z} = 0, \nonumber
+\end{align}
+where $\eta$ represents the free surface displacement, and $g$, $\rho$, and $\sigma$ represent the acceleration due to gravity, the fluid density, and the coefficient of surface tension respectively.
+
+By choosing a characteristic wave height $a$ and wave length $L$, all quantities can be non-dimensionalized via
+\begin{align*}
+\tilde{x} = \frac{x}{L}, ~\tilde{z} = \frac{z}{L}, ~ \tilde{t} = \sqrt{\frac{g}{L}}~t, ~\omega = \sqrt{\frac{g}{L}}~\tilde{\omega}, \\
+\eta = a \tilde{\eta}, ~ \phi  = a\sqrt{gL}~\tilde{\phi} , ~ \epsilon = \frac{a}{L}.
+\end{align*}
+By following the AFM approach described in \cite{afm,ashton} and dropping the tildes, the kinematic boundary condition, Equation \eqref{kine}, can be written in terms of surface variables alone via the integro-differential equation
+\begin{equation}
+\int_{\mathbb{R}}dx~e^{-ikx}e^{\epsilon|k|\eta}\left(\eta_{t} +
+  \epsilon \omega \eta \eta_{x} + i \mbox{sgn}(k)Q \right) = 0, ~
+k\neq 0,
+\label{kine2}
+\end{equation}
+where $Q=q_{x}$, $q(x,t) = \phi(x,\eta(x,t),t)$.
+By integrating over $\mathbb{R}$, we are assuming that both $\eta$ and $Q$ decay to zero sufficiently rapidly in the far field.
+We can also readily derive a nearly identical expression on domains periodic in the horizontal variable $x$, in which case Equation \eqref{kine2} becomes
+\begin{equation}
+\int_{-L_{p}/2}^{L_{p}/2}dx~e^{-ikx}e^{\epsilon|k|\eta}\left(\eta_{t} +
+  \epsilon \omega \eta \eta_{x} + i \mbox{sgn}(k)Q \right) = 0, ~ k = \frac{2\pi m}{L_{p}}, ~ m\in \mathbb{Z}\backslash{0},
+\label{integro1per}
+\end{equation}
+where $L_p$ is the spatial period.  Throughout the remainder of this section we only present results over the real line, $\mathbb{R}$, since identical results can be derived for the periodic case.
+Lastly, we note that if in Equation \eqref{kine2} we approach $k=0$ from both the left and the right, we get the equations
+\[
+\int_{\mathbb{R}}dx \left(\eta_{t} + i Q \right) = 0, ~ \int_{\mathbb{R}}dx \left(\eta_{t} - i Q \right) = 0.
+\]
+Thus we get the identities
+\begin{align}
+\pd_{t}\int_{\mathbb{R}} dx ~\eta(x,t) & = 0, \label{zavsurf}\\
+\int_{\mathbb{R}} dx ~Q(x,t) & = 0. \label{zavpot}
+\end{align}
+
+Taylor expanding Equation (\ref{kine2}) up to $\mathcal{O}(\epsilon^3)$ for $k\neq 0$ gives
+\begin{multline}
+\int_{\mathbb{R}}dx~e^{-ikx}\left(1 + \epsilon |k|\eta + \frac{\epsilon^{2}|k|^{2}\eta^{2}}{2} \right)\left(\eta_{t} + i \mbox{sgn}(k)Q \right) \\
++ \epsilon \omega \int_{\mathbb{R}}dx~e^{-ikx}\left(1 + \epsilon |k|\eta \right) \eta \eta_{x} = 0.
+\label{integro1}
+\end{multline}
+Transforming into surface variables and Taylor expanding Equation \eqref{bernou}, Bernoulli's equation, up to $\mathcal{O}(\epsilon^3)$ gives
+\begin{multline}
+Q_{t} + \omega \eta_{t} + \eta_{x} - \tilde{\sigma}\eta_{xxx} + \frac{\epsilon}{2}\pd_{x}\left(-\eta_{t}^{2} + (Q+\omega \eta)^{2} \right)\\
++ \epsilon^{2}\pd_{x}\left( \frac{3}{2}\tilde{\sigma}\eta_{x}^{2}\eta_{xx} - \eta_{t}\eta_{x} \left(Q  + \omega \eta \right) \right)  = 0,
+\label{berexp}
+\end{multline}
+where the reciprocal of the Bond number, $\tilde{\sigma}$, is given by
+\[
+\tilde{\sigma} = \frac{\sigma}{\rho g L^2}.
+\]
+We note that we have tacitly assumed $\omega = \mathcal{O}(1)$.
+In physical terms, this implies that $\omega$ is comparable to the natural time scale of this problem, $\sqrt{L/g}$.
+If we were to assume $\omega$ were of larger magnitude, the problem would no longer be weakly nonlinear and would be much less amenable to asymptotic analysis.
+Therefore, throughout the remainder of the paper, we assume that the vorticity is not too large.
+
 In non-dimensional variables, we model the evolution of a relatively small free-fluid surface over infinitely-deep water $z=\epsilon \eta(x,t)$ via the expansion 

### 2019-04-10 afm_dno_solver.m

@@ -1,9 +1,9 @@
 function etanls = afm_dno_solver(K,k0,ep,Llx,sig,om,Om,tf,dt,uintsac)
     
     uintphys = ifft(uintsac);
-    etan = ep*fft(2*real(uintphys));
+    etan = fft(2*real(uintphys));
     %etan(1) = 0;
-    Qn = ep*Om*fft(-2*real(uintphys));
+    Qn = Om*fft(-2*real(uintphys));
     Qn(1) = 0; %Enforce zero average in the surface velocity
     
     KT = 2*K;

### 2019-04-10 dno_nonlinearity.m

@@ -8,7 +8,7 @@ Kc = Kc+1;
 
 Dx = 1i*Kmesh;
 Dx2 = Dx.^2;
-Dx3 = Dx2.*Dx;
+%Dx3 = Dx2.*Dx;
 Hop = 1i*sign(Kmesh);
 
 eta(Kc:Kuc)=0;
@@ -28,11 +28,16 @@ Qpn2 = Qpn.^2;
 G0 = -real(ifft(Hop.*Q));
 G1 = real(ifft( Dx.*( Hop.*fft(etap.*G0) - fft(om/2*etap.^2 + etap.*Qp) )));
 G2 = real(ifft( Dx.*Hop.*fft( etap.*G1 ) + Dx2.*( fft( etap2.*G0/2 ) + Hop.*fft(om/3*etap3 + etap2.*Qp/2) )));
-G3 = real(ifft( Dx.*Hop.*fft( etap.*G2 ) + Dx2.*fft(etap2.*G1/2) + Dx3.*( Hop.*fft( -etap3.*G0/6 )+ fft(om/8*etap3.*etap + etap3.*Qp/6) )));
-rhs1 = fft( ep*G1 + ep^2*G2 + ep^3*G3 );
-rhs2 = -om*rhs1 + Dx.*fft( ep/2*( -Qpn2 + G0.^2 + 2*ep*G0.*G1 + ep^2*(2*G0.*G2 + G1.^2) ) ...
-                         + ep^2*( etax.*Qpn.*(G0 + ep*G1) - 3/2*sig*etaxp2.*etaxx ) ...
-                         + ep^3*etaxp2.*( Qpn2 - G0.^2 )/2 );
+%G3 = real(ifft( Dx.*Hop.*fft( etap.*G2 ) + Dx2.*fft(etap2.*G1/2) + Dx3.*( Hop.*fft( -etap3.*G0/6 )+ fft(om/8*etap3.*etap + etap3.*Qp/6) )));
+%rhs1 = fft( ep*G1 + ep^2*G2 + ep^3*G3 );
+%rhs2 = -om*rhs1 + Dx.*fft( ep/2*( -Qpn2 + G0.^2 + 2*ep*G0.*G1 + ep^2*(2*G0.*G2 + G1.^2) ) ...
+%                         + ep^2*( etax.*Qpn.*(G0 + ep*G1) - 3/2*sig*etaxp2.*etaxx ) ...
[…]

### 2019-04-10 nls_Dysthe_comparison.m

@@ -61,21 +61,23 @@ function nls_Dysthe_comparison(Llx,K,ep,tf,dt,om,sig,width,Nens)
     disp(sqrt(2*anl/ad))
     
     parfor jj=1:Nens
-        nls_sol(:,jj) = nls_solver_slider(K,Llx,nmax,ad,anl,dt,uints(:,jj));
-        dysthe_sol(:,jj) = vor_Dysthe_solver_slider(K,Llx,nmax,ad,anl,cg,k0,Om,om,sig,ep,dt,uints(:,jj));
-        act_sol(:,jj) = afm_dno_solver(K,k0,ep,Llxf,sig,om,Om,tf/ep^2,1e2*dt,uintsac(:,jj))
+        nls_sol(:,jj) = nls_solver(K,Llx,nmax,ad,anl,dt,uints(:,jj));
+        % dysthe_sol(:,jj) = vor_Dysthe_solver(K,Llx,nmax,ad,anl,cg,k0,Om,om,sig,ep,dt,uints(:,jj));
+        act_sol(:,jj) = afm_dno_solver(K,k0,ep,Llxf,sig,om,Om,tf/ep^2,dt,uintsac(:,jj))
     end
     
-    nls_spec_mean = ep^2*fftshift(abs(mean(nls_sol.*conj(nls_sol),2))/KT^2);
-    dysthe_spec_mean = ep^2*fftshift(abs(mean(dysthe_sol.*conj(dysthe_sol),2))/KT^2);
-    mean_int = ep^2*fftshift(abs(mean(uints.*conj(uints),2))/KT^2);
-    mean_act = fftshift(abs(mean(act_sol.*conj(act_sol),2))/KT^2);
+    nls_spec_mean = (ep/KT)^2*fftshift(abs(mean(nls_sol.*conj(nls_sol),2)));
+    % dysthe_spec_mean = ep^2*fftshift(abs(mean(dysthe_sol.*conj(dysthe_sol),2))/KT^2);
+    mean_int = (ep/KT)^2*fftshift(abs(mean(uints.*conj(uints),2)));
+    mean_act = (ep/KT)^2*fftshift(abs(mean(act_sol.*conj(act_sol),2)));
     
-    plot(ep*kvec+k0,mean_int,'k-',ep*kvec+k0,nls_spec_mean,'k--',ep*kvec+k0,dysthe_spec_mean,'k-.',ep*kvec+k0,mean_act,'k:','LineWidth',2)
[…]

### 2019-04-10 Writeup/random_vorticity.tex
Quick tweak to TeX File

@@
 
-In non-dimensional variables, we model the evolution of a relatively small free-fluid surface over infinitely-deep water $z=\epsilon \eta(x,t)$ via the expansion 
+By expanding Equation \eqref{integro1} and taking inverse Fourier transforms, we can derive the first three terms of the corresponding Dirichlet-to-Neumann Operator (DNO) expansion, so that we find
+\begin{equation}
+\eta_{t} = G_{0}Q + \epsilon G_{1}Q + \epsilon^{2}G_{2}Q,
+\label{dnoexp}
+\end{equation}
+where
+\begin{align*}
+G_{0}Q = & -\mathcal{H}Q\\
+G_{1}Q = & \mathcal{H}\pd_{x}\left(\eta G_{0}Q\right) - \pd_{x}\left(\eta Q + \frac{\omega}{2}\eta^{2}\right)\\
+G_{2}Q = & \mathcal{H}\pd_{x}\left(\eta G_{1}Q\right) + \frac{1}{2}\pd_{x}^{2}\left(\eta^{2}G_{0}Q \right) + \pd_{x}^{2}\mathcal{H}\left(\frac{1}{2}\eta^{2}Q + \frac{\omega}{3}\eta^{3}\right), 
+\end{align*}
+where $\mathcal{H}$ denotes the Hilbert transform so that 
+\[
+\mathcal{H}f = \frac{1}{2\pi}\int_{\mathbb{R}}dk~ e^{ikx}  i\mbox{sgn}(k) \hat{f}(k),
+\]
+where $\hat{f}(k)$ is the Fourier transform of $f(x)$ defined by
+\[
+\hat{f}(k) = \int_{\mathbb{R}}dx~ e^{-ikx}f(x).
+\]
+Using the DNO expansion in Equation \eqref{dnoexp}, we can rewrite Equation \eqref{berexp} an an evolution equation for $Q_{t}$, thereby giving us a closed system in terms of surface variables alone.
+The system of equations given by Equations \eqref{dnoexp} and \eqref{berexp} we denote as the Full-Euler Equations (FEEs).
+
+As is well known by now, in non-dimensional variables, we can model the evolution of a relatively small free-fluid surface over infinitely-deep water $z=\epsilon \eta(x,t)$ via the expansion 
 \begin{multline*}
@@
 
-The second model we study in this paper is a higher-order generalization of the NLSE which we call the vor-Dysthe Equation (VDE) which was derived in \cite{curtis8}.
-In this paper, we use it in the form
-\[
-i\pd_{\tau}\eta_{1} + \alpha_{d}\pd_{\xi}^{2}\eta_{1} + \alpha_{nl}\left|\eta_{1} \right|^{2}\eta_{1} - \epsilon\frac{i(2sc_{g}\alpha_{d}+\tilde{\sigma})}{\omega-2s\Omega}\pd^{3}_{\xi}\eta_{1} + i\epsilon\mathcal{N}(\eta_{1}) = 0,
-\]
-where
-\[
-\mathcal{N}(\eta_{1}) = \tilde{\alpha}_{1}\left|\eta_{1}\right|^{2}\pd_{\xi}\eta_{1} + \tilde{\alpha}_{2} \eta_{1}^{2}\pd_{\xi}\eta^{\ast}_{1} + i\tilde{\alpha}_{3}\eta_{1}\mathcal{H}\pd_{\xi}\left| \eta_{1}\right|^{2} + \tilde{\alpha}_{4}\pd_{\xi}\left| \eta_{1}\right|^{2}\eta_{1},
-\]
-and $\mathcal{H}$ is the Hilbert transform.
-The terms $\tilde{\alpha}_{j}$ are straightforward to derive with the aid of a program like SAGE, but they are lengthy and not illuminative, and thus they are ommitted for the sake of brevity.
-
-To generate simulations, we start both the NLSE and VDE from the same collection of $2L$ periodic-initial conditions
+To generate simulations, we start both the NLSE and FEE from the same collection of $2L$ periodic-initial conditions
 \[

### 2019-04-10 nls_Dysthe_comparison.m

@@ -59,13 +59,16 @@ function nls_Dysthe_comparison(Llx,K,ep,tf,dt,om,sig,width,Nens)
     disp(aval)
     disp('Minimal Stable Envelope Width')
     disp(sqrt(2*anl/ad))
-    
+    tic
+    %Since these work over such different time scales.  
     parfor jj=1:Nens
         nls_sol(:,jj) = nls_solver(K,Llx,nmax,ad,anl,dt,uints(:,jj));
-        % dysthe_sol(:,jj) = vor_Dysthe_solver(K,Llx,nmax,ad,anl,cg,k0,Om,om,sig,ep,dt,uints(:,jj));
-        act_sol(:,jj) = afm_dno_solver(K,k0,ep,Llxf,sig,om,Om,tf/ep^2,dt,uintsac(:,jj))
+        % dysthe_sol(:,jj) = vor_Dysthe_solver(K,Llx,nmax,ad,anl,cg,k0,Om,om,sig,ep,dt,uints(:,jj));        
+    end    
+    parfor jj=1:Nens
+       act_sol(:,jj) = afm_dno_solver(K,k0,ep,Llxf,sig,om,Om,tf/ep^2,dt,uintsac(:,jj)) 
     end
-    
+    toc
     nls_spec_mean = (ep/KT)^2*fftshift(abs(mean(nls_sol.*conj(nls_sol),2)));
     % dysthe_spec_mean = ep^2*fftshift(abs(mean(dysthe_sol.*conj(dysthe_sol),2))/KT^2);
     mean_int = (ep/KT)^2*fftshift(abs(mean(uints.*conj(uints),2)));

### 2019-05-30 dno_nonlinearity.m
Getting there finally

@@ -8,7 +8,7 @@ Kc = Kc+1;
 
 Dx = 1i*Kmesh;
 Dx2 = Dx.^2;
-%Dx3 = Dx2.*Dx;
+Dx3 = Dx2.*Dx;
 Hop = 1i*sign(Kmesh);
 
 eta(Kc:Kuc)=0;
@@ -28,15 +28,15 @@ Qpn2 = Qpn.^2;
 G0 = -real(ifft(Hop.*Q));
 G1 = real(ifft( Dx.*( Hop.*fft(etap.*G0) - fft(om/2*etap.^2 + etap.*Qp) )));
 G2 = real(ifft( Dx.*Hop.*fft( etap.*G1 ) + Dx2.*( fft( etap2.*G0/2 ) + Hop.*fft(om/3*etap3 + etap2.*Qp/2) )));
-%G3 = real(ifft( Dx.*Hop.*fft( etap.*G2 ) + Dx2.*fft(etap2.*G1/2) + Dx3.*( Hop.*fft( -etap3.*G0/6 )+ fft(om/8*etap3.*etap + etap3.*Qp/6) )));
-%rhs1 = fft( ep*G1 + ep^2*G2 + ep^3*G3 );
-%rhs2 = -om*rhs1 + Dx.*fft( ep/2*( -Qpn2 + G0.^2 + 2*ep*G0.*G1 + ep^2*(2*G0.*G2 + G1.^2) ) ...
-%                         + ep^2*( etax.*Qpn.*(G0 + ep*G1) - 3/2*sig*etaxp2.*etaxx ) ...
-%                         + ep^3*etaxp2.*( Qpn2 - G0.^2 )/2 );
+G3 = real(ifft( Dx.*Hop.*fft( etap.*G2 ) + Dx2.*fft(etap2.*G1/2) + Dx3.*( Hop.*fft( -etap3.*G0/6 )+ fft(om/8*etap3.*etap + etap3.*Qp/6) )));
+rhs1 = fft( ep*G1 + ep^2*G2 + ep^3*G3 );
+rhs2 = -om*rhs1 + Dx.*fft( ep/2*( -Qpn2 + G0.^2 + 2*ep*G0.*G1 + ep^2*(2*G0.*G2 + G1.^2) ) ...
+                         + ep^2*( etax.*Qpn.*(G0 + ep*G1) - 3/2*sig*etaxp2.*etaxx ) ...
[…]

### 2019-05-30 nls_Dysthe_comparison.m

@@ -1,4 +1,4 @@
-function nls_Dysthe_comparison(Llx,K,ep,tf,dt,om,sig,width,Nens)
+function [nls_avg,std_nls,dys_avg,std_dys,int_avg,std_int] = nls_Dysthe_comparison(Llx,K,ep,tf,dt,om,sig,width,k0,Nens)
     
     % [-Llx,Llx] is size of simulation
     % K is number of modes in simulation
@@ -12,14 +12,13 @@ function nls_Dysthe_comparison(Llx,K,ep,tf,dt,om,sig,width,Nens)
     Llxf = Llx;
     Llx = ep*Llx;
     
-    disp('Spatial mesh size is')
-    disp(Llx/K)
+    %disp('Spatial mesh size is')
+    %disp(Llx/K)
     
     nmax = round(tf/dt); % Step size for time integrator and number of time steps
-    k0 = 1;
     
-    disp('Carrier wave number is')
-    disp(k0)
+    %disp('Carrier wave number is')
+    %disp(k0)
     
     [Om,cg,ad,anl] = param_maker(k0,om,sig);    
     
@@ -33,54 +32,80 @@ function nls_Dysthe_comparison(Llx,K,ep,tf,dt,om,sig,width,Nens)
     Kmesh = pi/Llx*Kvec;
     Kmeshf = pi/Llxf*Kvec;
         
-    Xmesh = -Llx:Llx/K:Llx-Llx/K;
+    %Xmesh = -Llx:Llx/K:Llx-Llx/K;
     
     uints = zeros(KT,Nens);
     uintsac = zeros(KT,Nens);
     nls_sol = zeros(KT,Nens);
     dysthe_sol = zeros(KT,Nens);
-    act_sol = zeros(KT,Nens);
+    %act_sol = zeros(KT,Nens);
     
     rvec = exp(-(Kmesh).^2/(2*width^2));
     rvecac = exp(-(Kmeshf-k0).^2/(2*width^2*ep^2));
     kvec = pi/Llx*(-K+1:K)';
-    %deadinds = abs(ifftshift(kvec)) > sqrt(2)*width;
-    
+    deadinds = abs(Kmesh) > 2*k0;
+    %deadindsac = logical(1 - (Kmeshf>(k0-2*k0*ep)).*(Kmeshf<(k0+2*k0*ep)));
     for jj=1:Nens
[…]

### 2019-05-30 vor_Dysthe_solver.m

@@ -11,10 +11,13 @@ function w = vor_Dysthe_solver(K,Llx,nmax,ad,anl,cg,k0,Om,om,sig,ep,dt,uint)
     Dx = 1i*Kmesh;
     Dx2 = Dx.^2;
     s = sign(k0);
+    wto = om - 2*s*Om;
     %L0i = (1 + 2*1i*ep*s*cg/(om-2*s*Om)*Dx).^(-1);
-    L0i = (1 - 2*1i*ep*s*cg/(om-2*s*Om)*Dx);
+    %L0i = (1 - 2*1i*ep*s*cg/(om-2*s*Om)*Dx);
     %Lap = dt*L0i.*(1i*ad + ep*sig/(om-2*s*Om)*Dx).*Dx2;
-    Lap = dt*(L0i*1i*ad+ ep*sig/(om-2*s*Om)*Dx).*Dx2;
+    %Lap = dt*(L0i*1i*ad+ ep*sig/(om-2*s*Om)*Dx).*Dx2;
+    Lap = dt*(1i*ad + ep*(2*s*cg*ad+sig)/wto*Dx + ep^2*1i*(ad^2*s/wto - (2*s*cg*ad+sig)*2*s*cg/wto^2)*Dx2).*Dx2;
+    
     KT = 2*K;
     Kc = floor(KT/3);
     Kuc = KT - Kc + 1;

### 2019-05-30 nls_Dysthe_comparison.m
Ready to Pull

@@ -63,24 +63,34 @@ function nls_Dysthe_comparison(Llx,K,ep,tf,dt,om,sig,width,Nens)
     %Since these work over such different time scales.  
     parfor jj=1:Nens
         nls_sol(:,jj) = nls_solver(K,Llx,nmax,ad,anl,dt,uints(:,jj));
-        % dysthe_sol(:,jj) = vor_Dysthe_solver(K,Llx,nmax,ad,anl,cg,k0,Om,om,sig,ep,dt,uints(:,jj));        
+        dysthe_sol(:,jj) = vor_Dysthe_solver(K,Llx,nmax,ad,anl,cg,k0,Om,om,sig,ep,dt,uints(:,jj));        
     end    
     parfor jj=1:Nens
        act_sol(:,jj) = afm_dno_solver(K,k0,ep,Llxf,sig,om,Om,tf/ep^2,dt,uintsac(:,jj)) 
     end
     toc
-    nls_spec_mean = (ep/KT)^2*fftshift(abs(mean(nls_sol.*conj(nls_sol),2)));
-    % dysthe_spec_mean = ep^2*fftshift(abs(mean(dysthe_sol.*conj(dysthe_sol),2))/KT^2);
-    mean_int = (ep/KT)^2*fftshift(abs(mean(uints.*conj(uints),2)));
-    mean_act = (ep/KT)^2*fftshift(abs(mean(act_sol.*conj(act_sol),2)));
     
-    %plot(ep*kvec+k0,mean_int,'k-',ep*kvec+k0,nls_spec_mean,'k--',ep*kvec+k0,dysthe_spec_mean,'k-.',ep*kvec+k0,mean_act,'k:','LineWidth',2)
-    plot(ep*kvec+k0,mean_int,'k-',ep*kvec+k0,nls_spec_mean,'k--',ep*kvec+k0,mean_act,'k:','LineWidth',2)
+    ekT = (ep/KT)^2;
+    nls_spec_mean = ekT*fftshift(abs(mean(nls_sol.*conj(nls_sol),2)));
+    dysthe_spec_mean = ekT*fftshift(abs(mean(dysthe_sol.*conj(dysthe_sol),2)));
+    mean_int = ekT*fftshift(abs(mean(uints.*conj(uints),2)));
+    mean_act = ekT*fftshift(abs(mean(act_sol.*conj(act_sol),2)));
+    
+    dk = ep*pi/Llx;
+    
[…]

### 2019-07-17 Writeup/random_vorticity.tex
Proper Update and Draft

@@
 In particular, most understandings of MI rely on looking at the stability of perturbations to carrier waves at essentially a fixed wave number.
-To address this shortcoming, in a now seminal paper, \cite{alber} analytically studies families of perturbations of wave packets with narrow, but non-zero, spectral width around a central wave number.
-An analytic critierion determining when modulational instability (MI) is either manifested or supressed depending on this spectral width is derived.
+To address this shortcoming, in a now seminal paper, \cite{alber} analytically studied families of perturbations of wave packets with narrow, but non-zero, spectral width around a central wave number.
+An analytic critierion determining when modulational instability (MI) is either manifested or supressed depending on this spectral width, now refered to as a Benjamin--Feir index (BFI) was derived.
 The results in \cite{alber} were confirmed numerically in \cite{dysthe2,dysthe3} by examing the mean properties associated with ensembles of initial conditions.
@@
 In particular, vorticity was not taken into account, thereby ignoring a central mechanism for the transfer of energy across length scales in oceanic flows.
-Thus, building on the work in \cite{thomas2012nonlinear} and \cite{curtis8}, we numerically explore the analytic stability condition derived in \cite{alber}, focusing in particular on the impact that vorticity plays in controlling the effective width for the onset of MI.
-Moving beyond just examining the properties of the NLS equation with vorticity, we look at the statistical properties of solutions to a higher-order model, the vor-Dysthe equation derived in \cite{curtis8}.
-As shown in \cite{onorato}, the higher-order terms associated with the Dysthe equation can have significant impacts on the statistics of the waves, so it is a nontrivial question to examine how the vorticity will interact with these higher-order nonlinearities.
+Thus, building on the work in \cite{thomas2012nonlinear} and \cite{curtis8}, we numerically explore the BFI condition derived in \cite{alber}, focusing in particular on the impact that vorticity plays in controlling the effective width for the onset of MI.
+Moving beyond just examining the properties of the nonlinear Schr\"{o}dinger equation (NLSE) with vorticity, we look at the statistical properties of solutions to a higher-order model, the vor-Dysthe equation (VDE) derived in \cite{curtis8}.
+As shown in \cite{onorato}, the higher-order terms associated with the Dysthe equation can have significant impacts on the statistics of the waves, so it is a nontrivial question to examine how the vorticity interacts with these higher-order nonlinearities.
+
+Using our simulations, and as predicted in effect by the results in \cite{curtis8}, we show that we can exacerbate or suppress MI by changing the direction in which the constant vorticity shear profile moves relative to the propagating surface waves.
+Using ensembles of initial conditions for a given choice of spectral width, we see that both the NLSE and VDE exhibit either the MI or supression thereof predicted by the BFI as derived in \cite{alber}.
+However, as we show, starting from Gaussian initial conditions, the NLSE generates essentially Gaussian results.
+Moreover, it is incapable of shifting means, and thus is clearly of limited utility in describing realistic wave fields.
+In contrast, our results show that the VDE is able to capture non-Gaussian effects and allows for shifts in the mean relative to the distribution of initial conditions.
+
+A particularly interesting consequence of looking at wave packets is that the BFI condition gives one an estimate of what class of wavetrains would be stable given a particular shear profile.
+Thus, if one supposes that only stable profiles are physically observable, then the results in this paper give an indication of what wave trains one is most likely to see in oceanographic contexts.
 
@@
 
-Taylor expanding Equation (\ref{kine2}) up to $\mathcal{O}(\epsilon^3)$ for $k\neq 0$ gives
 \begin{multline}
-\int_{\mathbb{R}}dx~e^{-ikx}\left(1 + \epsilon |k|\eta + \frac{\epsilon^{2}|k|^{2}\eta^{2}}{2} \right)\left(\eta_{t} + i \mbox{sgn}(k)Q \right) \\
-+ \epsilon \omega \int_{\mathbb{R}}dx~e^{-ikx}\left(1 + \epsilon |k|\eta \right) \eta \eta_{x} = 0.
+\int_{\mathbb{R}}dx~e^{-ikx}\left(1 + \epsilon |k|\eta + \frac{\epsilon^{2}|k|^{2}\eta^{2}}{2} + \frac{\epsilon^{3}|k|^{3}\eta^{3}}{6}\right)\left(\eta_{t} + i \mbox{sgn}(k)Q \right) \\
++ \epsilon \omega \int_{\mathbb{R}}dx~e^{-ikx}\left(1 + \epsilon |k|\eta + \frac{\epsilon^{2}|k|^{2}\eta^{2}}{2}\right) \eta \eta_{x} = 0.
 \label{integro1}
@@
 Q_{t} + \omega \eta_{t} + \eta_{x} - \tilde{\sigma}\eta_{xxx} + \frac{\epsilon}{2}\pd_{x}\left(-\eta_{t}^{2} + (Q+\omega \eta)^{2} \right)\\
-+ \epsilon^{2}\pd_{x}\left( \frac{3}{2}\tilde{\sigma}\eta_{x}^{2}\eta_{xx} - \eta_{t}\eta_{x} \left(Q  + \omega \eta \right) \right)  = 0,
++ \epsilon^{2}\pd_{x}\left( \frac{3}{2}\tilde{\sigma}\eta_{x}^{2}\eta_{xx} - \eta_{t}\eta_{x} \left(Q  + \omega \eta \right) \right) - \frac{\epsilon^{3}}{2}\eta_{x}^{2}\left((\omega\eta+Q)^{2}-\eta_{t}^{2}\right) = 0,
 \label{berexp}
@@
 
-By expanding Equation \eqref{integro1} and taking inverse Fourier transforms, we can derive the first three terms of the corresponding Dirichlet-to-Neumann Operator (DNO) expansion, so that we find
-\begin{equation}
-\eta_{t} = G_{0}Q + \epsilon G_{1}Q + \epsilon^{2}G_{2}Q,
-\label{dnoexp}
-\end{equation}
-where
-\begin{align*}
-G_{0}Q = & -\mathcal{H}Q\\
-G_{1}Q = & \mathcal{H}\pd_{x}\left(\eta G_{0}Q\right) - \pd_{x}\left(\eta Q + \frac{\omega}{2}\eta^{2}\right)\\
-G_{2}Q = & \mathcal{H}\pd_{x}\left(\eta G_{1}Q\right) + \frac{1}{2}\pd_{x}^{2}\left(\eta^{2}G_{0}Q \right) + \pd_{x}^{2}\mathcal{H}\left(\frac{1}{2}\eta^{2}Q + \frac{\omega}{3}\eta^{3}\right), 
-\end{align*}
-where $\mathcal{H}$ denotes the Hilbert transform so that 
-\[
-\mathcal{H}f = \frac{1}{2\pi}\int_{\mathbb{R}}dk~ e^{ikx}  i\mbox{sgn}(k) \hat{f}(k),
-\]
-where $\hat{f}(k)$ is the Fourier transform of $f(x)$ defined by
-\[
-\hat{f}(k) = \int_{\mathbb{R}}dx~ e^{-ikx}f(x).
-\]
-Using the DNO expansion in Equation \eqref{dnoexp}, we can rewrite Equation \eqref{berexp} an an evolution equation for $Q_{t}$, thereby giving us a closed system in terms of surface variables alone.
+%By expanding Equation \eqref{integro1} and taking inverse Fourier transforms, we can derive the first three terms of the corresponding Dirichlet-to-Neumann Operator (DNO) expansion, so that we find
+%\begin{equation}
+%\eta_{t} = G_{0}Q + \epsilon G_{1}Q + \epsilon^{2}G_{2}Q,
+%\label{dnoexp}
+%\end{equation}
+%where
+%\begin{align*}
+%G_{0}Q = & -\mathcal{H}Q\\
+%G_{1}Q = & \mathcal{H}\pd_{x}\left(\eta G_{0}Q\right) - \pd_{x}\left(\eta Q + \frac{\omega}{2}\eta^{2}\right)\\
+%G_{2}Q = & \mathcal{H}\pd_{x}\left(\eta G_{1}Q\right) + \frac{1}{2}\pd_{x}^{2}\left(\eta^{2}G_{0}Q \right) + \pd_{x}^{2}\mathcal{H}\left(\frac{1}{2}\eta^{2}Q + \frac{\omega}{3}\eta^{3}\right), 
+%\end{align*}
+
+%Using the DNO expansion in Equation \eqref{dnoexp}, we can rewrite Equation \eqref{berexp} an an evolution equation for $Q_{t}$, thereby giving us a closed system in terms of surface variables alone.
 The system of equations given by Equations \eqref{dnoexp} and \eqref{berexp} we denote as the Full-Euler Equations (FEEs).
@@
 
-To generate simulations, we start both the NLSE and FEE from the same collection of $2L$ periodic-initial conditions
+However, the NLSE is a relatively low-order model known to be only of limited use in real-world modeling situations.
+A more robust model, following the work in \cite{dysthe1}, is found by going to the next higher order in $\epsilon$, giving us the
+\begin{multline}
+(\omega-2s\Omega)\pd_{\tau}\eta_{1} + i(c_{g}^{2}s-3k_{0}\tilde{\sigma})\pd_{\xi}^{2}\eta_{1} + 2i\epsilon s c_{g}\pd^{2}_{\xi\tau}\eta_{1} - \epsilon\tilde{\sigma} \pd^{3}_{\xi}\eta_{1} + i\epsilon^{2}c_{h}\pd_{\xi}^{4}\eta_{1} \\
+- i \alpha_{nl}(\omega-2s\Omega)\left|\eta_{1}\right|^{2}\eta_{1}
++ \epsilon \tilde{\alpha}_{1}\left|\eta_{1}\right|^{2}\pd_{\xi}\eta_{1}
++ \epsilon\tilde{\alpha}_{2}\eta_{1}^{2}\pd_{\xi}\eta^{\ast}_{1} +
+\epsilon i\tilde{\alpha}_{3}\eta_{1}\mathcal{H}\pd_{\xi}\left|\eta_{1}\right|^{2}= 0.
+\label{dysthe}
+\end{multline}
+where $\mathcal{H}$ denotes the Hilbert transform so that 
+\[
+\mathcal{H}f = \frac{1}{2\pi}\int_{\mathbb{R}}dk~ e^{ikx}  i\mbox{sgn}(k) \hat{f}(k),
+\]
+where $\hat{f}(k)$ is the Fourier transform of $f(x)$ defined by
+\[
+\hat{f}(k) = \int_{\mathbb{R}}dx~ e^{-ikx}f(x).
+\]
+We call Equation \eqref{dysthe} the Vor-Dysthe (VD) equation.
+Details of its derivation can be found in \cite{curtis8}, though we note a difference in this work with that in \cite{curtis8}.
+In this work, we found the inclusion of the higher-order dispersion associated with the coefficient $c_{h}$, given by
+\[
+c_{h} = 2sc_{g}(2sc_{g}\alpha_{d}+\tilde{\sigma})/(\omega-2s\Omega) - s\alpha_{d}^{2},
+\]
+was necessary to remove artificial numerical instabilities.
+Similar issues with Dysthe like equations have been noted in (add appropriate citation).
+Moreover, the coefficients $\tilde{\alpha}_{j}$ are readily derived, but their forms are cumbersome and thus we omit explicitly writing them for the sake of readability.
+
+To generate simulations, we start both the NLSE and VDE from the same collection of $2L$ periodic-initial conditions
 \[
@@
 \[
-\sigma \geq \epsilon_{rms}\sqrt{\frac{2\alpha_{nl}}{\alpha_{d}}}
+\sigma \geq \sigma_{c}, ~ \sigma_{c} = \epsilon_{rms}\sqrt{\frac{2\alpha_{nl}}{\alpha_{d}}}
 \]
@@
 De-aliasing is implemented via Orszag's `2/3-rule'.
-Averaging was done with $N_{e}=3200$ ensemble members, thereby ensuring relative averaging errors on the order of about $1\%$.
+Averaging was done with $N_{e}=512$ ensemble members, thereby ensuring relative averaging errors on the order of about $1\%$.
 Throughout this section, we present comparisons of the distributions for particular choices of $\sigma$ while also comparing the mean and standard deviations of the respective distributions so as to get a more quantitative understanding of the differences between the different distributions.
@@
 
-Motivated by the results in \cite{curtis8}, we anticipate that $\omega=-1$ presents the most case which would allow for the vorticity to have the strongest impact on the BFI condition, which for the parameter choices requires that $\sigma > 11.9$.
-This is borne out by numerical experiment as seen in Figures \ref{fig:meanstdomn1} and \ref{fig:pspecomn1}.
+Motivated by the results in \cite{curtis8}, we anticipate that $\omega=-1$ presents the most case which would allow for the vorticity to have the strongest impact on the BFI condition, which for the parameter choices gives us the BFI $\sigma_{c} \approx 11.9$.
+This is borne out by numerical experiment as seen in Figure \ref{fig:meanstdomn1}.
 In this case, we see that for the NLSE the BFI condition largely holds.
@@
 \begin{tabular}{c}
-\includegraphics[width=.85\textwidth]{omega_n1_ep0pt1_dt5en4}\\
-(a) $\epsilon=.1$ \\
-\includegraphics[width=.85\textwidth]{omega_n1_ep0pt04}\\
-(b) $\epsilon=.04$
+\includegraphics[width=.65\textwidth]{pdf_w_n1_ep_pt1_Nens_512} \\
+(a) $\epsilon=.1$\\
+\includegraphics[width=.65\textwidth]{omega_n1_mean_std_plot_ep_pt1}\\
+(b) $\epsilon=.1$ \\
+\includegraphics[width=.65\textwidth]{pdf_w_n1_ep_pt05_Nens_512}\\
+(c) $\epsilon=.05$ \\
+\includegraphics[width=.65\textwidth]{omega_n1_mean_std_plot_ep_pt05}\\
+(d) $\epsilon=.05$
 \end{tabular}
-\caption{For $\omega=-1$, comparison of the means and standard deviations between the initial averaged spectral density and those generated by the NLSE and VDE.
-This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).
+\caption{For $\omega=-1$, in (a) and (c) plots of the distributions for $\epsilon=.1$ and $\epsilon=.05$ respectively,  with $\sigma=1.05\sigma_{c}$.
+In (b) and (d) comparison of the means and standard deviations between the initial averaged spectral density for $\epsilon=.1$ and $\epsilon=.05$ respectively.
 The vertical line denotes the spectral width corresponding to the BFI condition.}
@@
 
-\begin{figure}[!ht]
-\centering
-\begin{tabular}{c}
-\includegraphics[width=.65\textwidth]{omega_n1_ep0pt1_dt5en4_width12} \\
-(a) $\epsilon=.1$\\
-\includegraphics[width=.65\textwidth]{omega_n1_ep0pt04_width12}\\
-(b) $\epsilon=.04$ 
-\end{tabular}
-\caption{For $\omega=-1$, comparison of the averaged spectral densities between the initial averaged spectral density and those generated by the NLSE and VDE.
-This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).}
-\label{fig:pspecomn1}
-\end{figure}
-
 However, the results for the VDE are markedly different.
-As seen in Figure \ref{fig:pspecomn1}, where $\sigma=12$ so that we are just past the BFI threshold, the most peculiar feature of the VDE is the bimodality of the distribution.
-There is clearly a frequency downshift of the central initial peak, but a secondary peak at high-wave number forms as well.
-This formation is less pronounced with decreasing $\epsilon$, but is clearly there nevertheless.
-This represents a kind of resonance condition for which no ready explanation exists.
-This feature is likewise responsible for the strong skewing of the means and standard deviations seen in Figure \ref{fig:meanstdomn1}.
+As seen in Figure \ref{fig:meanstdomn1}, where $\sigma=12$ so that we are just past the BFI threshold,there is clearly a frequency downshift of the central initial peak.
+This downshift is less pronounced with decreasing $\epsilon$, but is clearly there nevertheless.
+Likewise, noting that the overall mean is rightshifted in Figures \ref{fig:meansdomn1} (b) and (d), we see that the tails of the distributions are skewed in the VDE, in effect adding kurtosis and pushing the final distribution away from the Gaussian initial conditions.
 
@@
 Again though, the VDE produces markedly different statistical results from the NLS, though to nowhere near as drastic an extent as when $\omega=-1$.
-In particular, the bimodality and strong deformation of the profile is notably ameliorated for $\epsilon=.04$; see Figures \ref{fig:meanstdom0} and \ref{fig:pspecom0}.
+In particular, the bimodality and strong deformation of the profile is notably ameliorated for $\epsilon=.05$; see Figure \ref{fig:meanstdom0}.
 Thus in some sense, though not entirely, MI has more or less similar characteristics for both the NLSE and VDE.
@@
 \begin{tabular}{c}
-\includegraphics[width=.85\textwidth]{omega0_ep0pt1}\\
-(a) $\epsilon=.1$ \\
-\includegraphics[width=.85\textwidth]{omega0_ep0pt04}\\
-(b) $\epsilon=.04$
+\includegraphics[width=.65\textwidth]{pdf_w_0_ep_pt1_Nens_512} \\
+(a) $\epsilon=.1$\\
+\includegraphics[width=.65\textwidth]{omega_0_mean_std_plot_ep_pt1}\\
+(b) $\epsilon=.1$ \\
+\includegraphics[width=.65\textwidth]{pdf_w_0_ep_pt05_Nens_512}\\
+(c) $\epsilon=.05$ \\
+\includegraphics[width=.65\textwidth]{omega_0_mean_std_plot_ep_pt05}\\
+(d) $\epsilon=.05$
 \end{tabular}
-\caption{For $\omega=0$, comparison of the means and standard deviations between the initial averaged spectral density and those generated by the NLSE and VDE.
-This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).
+\caption{For $\omega=0$, in (a) and (c) plots of the distributions for $\epsilon=.1$ and $\epsilon=.05$ respectively,  with $\sigma=1.05\sigma_{c}$.
+In (b) and (d) comparison of the means and standard deviations between the initial averaged spectral density for $\epsilon=.1$ and $\epsilon=.05$ respectively.
 The vertical line denotes the spectral width corresponding to the BFI condition.}
@@
 
-\begin{figure}[!ht]
-\centering
-\begin{tabular}{c}
-\includegraphics[width=.65\textwidth]{omega0_ep0pt1_width6} \\
-(a) $\epsilon=.1$\\
-\includegraphics[width=.65\textwidth]{omega0_ep0pt04_width6}\\
-(b) $\epsilon=.04$ 
-\end{tabular}
-\caption{For $\omega=0$, comparison of the averaged spectral densities between the initial averaged spectral density and those generated by the NLSE and VDE.
-This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).}
-\label{fig:pspecom0}
-\end{figure}
-
-\subsection*{Strong-Co Shear: $\omega = 1$}
+\subsection*{Strong-Co Shear: $\omega = 1.12$}
 
 Finally, by letting $\omega=1$, we further reduce the BFI threshold to $\sigma > 2.7$, opening up a yet wider range of stable initial conditions.
-Likewise, the VDE dynamics are far more in line with those seen from the NLSE, even in the case that $\epsilon=.1$; see Figures \ref{fig:meanstdom1} and \ref{fig:pspecom1}.
+Likewise, the VDE dynamics are far more in line with those seen from the NLSE, even in the case that $\epsilon=.1$; see Figure \ref{fig:meanstdom1}.
 Therefore, we see how vorticity propagating in the correct direction can vastly reduce the impact of MI, even in a far more nonlinear model like the VDE.
@@
 \begin{tabular}{c}
-\includegraphics[width=.85\textwidth]{omega1_ep0pt1}\\
-(a) $\epsilon=.1$ \\
-\includegraphics[width=.85\textwidth]{omega1_ep0pt04}\\
-(b) $\epsilon=.04$
+\includegraphics[width=.65\textwidth]{pdf_w_1pt12_ep_pt1_Nens_512} \\
+(a) $\epsilon=.1$\\
+\includegraphics[width=.65\textwidth]{omega_1pt12_mean_std_plot_ep_pt1}\\
+(b) $\epsilon=.1$ \\
+\includegraphics[width=.65\textwidth]{pdf_w_1pt12_ep_pt05_Nens_512}\\
+(c) $\epsilon=.05$ \\
+\includegraphics[width=.65\textwidth]{omega_1pt12_mean_std_plot_ep_pt05}\\
+(d) $\epsilon=.05$
 \end{tabular}
-\caption{For $\omega=1$, comparison of the means and standard deviations between the initial averaged spectral density and those generated by the NLSE and VDE.
-This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).
+\caption{For $\omega=1.12$, in (a) and (c) plots of the distributions for $\epsilon=.1$ and $\epsilon=.05$ respectively,  with $\sigma=1.05\sigma_{c}$.
+In (b) and (d) comparison of the means and standard deviations between the initial averaged spectral density for $\epsilon=.1$ and $\epsilon=.05$ respectively.
 The vertical line denotes the spectral width corresponding to the BFI condition.}
@@
 
-\begin{figure}[!ht]
-\centering
-\begin{tabular}{c}
-\includegraphics[width=.65\textwidth]{omega1_ep0pt1_width3} \\
-(a) $\epsilon=.1$\\
-\includegraphics[width=.65\textwidth]{omega1_ep0pt04_width3}\\
-(b) $\epsilon=.04$ 
-\end{tabular}
-\caption{For $\omega=1$, comparison of the averaged spectral densities between the initial averaged spectral density and those generated by the NLSE and VDE.
-This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).}
-\label{fig:pspecom1}
-\end{figure}
+\section*{Conclusion}
 
-\section*{Conclusion}
+In this work, we have studied the modulational stability of nearly modulated deep-water waves under the influence of steady linear shear profiles in the more physically realistic case in which the modulation is done on wave packets.
+As we see, the vorticity profile has a strong influence on the stability properties of the wave train, with strong counter propagating shears exacerbating modulational instability effects.
+In contrast, strong copropagating shears all but supress modulational instability.
+Thus, we see vorticity effects can have significant implications for what the most likely observed deep water wavetrains are.
+Likewise, in trying to interpret oceanographic data, vortical effects should be taken into account in modeling so as to be able to properly interpret observed data.
+
 \bibliography{deep_water_shear_Feb2}

### 2019-07-17 nls_Dysthe_comparison.m

@@ -43,7 +43,7 @@ function [nls_avg,std_nls,dys_avg,std_dys,int_avg,std_int] = nls_Dysthe_comparis
     rvec = exp(-(Kmesh).^2/(2*width^2));
     rvecac = exp(-(Kmeshf-k0).^2/(2*width^2*ep^2));
     kvec = pi/Llx*(-K+1:K)';
-    deadinds = abs(Kmesh) > 2*k0;
+    deadinds = abs(Kmesh) > 16*k0;
     %deadindsac = logical(1 - (Kmeshf>(k0-2*k0*ep)).*(Kmeshf<(k0+2*k0*ep)));
     for jj=1:Nens
         arand = exp(2*pi*1i*rand(KT,1));
@@ -98,14 +98,16 @@ function [nls_avg,std_nls,dys_avg,std_dys,int_avg,std_int] = nls_Dysthe_comparis
     %std_act = dk/2*(stact(1)+stact(end)+2*sum(stact(2:end-1)));
     disp([nls_avg; dys_avg; int_avg])
     disp([std_nls; std_dys; std_int])
-    
-    %plot(rkvec,mean_int_plot,'k-',rkvec,nls_spec_plot,'k--',rkvec,dysthe_spec_plot,'k-.','LineWidth',2)
+    %{
+    figure(2)
+    plot(rkvec,mean_int_plot,'k-',rkvec,nls_spec_plot,'k--',rkvec,dysthe_spec_plot,'k-.','LineWidth',2)
     %plot(rkvec,mean_int_plot,'k-',rkvec,nls_spec_plot,'k--',rkvec,dysthe_spec_plot,'k-.',rkvec,mean_act_plot,'k:','LineWidth',2)
-    %h = set(gca,'FontSize',30);
-    %set(h,'Interpreter','LaTeX')
-    %xlabel('$k$','Interpreter','LaTeX','FontSize',30)
-    %ylabel('$\left<\left|\hat{\eta}(k,t_{f})\right|^{2}\right>$','Interpreter','LaTeX','FontSize',30)
-    %legend({'$Initial$','$NLS$','$Dysthe$'},'Interpreter','LaTeX','FontSize',30)
+    h = set(gca,'FontSize',30);
+    set(h,'Interpreter','LaTeX')
+    xlabel('$k$','Interpreter','LaTeX','FontSize',30)
[…]

### 2019-07-17 plot_maker.m

@@ -5,7 +5,7 @@ K = 256;
 ep = .05;
 tf = 1;
 dt = 5e-3;
-om = 1.12;
+om = 0;
 sig = 1e-5;
 Nens = 512;
 k0 = 1;
@@ -13,7 +13,7 @@ k0 = 1;
 mwid = sqrt(2*anl/ad);
 disp('Minimal Stable Envelope Width')
 disp(mwid)
-    
+
 nowdths = 4;
 widths = linspace(.5*mwid,1.5*mwid,nowdths);
 nls_avgs = zeros(nowdths,1);
@@ -22,7 +22,6 @@ act_avgs = zeros(nowdths,1);
 stds_act = zeros(nowdths,1);
 int_avgs = zeros(nowdths,1);
 stds_int = zeros(nowdths,1);
-
 for jj=1:nowdths
     [nls_avg,std_nls,act_avg,std_act,int_avg,std_int] = nls_Dysthe_comparison(Llx,K,ep,tf,dt,om,sig,widths(jj),k0,Nens);
     nls_avgs(jj) = nls_avg;
@@ -32,6 +31,10 @@ for jj=1:nowdths
     int_avgs(jj) = int_avg;
     stds_int(jj) = std_int;
 end
+
+%widths = 1.05*mwid;
+%disp(widths)
+%[nls_avg,std_nls,act_avg,std_act,int_avg,std_int] = nls_Dysthe_comparison(Llx,K,ep,tf,dt,om,sig,widths,k0,Nens);
 
 figure(1)
 %tstr = sprintf('$\omega=$%f, $\epsilon=$%f', om, ep);

### 2019-09-03 Writeup/random_vorticity.tex
Moving on

@@
 \section*{Results}
-To understand the relationship between the vorticity $\omega$ and the initial spectral width $\sigma$ of our initial conditions, we generate ensembles of initial conditions and then run these different initial conditions up to $\tau=1$, which corresponds to $t_{f} = 1/\epsilon^{2}$ in the original fast variables.
+To understand the relationship between the vorticity $\omega$ and the initial spectral width $\sigma$ of our initial conditions, we generate ensembles of initial conditions and then run these different initial conditions up to $\tau=2$ which corresponds to $t_{f} = 2/\epsilon^{2}$ in the original fast variables.
 The domain is $[-50\pi,50\pi]$.
-The number of modes is $K = 256$.
+The number of modes is $K = 512$.
 The surface tension is $\tilde{\sigma} = 10^{-5}$.

### 2019-12-16 Writeup/higher_order_dispersion_relationship_nost_writeup.tex
Updates

 \documentclass[a4paper,11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{epstopdf}
\epstopdfsetup{update}

\newcommand{\ba}{\begin{array}}
\newcommand{\ea}{\end{array}}

\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\newcommand{\bc}{\begin{center}}
\newcommand{\ec}{\end{center}}

\newcommand{\ds}{\displaystyle}

\newcommand{\bt}{\begin{tabular}}
\newcommand{\et}{\end{tabular}}

\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}

\newcommand{\bd}{\begin{description}}
\newcommand{\ed}{\end{description}}

\newcommand{\bp}{\begin{pmatrix}}
\newcommand{\ep}{\end{pmatrix}}

\newcommand{\pd}{\partial}
\newcommand{\sech}{\mbox{sech}}

\newcommand{\cf}{{\it cf.}~}

\newcommand{\ltwo}{L_{2}(\mathbb{R}^{2})}
\newcommand{\smooth}{C^{\infty}_{0}(\mathbb{R}^{2})}

\newcommand{\br}{{\bf r}}
\newcommand{\bk}{{\bf k}}
\newcommand{\bv}{{\bf v}}

\newcommand{\gnorm}[1]{\left|\left| #1\right|\right|}
\newcommand{\ipro}[2]{\left<#1,#2 \right>}

\begin{document}
\section*{Appendix}
\subsection*{Shear Profiles over Infinitely Deep Fluids}
As noted in the text, the assumption that the fluid velocity is to leading order given by 
\[
{\bf u} \approx \omega z \hat{{\bf i}}
\]
is clearly unrealistic insofar as it leads to currents of infinite speed as one descends through the fluid.  A more realistic, though also more complicated, ansatz is to suppose that to leading order we have that ${\bf u}\approx u(z)\hat{{\bf i}}$ where  
\[
u(z) = \left\{\ba{rl} \omega z, & -h_{0} \leq z < 0 \\
\frac{-\omega h_{0}}{h_{1}-h_{0}} (z+h_{1}), & -h_{1} \leq z < -h_{0} \\
0, & z < -h_{1}
\ea \right.
\] 
where $h_{1}>h_{0}\gg 1$, so that we are looking at a deep, continuous shear profile which is zero at or near the surface $z=\epsilon \eta(x,t)$ and at the depth $z=-h_{1}$, after which the fluid is to leading order quiescent.  Note, we can also see this profile as satisfying to leading order two `no-slip' conditions, one near the free surface at $z=0$ and one near $z=-h_{1}$.

While a full, nonlinear description of the above shear profile would require two more freely evolving interfaces, and thus is beyond the scope of this paper, we can readily find the dispersion relationship affiliated with this profile.  This then allows us to provide some analytic argument for why we study an otherwise unphysical velocity profile in the main body of the text.  Likewise, information from the dispersion relationship provides us with a better understanding of how depth varying shear profiles induce both surface and internal waves.

Following relatively classical approaches, we introduce three fluid velocities
\begin{align*}
{\bf u}_{1} = \omega z \hat{{\bf i}} + \epsilon \nabla \phi_{1}, &  ~-h_{0}+\epsilon \eta_{2} < z < \epsilon \eta_{1}\\
{\bf u}_{2} = \frac{-\omega h_{0}}{h_{1}-h_{0}} (z+h_{1}) \hat{{\bf i}} + \epsilon \nabla \phi_{2}, & ~-h_{1}+\epsilon \eta_{3} < z < -h_{0}+\epsilon \eta_{2}\\
{\bf u}_{3} = \epsilon \nabla \phi_{3}, & ~z < -h_{1} + \epsilon \eta_{3}
\end{align*}    
so that after linearizing around the small disturbances, we have the following system of evolution equations describing the behavior of small disturbances
\[
\pd_{t}\phi_{1} + g\eta_{1} + \omega \pd_{t}\pd_{x}^{-1}\eta_{1} = 0, ~ z = 0 
\]
\[
\pd_{t}\eta_{1} = \pd_{z}\phi_{1}, ~ z = 0, 
\]
\[
\pd_{t}\phi_{2} - \frac{\omega h_{1}}{h_{1}-h_{0}} \pd_{t}\pd_{x}^{-1}\eta_{2} = \pd_{t}\phi_{1} , ~ z = -h_{0} 
\]
\begin{align*}
\pd_{t}\eta_{2} = \pd_{z}\phi_{2}, & ~ z = -h_{0},\\
\pd_{t}\eta_{2} = \pd_{z}\phi_{1}, & ~ z = -h_{0}, 
\end{align*}
and
\[
\pd_{t}\phi_{3} = \pd_{t}\phi_{2} - \frac{\omega h_{0}}{h_{1}-h_{0}} \pd_{t}\pd_{x}^{-1}\eta_{3}, ~ z = -h_{1} 
\]
\begin{align*}
\pd_{t}\eta_{3} = \pd_{z}\phi_{3}, & ~ z = -h_{1},\\
\pd_{t}\eta_{3} = \pd_{z}\phi_{2}, & ~ z = -h_{1}. 
\end{align*}
Taking each function $\phi_{j}$ to be harmonic and letting $\pd_{z}\phi_{3}\rightarrow 0 $ as $z\rightarrow -\infty$ gives us the following solutions for each of the scalar velocity potentials
\[
\phi_{1}(x,z,t) = \left(\alpha_{11}\cosh(kz)+\alpha_{12}\sinh(kz)\right)e^{i\theta(x,t)} + \mbox{cc}
\]
\[
\phi_{2}(x,z,t) = \left(\alpha_{21}\cosh(k(z+h_{0}))+\alpha_{22}\sinh(k(z+h_{0}))\right)e^{i\theta(x,t)} + \mbox{cc}
\]
and
\[
\phi_{3}(x,z,t) = \alpha_{3}e^{|k|(z+h_{1})}e^{i\theta(x,t)} + \mbox{cc}
\]
where $\theta(x,t) = kx + \Omega(k,\omega)t$, and `cc' denotes the complex conjugate.

This then gives us the linear system of equations 
\begin{align*}
\Omega \alpha_{11} = & \left(\frac{gk}{\Omega} + \omega \right)\alpha_{12} \\
\Omega\left(\alpha_{11}\cosh(kh_{0}) -\alpha_{12}\sinh(kh_{0}) \right) = &  \Omega \alpha_{21} + \frac{\omega h_{1}}{\delta h}\alpha_{22}\\
\Omega\left(\alpha_{21}\cosh(k\delta h) - \alpha_{22}\sinh(k \delta h) \right) = & \left(\Omega - \frac{\omega h_{0} s}{\delta h} \right)\alpha_{3}\\
\alpha_{22} = & -\alpha_{11}\sinh(kh_{0}) + \alpha_{12}\cosh(kh_{0})  \\
\alpha_{3} = & s_{k}\left(-\alpha_{21}\sinh(k\delta h) + \alpha_{22}\cosh(k\delta h)\right)
\end{align*}
where $\delta h = h_{1}-h_{0}$ and $s=\mbox{sgn}(k)$.  From this we derive the dispersion relationship 
\[
\frac{gk}{\Omega} + \omega -\Omega \tanh(kh_{0}) + \left(\frac{1}{\Omega}\left(\frac{gk}{\Omega} + \omega\right)\tanh(kh_{0}) - 1\right)\left(\frac{\omega h_{1}}{\delta h} + \Omega \Omega_{1}(k\delta h)\right) = 0
\]
where
\[
\Omega_{1}(k\delta h) = \frac{s\left(\Omega -\frac{\omega h_{0}}{\delta h}s\right)+\Omega \tanh(k\delta h)}{\Omega +s\left(\Omega - \frac{\omega h_{0}}{\delta h}s\right)\tanh(k\delta h)}.
\]
While in general we would have to find the roots of a fifth-order polynomial to determine the values of $\Omega$, we see for $h_{1}\gg h_{0}$ that $\Omega_{1}\sim s$, and thus we get that the dispersion relationship simplifies to 
\[
\frac{gk}{\Omega} + \omega -\Omega \tanh(kh_{0}) + \left(\frac{1}{\Omega}\left(\frac{gk}{\Omega} + \omega\right)\tanh(kh_{0}) - 1\right)\left(\tilde{\omega} + s\Omega \right) \sim 0,
\]
where 
\[
\tilde{\omega} = \frac{\omega h_{1}}{\delta h}.
\]
Since $\tilde{\omega}$ only approaches $\omega$ at an algebraic rate, it seems appropriate to keep it included in the analysis.  Letting 
\[
\left|\tanh(kh_{0}) \right| = 1 - \tilde{\epsilon}, 
\]
and noting that $\tilde{\epsilon}$ vanishes to zero exponentially fast as we increase $h_{0}$, we see the reduced dispersion relationship factors into the form
\[
\left(2\Omega + s\tilde{\omega} \right)\left(g\left|k\right| + s\omega \Omega - \Omega^{2} \right) + \tilde{\epsilon}\left(\Omega^{3} - (s\tilde{\omega} + \Omega)(g|k|+\omega s\Omega) \right) \sim 0.
\]
Thus we have a completely regular perturbation problem for the roots, which we readily see are given by
\[
\Omega \sim \frac{1}{2}\left(-s\omega \pm \sqrt{\omega^{2}+4g|k|} \right), ~ -\frac{s\tilde{\omega}}{2}.
\]
Thus of the three roots we find, two give the dispersion relationship we find in the body of the text using ${\bf u}\sim \omega z \hat{{\bf i}}$.  Looking at the affiliated disturbances of the relevant free surface and internal wave, we find that  
\[
\eta_{1} \sim \alpha_{12}\frac{ik}{\Omega}e^{i\theta} + \mbox{cc}, ~ \eta_{2} \sim \alpha_{12}\frac{ik}{\Omega^{3}}\cosh(kh_{0})\left(\Omega^{2}-(1-\tilde{\epsilon})\left( g|k|+s\omega\Omega\right)\right)e^{i\theta}+ \mbox{cc}.
\]
Thus if we choose $\Omega$ so that $\Omega^{2}-\left( g|k|+s\omega\Omega\right)= 0$, the magnitude of the internal wave essentially vanishes, thereby localizing dynamics along the free surface near $z=0$.  Therefore, while not necessarily physically justifiable throughout the bulk of the fluid, our simplified shear profile assumption produces results which are asymptotically consistent with a more sophisticated treatment of the shear profile.

We note though that the more physically realistic shear profile shows that there is a choice of $\Omega$ which corresponds to a non-trivial internal mode near $z=-h_{0}$.  We also note that this choice of $\Omega$ makes many of the terms in our modulation theory singular, thus showing that this case would require a markedly different treatment.  While interesting, this issue is beyond the scope of the current paper and will be addressed in future research.

\subsection*{Higher-Order-Dispersive Corrections to the VDE}

To better understand why we must include the higher-order corrections to the dispersion used in the VDE, we examine how such terms are found from first principles.  To leading order from the kinetic equation over an infinitely deep fluid we have that 
\[
\eta_{t} = -\mathcal{H}Q,
\]
where $q(x,t) = \phi(x,\eta(x,t),t)$, $Q = q_{x}$, and $\mathcal{H}$ is the Hilbert transform.  Likewise, from the Bernoulli equation we have to leading order that
\[
Q _{t}+ \omega\eta_{t} + \eta_{x}= 0, 
\]
so that by combining the two expressions we have the wave equation in $\eta$ alone given by 
\begin{equation}
\eta_{tt}- \omega\mathcal{H}\eta_{t} - \mathcal{H}\eta_{x}= 0.
\label{linearwaveeq}
\end{equation}
Noting that 
\[
\left(-\mathcal{H} \pd_{x}\right)^{\widehat{}} = \left|k\right|,
\]
we then define the self-adjoint pseudo-differential operator $\mathcal{L}_{+}$ with symbol 
\[
\mathcal{\widehat{L}}_{+} = \sqrt{\left|k\right| }.
\]
Multiplying by $\eta_{t}$ and integrating over space gives us the conserved quantity
\[
\int_{\mathbb{R}} \left(\frac{1}{2}\eta_{t}^{2} + \frac{1}{2}\left(\mathcal{L}_{+}\eta \right)^{2}\right) dx = \tilde{E}.
\]
Note, 
\[
\int_{\mathbb{R}}\eta_{t}\mathcal{H}\eta_{t} dx = \frac{1}{2\pi}\int_{\mathbb{R}} i\mbox{sgn}(k) \left|\hat{\eta}_{t}(k,t) \right|^{2} dk = 0,
\]
since we assume $\eta_{t}$ is real.

Letting
\[
\eta(x,t) = \eta_{1}(\xi,\tau)e^{i\theta}+\mbox{cc}, ~ \xi = \epsilon(x+c_{g}t), ~ \tau = \epsilon^{2}t, ~ \theta = k_{0}x + \Omega(\omega,k_{0})t,
\]
and using the corresponding multiple-scales expansions
\[
\pd_{x} = ik_{0} + \epsilon \pd_{\xi}, ~ \pd_{t} = i\Omega + \epsilon c_{g}\pd_{\xi} + \epsilon^{2}\pd_{\tau}, 
\]
Equation \eqref{linearwaveeq} then becomes 
\begin{multline*}
\left(\Omega^{2} -s\omega\Omega - |k_{0}|\right)\eta_{1} + 2i\epsilon\left((2\Omega-s\omega)c_{g} - s\right)\pd_{\xi}\eta_{1}\\
+ \epsilon^{2}\left(i(2\Omega-s\omega)\pd_{\tau}+c^{2}_{g}\pd^{2}_{\xi} + 2\epsilon c_{g}\pd^{2}_{\xi \tau} + \epsilon^{2}\pd^{2}_{\tau}\right)\eta_{1} = 0.
\end{multline*}
Standard NLS theory has the first two terms vanishing, thus leaving us with the slowly evolving wave equation 
\[
\left(i(2\Omega-s\omega)\pd_{\tau}+c^{2}_{g}\pd^{2}_{\xi} + 2\epsilon c_{g}\pd^{2}_{\xi \tau} + \epsilon^{2}\pd^{2}_{\tau}\right)\eta_{1} = 0.
\]
Taking a Fourier transform in space so that $\pd_{\xi}\rightarrow i\tilde{k}$, we see the above wave equation propagates information temporally at the characteristic frequencies $\lambda_{\pm}$ where
\[
\lambda_{\pm} = \frac{(2\Omega-s\omega)}{2\epsilon^{2}}\left(-\left(1+\frac{2\epsilon c_{g}\tilde{k}}{2\Omega-s\omega}\right) \pm \left(1 + \frac{4\epsilon c_{g}\tilde{k}}{2\Omega-s\omega} \right)^{1/2} \right)
\]
Likewise, proceeding as above, we can readily derive the corresponding conserved quantity
\[
\frac{s\epsilon^{2}}{\omega-2s\Omega}\int_{\mathbb{R}}\left| \pd_{\tau}\eta_{1}\right|^{2}d\xi + \alpha_{d}\int_{\mathbb{R}} \left| \pd_{\xi}\eta_{1}\right|^{2}d\xi + \frac{\epsilon\tilde{\sigma}i}{\omega-2s\Omega}\int_{\mathbb{R}}\pd_{\xi}\eta^{\ast}_{1}\pd^{2}_{\xi}\eta_{1}d\xi= \bar{E}.
\]

\end{document}

### 2019-12-18 Writeup/higher_order_dispersion_relationship_writeup.tex
Updates
cleaning house

--- 2019-12-16 Writeup/higher_order_dispersion_relationship_nost_writeup.tex
@@
 \begin{document}
-\section*{Appendix}
-\subsection*{Shear Profiles over Infinitely Deep Fluids}
 As noted in the text, the assumption that the fluid velocity is to leading order given by 
@@
 where $h_{1}>h_{0}\gg 1$, so that we are looking at a deep, continuous shear profile which is zero at or near the surface $z=\epsilon \eta(x,t)$ and at the depth $z=-h_{1}$, after which the fluid is to leading order quiescent.
-Note, we can also see this profile as satisfying to leading order two `no-slip' conditions, one near the free surface at $z=0$ and one near $z=-h_{1}$.
 
@@
 \[
-\phi_{1}(x,z,t) = \left(\alpha_{11}\cosh(kz)+\alpha_{12}\sinh(kz)\right)e^{i\theta(x,t)} + \mbox{cc}
+\phi_{1}(x,z,t) = \left(\alpha_{11}\cosh(kz)+\alpha_{12}\sinh(kz)\right)e^{i\theta(x,t)}
 \]
 \[
-\phi_{2}(x,z,t) = \left(\alpha_{21}\cosh(k(z+h_{0}))+\alpha_{22}\sinh(k(z+h_{0}))\right)e^{i\theta(x,t)} + \mbox{cc}
+\phi_{2}(x,z,t) = \left(\alpha_{21}\cosh(k(z+h_{0}))+\alpha_{22}\sinh(k(z+h_{0}))\right)e^{i\theta(x,t)}
 \]
@@
 \[
-\phi_{3}(x,z,t) = \alpha_{3}e^{|k|(z+h_{1})}e^{i\theta(x,t)} + \mbox{cc}
+\phi_{3}(x,z,t) = \alpha_{3}e^{|k|(z+h_{1})}e^{i\theta(x,t)}
 \]
-where $\theta(x,t) = kx + \Omega(k,\omega)t$, and `cc' denotes the complex conjugate.
+where $\theta(x,t) = kx + \Omega(k,\omega)t$.
 
@@
 \Omega\left(\alpha_{11}\cosh(kh_{0}) -\alpha_{12}\sinh(kh_{0}) \right) = &  \Omega \alpha_{21} + \frac{\omega h_{1}}{\delta h}\alpha_{22}\\
-\Omega\left(\alpha_{21}\cosh(k\delta h) - \alpha_{22}\sinh(k \delta h) \right) = & \left(\Omega - \frac{\omega h_{0} s}{\delta h} \right)\alpha_{3}\\
+\Omega\left(\alpha_{21}\cosh(k\delta h) - \alpha_{22}\sinh(k \delta h) \right) = & \left(\Omega - \frac{\omega h_{0} s_{k}}{\delta h} \right)\alpha_{3}\\
 \alpha_{22} = & -\alpha_{11}\sinh(kh_{0}) + \alpha_{12}\cosh(kh_{0})  \\
@@
 \end{align*}
-where $\delta h = h_{1}-h_{0}$ and $s=\mbox{sgn}(k)$.  From this we derive the dispersion relationship 
+where $\delta h = h_{1}-h_{0}$ and $s_{k}=\mbox{sgn}(k)$.  From this we derive the dispersion relationship 
 \[
@@
 \[
-\Omega_{1}(k\delta h) = \frac{s\left(\Omega -\frac{\omega h_{0}}{\delta h}s\right)+\Omega \tanh(k\delta h)}{\Omega +s\left(\Omega - \frac{\omega h_{0}}{\delta h}s\right)\tanh(k\delta h)}.
+\Omega_{1}(k\delta h) = \frac{s_{k}\left(\Omega -\frac{\omega h_{0}}{\delta h}s_{k}\right)+\Omega \tanh(k\delta h)}{\Omega +s_{k}\left(\Omega - \frac{\omega h_{0}}{\delta h}s_{k}\right)\tanh(k\delta h)}.
 \]
-While in general we would have to find the roots of a fifth-order polynomial to determine the values of $\Omega$, we see for $h_{1}\gg h_{0}$ that $\Omega_{1}\sim s$, and thus we get that the dispersion relationship simplifies to
+While in general we would have to find the roots of a fifth-order polynomial to determine the values of $\Omega$, we see for $h_{1}\gg h_{0}$ that $\Omega_{1}\sim s_{k}$, and thus we get that the dispersion relationship simplifies to
 \[
-\frac{gk}{\Omega} + \omega -\Omega \tanh(kh_{0}) + \left(\frac{1}{\Omega}\left(\frac{gk}{\Omega} + \omega\right)\tanh(kh_{0}) - 1\right)\left(\tilde{\omega} + s\Omega \right) \sim 0,
+\frac{gk}{\Omega} + \omega -\Omega \tanh(kh_{0}) + \left(\frac{1}{\Omega}\left(\frac{gk}{\Omega} + \omega\right)\tanh(kh_{0}) - 1\right)\left(\tilde{\omega} + s_{k}\Omega \right) \sim 0,
 \]
@@
 \[
-\left(2\Omega + s\tilde{\omega} \right)\left(g\left|k\right| + s\omega \Omega - \Omega^{2} \right) + \tilde{\epsilon}\left(\Omega^{3} - (s\tilde{\omega} + \Omega)(g|k|+\omega s\Omega) \right) \sim 0.
+\left(2\Omega + s_{k}\tilde{\omega} \right)\left(g\left|k\right| + s_{k}\omega \Omega - \Omega^{2} \right) + \tilde{\epsilon}\left(\Omega^{3} - (s_{k}\tilde{\omega} + \Omega)(g|k|+\omega s_{k}\Omega) \right) \sim 0.
 \]
@@
 \[
-\Omega \sim \frac{1}{2}\left(-s\omega \pm \sqrt{\omega^{2}+4g|k|} \right), ~ -\frac{s\tilde{\omega}}{2}.
+\Omega \sim \frac{1}{2}\left(-s_{k}\omega \pm \sqrt{\omega^{2}+4g|k|} \right), ~ -\frac{s_{k}\tilde{\omega}}{2}.
 \]
-Thus of the three roots we find, two give the dispersion relationship we find in the body of the text using ${\bf u}\sim \omega z \hat{{\bf i}}$.
-Looking at the affiliated disturbances of the relevant free surface and internal wave, we find that
-\[
-\eta_{1} \sim \alpha_{12}\frac{ik}{\Omega}e^{i\theta} + \mbox{cc}, ~ \eta_{2} \sim \alpha_{12}\frac{ik}{\Omega^{3}}\cosh(kh_{0})\left(\Omega^{2}-(1-\tilde{\epsilon})\left( g|k|+s\omega\Omega\right)\right)e^{i\theta}+ \mbox{cc}.
-\]
-Thus if we choose $\Omega$ so that $\Omega^{2}-\left( g|k|+s\omega\Omega\right)= 0$, the magnitude of the internal wave essentially vanishes, thereby localizing dynamics along the free surface near $z=0$.
-Therefore, while not necessarily physically justifiable throughout the bulk of the fluid, our simplified shear profile assumption produces results which are asymptotically consistent with a more sophisticated treatment of the shear profile.
-
-We note though that the more physically realistic shear profile shows that there is a choice of $\Omega$ which corresponds to a non-trivial internal mode near $z=-h_{0}$.
-We also note that this choice of $\Omega$ makes many of the terms in our modulation theory singular, thus showing that this case would require a markedly different treatment.
-While interesting, this issue is beyond the scope of the current paper and will be addressed in future research.
-
-\subsection*{Higher-Order-Dispersive Corrections to the VDE}
-
-To better understand why we must include the higher-order corrections to the dispersion used in the VDE, we examine how such terms are found from first principles.
-To leading order from the kinetic equation over an infinitely deep fluid we have that
-\[
-\eta_{t} = -\mathcal{H}Q,
-\]
-where $q(x,t) = \phi(x,\eta(x,t),t)$, $Q = q_{x}$, and $\mathcal{H}$ is the Hilbert transform.  Likewise, from the Bernoulli equation we have to leading order that
-\[
-Q _{t}+ \omega\eta_{t} + \eta_{x}= 0, 
-\]
-so that by combining the two expressions we have the wave equation in $\eta$ alone given by 
-\begin{equation}
-\eta_{tt}- \omega\mathcal{H}\eta_{t} - \mathcal{H}\eta_{x}= 0.
-\label{linearwaveeq}
-\end{equation}
-Noting that 
-\[
-\left(-\mathcal{H} \pd_{x}\right)^{\widehat{}} = \left|k\right|,
-\]
-we then define the self-adjoint pseudo-differential operator $\mathcal{L}_{+}$ with symbol 
-\[
-\mathcal{\widehat{L}}_{+} = \sqrt{\left|k\right| }.
-\]
-Multiplying by $\eta_{t}$ and integrating over space gives us the conserved quantity
-\[
-\int_{\mathbb{R}} \left(\frac{1}{2}\eta_{t}^{2} + \frac{1}{2}\left(\mathcal{L}_{+}\eta \right)^{2}\right) dx = \tilde{E}.
-\]
-Note, 
-\[
-\int_{\mathbb{R}}\eta_{t}\mathcal{H}\eta_{t} dx = \frac{1}{2\pi}\int_{\mathbb{R}} i\mbox{sgn}(k) \left|\hat{\eta}_{t}(k,t) \right|^{2} dk = 0,
-\]
-since we assume $\eta_{t}$ is real.
-
-Letting
-\[
-\eta(x,t) = \eta_{1}(\xi,\tau)e^{i\theta}+\mbox{cc}, ~ \xi = \epsilon(x+c_{g}t), ~ \tau = \epsilon^{2}t, ~ \theta = k_{0}x + \Omega(\omega,k_{0})t,
-\]
-and using the corresponding multiple-scales expansions
-\[
-\pd_{x} = ik_{0} + \epsilon \pd_{\xi}, ~ \pd_{t} = i\Omega + \epsilon c_{g}\pd_{\xi} + \epsilon^{2}\pd_{\tau}, 
-\]
-Equation \eqref{linearwaveeq} then becomes 
-\begin{multline*}
-\left(\Omega^{2} -s\omega\Omega - |k_{0}|\right)\eta_{1} + 2i\epsilon\left((2\Omega-s\omega)c_{g} - s\right)\pd_{\xi}\eta_{1}\\
-+ \epsilon^{2}\left(i(2\Omega-s\omega)\pd_{\tau}+c^{2}_{g}\pd^{2}_{\xi} + 2\epsilon c_{g}\pd^{2}_{\xi \tau} + \epsilon^{2}\pd^{2}_{\tau}\right)\eta_{1} = 0.
-\end{multline*}
-Standard NLS theory has the first two terms vanishing, thus leaving us with the slowly evolving wave equation 
-\[
-\left(i(2\Omega-s\omega)\pd_{\tau}+c^{2}_{g}\pd^{2}_{\xi} + 2\epsilon c_{g}\pd^{2}_{\xi \tau} + \epsilon^{2}\pd^{2}_{\tau}\right)\eta_{1} = 0.
-\]
-Taking a Fourier transform in space so that $\pd_{\xi}\rightarrow i\tilde{k}$, we see the above wave equation propagates information temporally at the characteristic frequencies $\lambda_{\pm}$ where
-\[
-\lambda_{\pm} = \frac{(2\Omega-s\omega)}{2\epsilon^{2}}\left(-\left(1+\frac{2\epsilon c_{g}\tilde{k}}{2\Omega-s\omega}\right) \pm \left(1 + \frac{4\epsilon c_{g}\tilde{k}}{2\Omega-s\omega} \right)^{1/2} \right)
-\]
-Likewise, proceeding as above, we can readily derive the corresponding conserved quantity
-\[
-\frac{s\epsilon^{2}}{\omega-2s\Omega}\int_{\mathbb{R}}\left| \pd_{\tau}\eta_{1}\right|^{2}d\xi + \alpha_{d}\int_{\mathbb{R}} \left| \pd_{\xi}\eta_{1}\right|^{2}d\xi + \frac{\epsilon\tilde{\sigma}i}{\omega-2s\Omega}\int_{\mathbb{R}}\pd_{\xi}\eta^{\ast}_{1}\pd^{2}_{\xi}\eta_{1}d\xi= \bar{E}.
-\]
-
+Thus we get back the dispersion relationship we would expect.
+However, we also find another mode which would correspond to an infinite group velocity for the modulated wave packets we study via the NLSE and VDE.
 \end{document}

### 2019-12-19 Writeup/higher_order_dispersion_relationship_nost_writeup.tex

@@
 \end{equation}
-Noting that 
+Using Fourier transforms, we can readily write a solution to the affiliated initial-value problem in the form
 \[
-\left(-\mathcal{H} \pd_{x}\right)^{\widehat{}} = \left|k\right|,
+\eta(x,t) = \frac{1}{2\pi}\int_{\mathbb{R}}\hat{\eta}_{0}(k) e^{ikx+i\Omega_{+}(k)t} dk + \frac{1}{2\pi}\int_{\mathbb{R}}\hat{\eta}_{0}(k) e^{ikx+i\Omega_{-}(k)t} dk
 \]
-we then define the self-adjoint pseudo-differential operator $\mathcal{L}_{+}$ with symbol 
+where
 \[
-\mathcal{\widehat{L}}_{+} = \sqrt{\left|k\right| }.
+\Omega_{\pm}(k) = \frac{1}{2}\left(\omega s \pm \sqrt{\omega^{2}+4|k|} \right).
 \]
-Multiplying by $\eta_{t}$ and integrating over space gives us the conserved quantity
+If we assume, as per usual when deriving NLS-type models, that 
 \[
-\int_{\mathbb{R}} \left(\frac{1}{2}\eta_{t}^{2} + \frac{1}{2}\left(\mathcal{L}_{+}\eta \right)^{2}\right) dx = \tilde{E}.
+\hat{\eta}_{0}(k) = \frac{1}{\epsilon}\hat{A}\left(\frac{k-k_{0}}{\epsilon} \right), 
 \]
-Note, 
+representing the assumption that the initial conditions are narrowly banded around a carrier wave number $k_{0}$, then we see that, taking only the plus branch of $\Omega_{\pm}$, that  
+\begin{align*}
+\eta(x,t) = & \frac{e^{i\theta(x,t,k_{0})}}{2\pi}\int_{\mathbb{R}}\hat{A}(\tilde{k})e^{i\tilde{k}\xi} e^{i\tau \tilde{\Omega}(k_{0},\tilde{k})}d\tilde{k}\\
+\sim & \frac{e^{i\theta(x,t,k_{0})}}{2\pi}\int_{\mathbb{R}}\hat{A}(\tilde{k})e^{i\tilde{k}\xi}e^{i\tau\left(\frac{\Omega''(k_{0})}{2}\tilde{k}^{2} + \frac{\epsilon}{6}\Omega'''(k_{0})\tilde{k}^{3} + \frac{\epsilon^{2}}{24}\Omega''''(k_{0})\tilde{k}^{4}\right)}d\tilde{k},
+\end{align*}
+where
 \[
-\int_{\mathbb{R}}\eta_{t}\mathcal{H}\eta_{t} dx = \frac{1}{2\pi}\int_{\mathbb{R}} i\mbox{sgn}(k) \left|\hat{\eta}_{t}(k,t) \right|^{2} dk = 0,
+\theta(x,t,k_{0}) = k_{0}x + \Omega(k_{0})t, ~c_{g}(k_{0}) = \Omega'(k_{0}), ~ \xi = \epsilon(x+c_{g}t), ~ \tau = \epsilon^{2}t, 
 \]
-since we assume $\eta_{t}$ is real.
-
-Letting
+and where
 \[
-\eta(x,t) = \eta_{1}(\xi,\tau)e^{i\theta}+\mbox{cc}, ~ \xi = \epsilon(x+c_{g}t), ~ \tau = \epsilon^{2}t, ~ \theta = k_{0}x + \Omega(\omega,k_{0})t,
+\tilde{\Omega}(k_{0},\tilde{k}) = \frac{1}{\epsilon^{2}}\left(\Omega(k_{0}+\epsilon \tilde{k})-\Omega(k_{0})-\Omega'(k_{0})\epsilon\tilde{k}\right) 
 \]
-and using the corresponding multiple-scales expansions
+and where we have expanded the dispersion relationship up to $\mathcal{O}(\epsilon^{3})$ as is done in the main body of the text.
+This expansion corresponds to the affiliated linear-evolution equation for the slowly-evolving envelope $A(\xi,\tau)$
 \[
-\pd_{x} = ik_{0} + \epsilon \pd_{\xi}, ~ \pd_{t} = i\Omega + \epsilon c_{g}\pd_{\xi} + \epsilon^{2}\pd_{\tau}, 
-\]
-Equation \eqref{linearwaveeq} then becomes 
-\begin{multline*}
-\left(\Omega^{2} -s\omega\Omega - |k_{0}|\right)\eta_{1} + 2i\epsilon\left((2\Omega-s\omega)c_{g} - s\right)\pd_{\xi}\eta_{1}\\
-+ \epsilon^{2}\left(i(2\Omega-s\omega)\pd_{\tau}+c^{2}_{g}\pd^{2}_{\xi} + 2\epsilon c_{g}\pd^{2}_{\xi \tau} + \epsilon^{2}\pd^{2}_{\tau}\right)\eta_{1} = 0.
-\end{multline*}
-Standard NLS theory has the first two terms vanishing, thus leaving us with the slowly evolving wave equation 
-\[
-\left(i(2\Omega-s\omega)\pd_{\tau}+c^{2}_{g}\pd^{2}_{\xi} + 2\epsilon c_{g}\pd^{2}_{\xi \tau} + \epsilon^{2}\pd^{2}_{\tau}\right)\eta_{1} = 0.
-\]
-Taking a Fourier transform in space so that $\pd_{\xi}\rightarrow i\tilde{k}$, we see the above wave equation propagates information temporally at the characteristic frequencies $\lambda_{\pm}$ where
-\[
-\lambda_{\pm} = \frac{(2\Omega-s\omega)}{2\epsilon^{2}}\left(-\left(1+\frac{2\epsilon c_{g}\tilde{k}}{2\Omega-s\omega}\right) \pm \left(1 + \frac{4\epsilon c_{g}\tilde{k}}{2\Omega-s\omega} \right)^{1/2} \right)
-\]
-Likewise, proceeding as above, we can readily derive the corresponding conserved quantity
-\[
-\frac{s\epsilon^{2}}{\omega-2s\Omega}\int_{\mathbb{R}}\left| \pd_{\tau}\eta_{1}\right|^{2}d\xi + \alpha_{d}\int_{\mathbb{R}} \left| \pd_{\xi}\eta_{1}\right|^{2}d\xi + \frac{\epsilon\tilde{\sigma}i}{\omega-2s\Omega}\int_{\mathbb{R}}\pd_{\xi}\eta^{\ast}_{1}\pd^{2}_{\xi}\eta_{1}d\xi= \bar{E}.
+i\pd_{\tau}A = \left(\frac{\Omega''(k_{0})}{2}\pd_{\xi}^{2}-\frac{i\epsilon}{6}\Omega'''(k_{0})\pd_{\xi}^{3} - \frac{\epsilon^{2}}{24}\Omega''''(k_{0})\pd_{\xi}^{4}\right)A. 
 \]
 
+As to the issue of why to include the $\epsilon^{2}$ term, we note that the affiliated group velocity of the slowly-evolving envelope is given by 
+\begin{align*}
+\frac{d}{d\tilde{k}}\tilde{\Omega} = & \frac{1}{\epsilon}\left(\Omega'(k_{0}+\epsilon \tilde{k}) - \Omega'(k_{0})\right)\\
+\sim & \Omega''(k_{0})\tilde{k}\left(1 - \frac{3\epsilon c_{g}^{2}}{s}\tilde{k} + 10 \epsilon^{2}c_{g}^{4}\tilde{k}^{2}\right).
+\end{align*}
+We see that the full dispersion relationship has exactly one root at $\tilde{k}=0$, so that this root is the only point of stationary phase in the corresponding integral representation of the solution $\eta(x,t)$.
+However, using our Taylor series expansion, were we to ignore the $\epsilon^{2}$ term, we would introduce a new stationary-phase point at $\tilde{k}_{\ast}$ where
+\[
+\tilde{k}_{\ast} = \frac{s}{3\epsilon c_{g}^{2}}.
+\]
+The presence of this stationary-phase point allows for the accumulation of energy at high-frequencies, which without a corresponding band-limiting requirement introduces non-physical effects into the model.
+By including the $\epsilon^{2}$ term though, we remove this spurious high-frequency stationary-phase point, thereby avoiding the inclusion of an otherwise unnecessary bandwidth condition on the slowly-evolving envelope.
 \end{document}

### 2019-12-19 Writeup/random_vorticity.tex

--- 2019-04-10 Writeup/random_vorticity.tex
@@
 
-\title{Evolution of Narrow-Band Spectra in Deep-Water Constant Vorticity Flows}
+\title{Evolution of Spectral Distributions in Deep-Water Constant Vorticity Flows}
 \author{Christopher W. Curtis and Mackensie Murphy}
@@
 \maketitle
-\section*{Introduction}
-There is now a relatively wide range of literature which shows that nonlinear instabilities, in particular the modulational instability (MI), are responsible for significant modifications to the statistical properties of water waves; see \cite{dysthe2,dysthe3,onorato,thomas2012nonlinear} among others.
-These results have significant impact on the understanding of deep-water rogue wave formation among other oceanographic phenomena.
-However, much of our understanding around MI relies on highly idealized assumptions which cannot be expected to hold in natural settings.
+\section*{Abstract}
+A central question in sea-state modeling is the role that various physical effects have on the evolution of the statistical properties of random sea states.
+This becomes a critical issue when one is concerned with the likelihood of rare events such as rogue, or freak, waves which can have significant destructive potential on deep sea ships and other offshore structures.
+In this paper then, using a recently derived higher-order model of deep water nonlinear waves, we examine the impact of constant vorticity currents on the statistical properties of nonlinearly evolving random sea states.
+As we show, these currents can both decrease and increase the kurtosis of the affiliated distributions of the sea states, thereby diminishing or enhancing the likelihood of rare events.
+We likewise numerically study the relationship between the kurtosis and a non-dimensional parameter, the Benjamin--Feir Index, which has proven to be a useful measure of when rare events are likely in oceanographic application.
+
+\section{Introduction}
+There is now a wide range of literature which shows that nonlinear instabilities, in particular the modulational instability (MI), are responsible for significant modifications to the statistical properties of water waves; see \cite{dysthe2,dysthe3,onorato,thomas2012nonlinear} among others.
+As explained in \cite{janssen}, MI is especially important due to the relative rapidity with which it acts, being orders of magnitude faster than say four-wave resonant interactions which form the traditional backbone of our understanding of how nonlinearity drives changes in statistics of random-wave fields \cite{hasselmann,holthuijsen}.
+These results have significant impact on the understanding of deep-water-rogue wave formation among other oceanographic phenomena.
+
+However, much of the understanding around MI relies on highly idealized assumptions which cannot be expected to hold in natural settings.
 In particular, most understandings of MI rely on looking at the stability of perturbations to carrier waves at essentially a fixed wave number.
-To address this shortcoming, in a now seminal paper, \cite{alber} analytically studies families of perturbations of wave packets with narrow, but non-zero, spectral width around a central wave number.
-An analytic critierion determining when modulational instability (MI) is either manifested or supressed depending on this spectral width is derived.
-The results in \cite{alber} were confirmed numerically in \cite{dysthe2,dysthe3} by examing the mean properties associated with ensembles of initial conditions.
+To address this shortcoming, in now seminal papers, \cite{alber} and \cite{alber2} analytically studied families of perturbations of wave packets with narrow, but non-zero, spectral width around a central wave number.
+Using the Nonlinear Schr\"{o}dinger equation (NLSE), an analytic criterion determining when MI is either manifested or suppressed depending on this spectral width, now referred to as a Benjamin--Feir index (BFI) was derived.
+The results in \cite{alber} were explored numerically in \cite{dysthe2,dysthe3} by examining the mean properties associated with ensembles of initial conditions.
 
-However, many physical effects were ignored in the derivation and verification of the results described above.
-In particular, vorticity was not taken into account, thereby ignoring a central mechanism for the transfer of energy across length scales in oceanic flows.
-Thus, building on the work in \cite{thomas2012nonlinear} and \cite{curtis8}, we numerically explore the analytic stability condition derived in \cite{alber}, focusing in particular on the impact that vorticity plays in controlling the effective width for the onset of MI.
-Moving beyond just examining the properties of the NLS equation with vorticity, we look at the statistical properties of solutions to a higher-order model, the vor-Dysthe equation derived in \cite{curtis8}.
-As shown in \cite{onorato}, the higher-order terms associated with the Dysthe equation can have significant impacts on the statistics of the waves, so it is a nontrivial question to examine how the vorticity will interact with these higher-order nonlinearities.
+As shown in \cite{janssen} though, the movement of eigenvalues and corresponding bifurcation in stability predicted in \cite{alber} appears at best in an ambiguous fashion.
+Regardless though, the BFI was still shown in \cite{janssen} to be an especially useful parameter since it allows for a ready prediction of the kurtosis of a random wave field as a quadratic function of the BFI.
+Thus, the BFI becomes a useful predictive tool for the appearance of rare events, such as rogue waves in a random-nonlinear-wave field.
+Experimental explorations of the results in \cite{janssen} and \cite{alber} were undertaken in \cite{onorato}, where the role of using higher-order nonlinear models than the NLSE were examined as well.
 
-\section*{Modeling and Stability Theory}
-We examine the unsteady nonlinear wave propagation over a constant shear current.  To do this, we assume the fluid velocity has the form
+Later work has simulated the impact of wind and damping on the interplay between the BFI and the kurtosis of a wave field; see \cite{slunyaev,eeltink}.
+However, the impact of vorticity was not taken into account in these works, thereby ignoring a central mechanism for the transfer of energy across length scales in oceanic flows.
+Moreover, as shown recently in \cite{curtis8}, linear shear profiles can have strong impacts on the behavior of mean properties of surface wave flows with some shear profiles being able to suppress the Stokes Drift at the surface.
+Therefore while a relatively idealized case of shear, the interplay between nonlinearity and constant vorticity can manifest in markedly different physics.
+Building then on the work in \cite{thomas2012nonlinear} and \cite{curtis8}, and complementing the results in \cite{touboul}, we numerically explore the statistical properties of random-nonlinear-wave fields moving over deep water and linear shear profiles.
+Moving beyond just examining the properties of the NLSE with vorticity, we look at the statistical properties of solutions to a higher-order model, the vor-Dysthe equation (VDE) derived in \cite{curtis8}.
+As shown in \cite{onorato,janssen} and \cite{eeltink}, the higher-order terms associated with the Dysthe equation can have significant impacts on the statistics of the waves, so it is a nontrivial question to examine how the vorticity interacts with these higher-order nonlinearities.
+
+Using our simulations, starting from normally distributed random-wave fields, we show that the BFI condition with respect to predicting stability is not especially useful, conforming to the results in \cite{janssen}.
+While we are able to show that for sufficiently wide enough initial distributions that the temporal modifications to the profile are slight, whether or not this corresponds to an actual bifurcation in stability or simply reflects that for a wide enough profile the sideband growth affiliated with the MI has little impact is not clear.
+We note though that classic MI analysis does provide a useful prediction of the time scale over which the most significant dynamics in the statistics of the wave field occur.
+
+Beyond the issue of MI, we see several interesting results from our work.
+In particular, we see that increasing vorticity enhances the kurtosis of the random-wave field.
+Moreover, the VDE always predicts a larger kurtosis than the NLSE, with this difference between the models exacerbated by increasing vorticity.
+Likewise the affiliated statistics of the spectral distribution show marked differences between the two models, with the VDE tending to strongly skew distributions away from Gaussian conditions through peak down shifting and one sided tail enhancement.
+Finally, we study the behavior of the kurtosis as a function of the BFI parameter.
+The quadratic relationship found in \cite{janssen} and \cite{eeltink} is shown to hold, though the nature of the fit is clearly dependent on the vorticity, conforming to the dependence of the fit on physical parameters found in \cite{eeltink}.
+
+ Thus we show that vorticity and higher-order deep water models can have significant impacts on our understanding of the likelihood of rare events, like rogue waves, in deep-water-surface flows.
+In particular, shear profiles which tend to propagate in the direction of the surface wave-field acting on relatively narrow distributions of waves can create strongly non-Gaussian distributions with significantly enhanced tails thereby encouraging the occurrence of rare events.
+These results motivate further study of currents in random-wave fields, especially as they pertain to the presence of wind and wave dissipation.
+
+The structure of the paper is an explanation of the model and discussion of the BFI in Section 2.  Section 3 presents our numerical results, while Section 4 presents a conclusion.
+
+\section{Modeling and Stability Theory}
+We examine the unsteady nonlinear wave propagation over a constant shear current.  To do this, we suppose the fluid velocity is of the form
 \[
@@
 We restrict fluid motion to the $(x,z)$-plane, thereby ignoring transverse variations in the $y$ dimension.
-Following standard arguments, e.g.~see \cite{ashton}, the dynamics of the fluid can be determined by solving the free boundary value problem
+Using well established arguments, e.g.~see \cite{ashton}, the dynamics of the fluid are found by solving the free boundary value problem
 \begin{align}
@@
 where $\eta$ represents the free surface displacement, and $g$, $\rho$, and $\sigma$ represent the acceleration due to gravity, the fluid density, and the coefficient of surface tension respectively.
-
-By choosing a characteristic wave height $a$ and wave length $L$, all quantities can be non-dimensionalized via
+The system is non-dimensionalized via the choices
 \begin{align*}
-\tilde{x} = \frac{x}{L}, ~\tilde{z} = \frac{z}{L}, ~ \tilde{t} = \sqrt{\frac{g}{L}}~t, ~\omega = \sqrt{\frac{g}{L}}~\tilde{\omega}, \\
-\eta = a \tilde{\eta}, ~ \phi  = a\sqrt{gL}~\tilde{\phi} , ~ \epsilon = \frac{a}{L}.
+\tilde{x} = \frac{x}{L_{p}}, ~\tilde{z} = \frac{z}{L_{p}}, ~ \tilde{t} = \sqrt{\frac{g}{L_{p}}}~t, ~\omega = \sqrt{\frac{g}{L_{p}}}~\tilde{\omega}, \\
+\eta = a \tilde{\eta}, ~ \phi  = a\sqrt{gL_{p}}~\tilde{\phi} , ~ \epsilon = \frac{a}{L_{p}}.
 \end{align*}
-By following the AFM approach described in \cite{afm,ashton} and dropping the tildes, the kinematic boundary condition, Equation \eqref{kine}, can be written in terms of surface variables alone via the integro-differential equation
-\begin{equation}
-\int_{\mathbb{R}}dx~e^{-ikx}e^{\epsilon|k|\eta}\left(\eta_{t} +
-  \epsilon \omega \eta \eta_{x} + i \mbox{sgn}(k)Q \right) = 0, ~
-k\neq 0,
-\label{kine2}
-\end{equation}
-where $Q=q_{x}$, $q(x,t) = \phi(x,\eta(x,t),t)$.
-By integrating over $\mathbb{R}$, we are assuming that both $\eta$ and $Q$ decay to zero sufficiently rapidly in the far field.
-We can also readily derive a nearly identical expression on domains periodic in the horizontal variable $x$, in which case Equation \eqref{kine2} becomes
-\begin{equation}
-\int_{-L_{p}/2}^{L_{p}/2}dx~e^{-ikx}e^{\epsilon|k|\eta}\left(\eta_{t} +
-  \epsilon \omega \eta \eta_{x} + i \mbox{sgn}(k)Q \right) = 0, ~ k = \frac{2\pi m}{L_{p}}, ~ m\in \mathbb{Z}\backslash{0},
-\label{integro1per}
-\end{equation}
-where $L_p$ is the spatial period.  Throughout the remainder of this section we only present results over the real line, $\mathbb{R}$, since identical results can be derived for the periodic case.
-Lastly, we note that if in Equation \eqref{kine2} we approach $k=0$ from both the left and the right, we get the equations
-\[
-\int_{\mathbb{R}}dx \left(\eta_{t} + i Q \right) = 0, ~ \int_{\mathbb{R}}dx \left(\eta_{t} - i Q \right) = 0.
-\]
-Thus we get the identities
-\begin{align}
-\pd_{t}\int_{\mathbb{R}} dx ~\eta(x,t) & = 0, \label{zavsurf}\\
-\int_{\mathbb{R}} dx ~Q(x,t) & = 0. \label{zavpot}
-\end{align}
-
-Taylor expanding Equation (\ref{kine2}) up to $\mathcal{O}(\epsilon^3)$ for $k\neq 0$ gives
-\begin{multline}
-\int_{\mathbb{R}}dx~e^{-ikx}\left(1 + \epsilon |k|\eta + \frac{\epsilon^{2}|k|^{2}\eta^{2}}{2} \right)\left(\eta_{t} + i \mbox{sgn}(k)Q \right) \\
-+ \epsilon \omega \int_{\mathbb{R}}dx~e^{-ikx}\left(1 + \epsilon |k|\eta \right) \eta \eta_{x} = 0.
-\label{integro1}
-\end{multline}
-Transforming into surface variables and Taylor expanding Equation \eqref{bernou}, Bernoulli's equation, up to $\mathcal{O}(\epsilon^3)$ gives
-\begin{multline}
-Q_{t} + \omega \eta_{t} + \eta_{x} - \tilde{\sigma}\eta_{xxx} + \frac{\epsilon}{2}\pd_{x}\left(-\eta_{t}^{2} + (Q+\omega \eta)^{2} \right)\\
-+ \epsilon^{2}\pd_{x}\left( \frac{3}{2}\tilde{\sigma}\eta_{x}^{2}\eta_{xx} - \eta_{t}\eta_{x} \left(Q  + \omega \eta \right) \right)  = 0,
-\label{berexp}
-\end{multline}
-where the reciprocal of the Bond number, $\tilde{\sigma}$, is given by
-\[
-\tilde{\sigma} = \frac{\sigma}{\rho g L^2}.
-\]
-We note that we have tacitly assumed $\omega = \mathcal{O}(1)$.
-In physical terms, this implies that $\omega$ is comparable to the natural time scale of this problem, $\sqrt{L/g}$.
-If we were to assume $\omega$ were of larger magnitude, the problem would no longer be weakly nonlinear and would be much less amenable to asymptotic analysis.
+Note that we are assuming $\omega = \mathcal{O}(1)$ throughout the remainder of the paper.
+In physical terms, this implies that $\omega$ is comparable to the natural time scale of this problem, $\sqrt{L_{p}/g}$.
+If we were to assume $\omega$ were of larger magnitude, the problem would no longer be weakly nonlinear and asymptotic analysis would be siginificantly more difficult.
 Therefore, throughout the remainder of the paper, we assume that the vorticity is not too large.
 
-By expanding Equation \eqref{integro1} and taking inverse Fourier transforms, we can derive the first three terms of the corresponding Dirichlet-to-Neumann Operator (DNO) expansion, so that we find
-\begin{equation}
-\eta_{t} = G_{0}Q + \epsilon G_{1}Q + \epsilon^{2}G_{2}Q,
-\label{dnoexp}
-\end{equation}
-where
-\begin{align*}
-G_{0}Q = & -\mathcal{H}Q\\
-G_{1}Q = & \mathcal{H}\pd_{x}\left(\eta G_{0}Q\right) - \pd_{x}\left(\eta Q + \frac{\omega}{2}\eta^{2}\right)\\
-G_{2}Q = & \mathcal{H}\pd_{x}\left(\eta G_{1}Q\right) + \frac{1}{2}\pd_{x}^{2}\left(\eta^{2}G_{0}Q \right) + \pd_{x}^{2}\mathcal{H}\left(\frac{1}{2}\eta^{2}Q + \frac{\omega}{3}\eta^{3}\right), 
-\end{align*}
-where $\mathcal{H}$ denotes the Hilbert transform so that 
-\[
-\mathcal{H}f = \frac{1}{2\pi}\int_{\mathbb{R}}dk~ e^{ikx}  i\mbox{sgn}(k) \hat{f}(k),
-\]
-where $\hat{f}(k)$ is the Fourier transform of $f(x)$ defined by
-\[
-\hat{f}(k) = \int_{\mathbb{R}}dx~ e^{-ikx}f(x).
-\]
-Using the DNO expansion in Equation \eqref{dnoexp}, we can rewrite Equation \eqref{berexp} an an evolution equation for $Q_{t}$, thereby giving us a closed system in terms of surface variables alone.
-The system of equations given by Equations \eqref{dnoexp} and \eqref{berexp} we denote as the Full-Euler Equations (FEEs).
-
-As is well known by now, in non-dimensional variables, we can model the evolution of a relatively small free-fluid surface over infinitely-deep water $z=\epsilon \eta(x,t)$ via the expansion 
+For a relatively small free-fluid surface over infinitely-deep water $z=\epsilon \eta(x,t)$, we model a modulated-wave packet via the expansion 
 \begin{multline*}
@@
 \]
-Here, $s = \mbox{sgn}(k_{0})$, $\omega$ is the non-dimensionalized magnitude of the vorticity of the flow, and $\tilde{\sigma}$ is the surface tension.
+We note that the form for $\phi$ follows directly from this expansion, allowing for everything to be modeled by the terms in the expansion in $\eta$.
+Here, $s = \mbox{sgn}(k_{0})$, $\omega$ is the non-dimensionalized magnitude of the vorticity of the flow, and $\tilde{\sigma}$ is the surface tension, or where the reciprocal of the Bond number, $\tilde{\sigma}$, is given by
+\[
+\tilde{\sigma} = \frac{\sigma}{\rho g L_{p}^2}.
+\]
+Throughout the remainder of this paper, we take the positive branch $+$ of the dispersion relationship, thereby ensuring that $\Omega > 0$.
+Thus, for $k_{0}>0$, the surface waves to leading order propagate to the left.
+Thus, if $\omega>0$, then throughout the bulk of the fluid, the shear current propagates with the surface waves, or is {\it co-propagating}, while for $\omega<0$, we say the shear current is {\it counter-propagating}.
 
-To describe the evolution of the slowly evolving envelope $\eta_{1}(\xi,\tau)$, we use two asymptotically accurate models.
+To describe the evolution of the slowly evolving envelope $\eta_{1}(\xi,\tau)$, we use two models.
 The first model is the the nonlinear Schr\"{o}dinger equation (NLSE) derived in \cite{curtis8}, though see also \cite{thomas2012nonlinear}, given by
@@
 \end{align*}
+However, the NLSE is a relatively low-order model known to be only of limited use in real-world modeling situations.
+A more robust model, following the work in \cite{dysthe1}, is found by going to the next higher order in $\epsilon$, giving us the equation
+\begin{multline}
+(\omega-2s\Omega)\pd_{\tau}\eta_{1} + i(c_{g}^{2}s-3k_{0}\tilde{\sigma})\pd_{\xi}^{2}\eta_{1} + 2i\epsilon s c_{g}\pd^{2}_{\xi\tau}\eta_{1} - \epsilon\tilde{\sigma} \pd^{3}_{\xi}\eta_{1} + i\epsilon^{2}c_{h}\pd_{\xi}^{4}\eta_{1} \\
+- i \alpha_{nl}(\omega-2s\Omega)\left|\eta_{1}\right|^{2}\eta_{1}
++ \epsilon \tilde{\alpha}_{1}\left|\eta_{1}\right|^{2}\pd_{\xi}\eta_{1}\\
++ \epsilon\tilde{\alpha}_{2}\eta_{1}^{2}\pd_{\xi}\eta^{\ast}_{1} +
+\epsilon i\tilde{\alpha}_{3}\eta_{1}\mathcal{H}\pd_{\xi}\left|\eta_{1}\right|^{2}= 0.
+\label{dysthe}
+\end{multline}
+where $\mathcal{H}$ denotes the Hilbert transform so that 
+\[
+\mathcal{H}f = \frac{1}{2\pi}\int_{\mathbb{R}}dk~ e^{ikx}  i\mbox{sgn}(k) \hat{f}(k),
+\]
+and where $\hat{f}(k)$ is the Fourier transform of $f(x)$ defined by
+\[
+\hat{f}(k) = \int_{\mathbb{R}}dx~ e^{-ikx}f(x).
+\]
 
-To generate simulations, we start both the NLSE and FEE from the same collection of $2L$ periodic-initial conditions
+We call Equation \eqref{dysthe} the Vor-Dysthe equation (VDE).
+Details of its derivation can be found in \cite{curtis8}, though we note a difference in this work with that in \cite{curtis8}.
+In this work, we found the inclusion of the higher-order dispersion associated with the coefficient $c_{h}$, given by
+\[
+c_{h} = 2sc_{g}(2sc_{g}\alpha_{d}+\tilde{\sigma})/(\omega-2s\Omega) - s\alpha_{d}^{2},
+\]
+was necessary to remove artificial numerical instabilities.
+Similar issues with Dysthe like equations have been noted in \cite{eeltink} among other sources.
+Moreover, the coefficients $\tilde{\alpha}_{j}$ are readily derived, but their forms are cumbersome and thus we omit explicitly writing them for the sake of readability.
+
+To generate simulations, we start both the NLSE and VDE from the same collection of $2L$ periodic-initial conditions
 \[
@@
 \]
-where $\delta \tilde{k}=\pi/L$ and the phases $\theta_{k}$ are randomly chosen uniformly between $0$ and $2\pi$.  We can then readily show that, for $L\gg 1$, 
+where $\delta \tilde{k}=\pi/L$ and the phases $\theta_{k}$ are randomly and independently chosen uniformly between $0$ and $2\pi$.  We can then readily show that, for $L\gg 1$, 
 \[
-\epsilon_{rms} \approx \overline{\left|\eta_{1}\right|^{2}}^{1/2},
+\epsilon_{rms} \approx \overline{\left|\eta_{1}(\xi,0)\right|^{2}}^{1/2},
 \]
 where $\overline{()}$ denotes the ensemble average.
-In our coordinates then, the work in \cite{alber} and later confirmed in \cite{dysthe2} shows that for the NLSE the MI is suppressed for spectral widths $\sigma$ such that
+In our coordinates then, motivated by the work of \cite{alber} and \cite{dysthe2} and more fully explored in \cite{janssen}, we study the stability of the initial wavepacket distributions with regards to the magnitude of the Benjamin--Feir Index (BFI), where the BFI is defined to be
 \[
-\sigma \geq \epsilon_{rms}\sqrt{\frac{2\alpha_{nl}}{\alpha_{d}}}
+\mbox{BFI} = \sqrt{\frac{\alpha_{nl}}{\alpha_{d}}}\frac{\epsilon_{rms}}{\sigma}.
 \]
-This work has helped motivate defining this parameter as the Benjamin--Feir Index (BFI), which has been shown to be a critical parameter to understanding the statstical properties of nonlinear free-surface flows; see \cite{onorato, thomas2012nonlinear}.
-\section*{Results}
-To understand the relationship between the vorticity $\omega$ and the initial spectral width $\sigma$ of our initial conditions, we generate ensembles of initial conditions and then run these different initial conditions up to $\tau=1$, which corresponds to $t_{f} = 1/\epsilon^{2}$ in the original fast variables.
-The domain is $[-50\pi,50\pi]$.
-The number of modes is $K = 256$.
+Note, this definition is identical with the spectral width criterion of \cite{alber} and the BFI condition given in \cite{thomas2012nonlinear}; see also \cite{onorato}.
+It was claimed in \cite{alber}, and in some sense verified in \cite{dysthe2}, that wave packets should be stable with respect to MI's for $\mbox{BFI}<1$, which corresponds to the initial spectral width $\sigma>\sigma_{c}$, where
+\[
+\sigma_{c} = \sqrt{\frac{\alpha_{nl}}{\alpha_{d}}}\epsilon_{rms}.
+\]
+
+However, \cite{alber} does not present numerical verification of their results, and \cite{dysthe2} is relatively brief in their verification of the BFI condition.
+The work in \cite{janssen} shows that the BFI condition does not yield a clear bifurcation in stability.
+Our results likewise support this.
+However, also shown in \cite{janssen} and followed up on in several later works; see \cite{onorato,slunyaev,eeltink}, it is shown that the kurtosis of the corresponding distribution of the surface waves, say $\tilde{k}(\tau)$ is a quadratic function of the time evolving $\mbox{BFI}(\tau)$.
+We likewise aim to establish a similar relationship, thereby maintaining the utility of the BFI insofar as it then helps predict a key feature of nonlinear-random wave fields.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results}
+To understand the relationship between the vorticity $\omega$ and the initial spectral width $\sigma$ of our initial conditions, we generate ensembles of initial conditions and then run these different initial conditions up to $\tau_{f}$ where $\tau_{f}$ is given by
+\[
+\tau_{f}=\max\left\{1.5/\alpha_{nl},1\right\},
+\]
+so that MI has a sufficiently long enough time scale over which to act; see \cite{curtis8} for a derivation of this result.
+We choose $L=5\pi$ so that the computational domain is $[-5\pi,5\pi]$.
+The number of modes is $K_{T} = 1024$ or $K=512$.
 The surface tension is $\tilde{\sigma} = 10^{-5}$.
-We then plot the averaged spectral density $S(k,t_{f};\sigma)$ where
+We then compute the averaged spectral density $S(k,\tau;\sigma)$ where
 \[
-S\left(k,t_{f};\sigma\right) = \overline{\left( \left|\frac{1}{\epsilon}\hat{\eta}_{1}\left(\frac{k-k_{0}}{\epsilon},1;\sigma \right)\right|^{2}\right)}.
+S\left(k,\tau;\sigma\right) \sim \overline{\left( \left|\hat{\eta}_{1}\left(k,\tau;\sigma \right)\right|^{2}\right)},
 \]
 which corresponds to the autocorrelation of the surface profile with respect to space.
-Note we have emphasized the dependence on the initial spectral width $\sigma$ for the sake of clarity.
-Following the approach in \cite{dysthe2,dysthe3}, by comparing the initial spectral density to that at later times, we can see directly whether MI manifests as significant distortions of the statistics, or if through supression the distributions remain relatively close to one another.
+Note, we normalize $S$ so that it is a probability distribution.
+Using $S(k,\tau_{f},\sigma)$, we then plot the time dynamics of the mean $m(\tau)$
+\[
+m(\tau) = \int_{\mathbb{R}} k S\left(k,\tau;\sigma\right) dk,
+\]
+standard deviation $\tilde{\sigma}(\tau)$,
+\[
+\tilde{\sigma}(\tau) = \left(\int_{\mathbb{R}} (k-m(\tau))^{2} S\left(k,\tau;\sigma\right) dk\right)^{1/2},
+\]
+and the kurtosis $\tilde{k}(\tau)$
+\[
+\tilde{k}(\tau) = \frac{1}{3\tilde{\sigma}^{4}(\tau)}\int_{\mathbb{R}} (k-m(\tau))^{4} S\left(k,\tau;\sigma\right) dk - 1,
+\]
+so as to examine the impact of various shear profiles on the statistics of the waves as well as establishing that statistical equilibrium has been achieved by $\tau_{f}$.
+Note, we interpret the kurtosis $\tilde{k}$ as the degree to which the distribution $S$ exhibits `heavy-tail phenomena', or the enhancement of the probability of what should otherwise be rare events.
+Given that all Gaussian distributions have $\tilde{k}=0$, all measurements are then made relative to this fixed level.
+Moreover, following the approach in \cite{dysthe2,dysthe3}, by comparing the initial spectral density to that at later times, we can see directly whether MI manifests as significant distortions of the statistics, or if through suppression the distributions remain relatively close to one another.
 
-Throughout our results, we have used a psuedo-spectral method in space coupled with a 4th-order Runge-Kutta scheme using integrating factors in time, with a time step of $dt = 2.5 \times 10^{-3}$.
+Throughout our results, we have used a pseudo-spectral method in space coupled with a 4th-order Runge-Kutta scheme using integrating factors in time, with a time step of $dt = 1 \times 10^{-3}$.
 De-aliasing is implemented via Orszag's `2/3-rule'.
-Averaging was done with $N_{e}=3200$ ensemble members, thereby ensuring relative averaging errors on the order of about $1\%$.
-Throughout this section, we present comparisons of the distributions for particular choices of $\sigma$ while also comparing the mean and standard deviations of the respective distributions so as to get a more quantitative understanding of the differences between the different distributions.
+Averaging was done with $N_{e}=512$ ensemble members, thereby ensuring relative averaging errors on the order of about $1\%$.
+We set $\epsilon=.05$ throughout the remainder of the paper.
 
-\subsection*{Strong-Counter Shear: $\omega = -1$}
+\subsection*{Vorticity-Free Flow: $\omega = 0$}
 
-Motivated by the results in \cite{curtis8}, we anticipate that $\omega=-1$ presents the most case which would allow for the vorticity to have the strongest impact on the BFI condition, which for the parameter choices requires that $\sigma > 11.9$.
-This is borne out by numerical experiment as seen in Figures \ref{fig:meanstdomn1} and \ref{fig:pspecomn1}.
-In this case, we see that for the NLSE the BFI condition largely holds.
-In particular, we see that while the mean never changes in any significant way from that of the initial distribution, increasing the width $\sigma$ causes the standard deviation of the distribution determined via evolution of the NLSE to collapse onto that of the initial, thereby showing that wider spreads in initial conditions are overall less susceptible to MI.
+We first examine the vorticity free case so as to establish a base line against which to measure non-zero shear.
+Looking at the time dynamics of the affiliated statistics when the initial width $\sigma=1.05\sigma_{c}$ where $\sigma_{c}\approx 4$, we see in Figure \ref{fig:stbleom0} the final steady distribution $S$ for $\tau_{f}=1$ as well as the affiliated time dynamics of the three primary moments $m(\tau)$, $\tilde{\sigma}(\tau)$, and $\tilde{k}(\tau)$.
+As can be seen, any claim of stability of the distribution for either the NLSE or VDE is tenuous at best.
+Note, the timescale for MI is $1/\alpha_{nl} = .5$ in this case.
+We see this result borne out in the plots in Figures \ref{fig:stbleom0} (b)-(d), which shows that saturation of the affiliated statistics does not set in till somewhat after this MI timescale.
+Thus, arguing that instability with respect to MI drives the observed statistical dynamics is reasonable.
+Comparing the VDE to the NLSE, we see the added nonlinearity in the VDE allows for yet greater deviations away from normality through the downshifting of the spectral peak; see Figure \ref{fig:stbleom0} (a), and the asymmetric widening of the VDE profile as evidenced by the enhanced kurtosis and strong shift in the mean as seen in Figures \ref{fig:stbleom0} (d) and (b) respectively.
+
+In contrast, by more than doubling the initial spectral width to $\sigma = 2\sqrt{2}\sigma_{c}$, we see in Figure \ref{fig:stbleom0wide} that an argument for stability against MI can be made.
+As can be seen for both the NLSE and VDE, all of the affiliated statistics remain close to their initial values and saturate rapidly, though we again see time of onset of saturation corresponds to the MI timescale.
+But as predicted in \cite{alber}, there should be profiles wide enough such that the impact of MI is nominal, though it is, as noted in \cite{janssen}, difficult to argue that what is being observed is a bifurcation as opposed to a weak nonlinear effect.
+Despite the relative stability, we again note that the VDE still has significant impacts on the statistics of the spectral distribution $S$, with in particular the VDE providing a slight enhancement of the kurtosis relative to the NLSE, which essentially maintains the normality of the initial distribution.
+This effect is also seen in the absence of an inversion in the magnitude of $\tilde{\sigma}(\tau)$, i.e.
+compare Figures \ref{fig:stbleom0} (b) to \ref{fig:stbleom0wide} (b), which shows that in the wider initial spectral width case, the VDE is wider than the NLSE, whereas in the more narrow initial spectral width case the roles between the two equations are reversed, reflecting the VDE's tendency to strongly skew $S(k,\tau_{f},\sigma)$.
 
@@
 \centering
-\begin{tabular}{c}
-\includegraphics[width=.85\textwidth]{omega_n1_ep0pt1_dt5en4}\\
-(a) $\epsilon=.1$ \\
-\includegraphics[width=.85\textwidth]{omega_n1_ep0pt04}\\
-(b) $\epsilon=.04$
+\begin{tabular}{cc}
+\includegraphics[width=60mm, height=39mm]{pdf_w_0_ep_pt05_Nens_512_sig0_1pt05} & \includegraphics[width=.48\textwidth]{mean_w_0_ep_pt05_Nens_512_sig0_1pt05} \\
+(a) $S(k,\tau_{f},1.05\sigma_{c})$ & (b) $m(\tau)$, $\sigma = 1.05\sigma_{c}$\\
+\includegraphics[width=.48\textwidth]{std_w_0_ep_pt05_Nens_512_sig0_1pt05} & \includegraphics[width=.48\textwidth]{kts_w_0_ep_pt05_Nens_512_sig0_1pt05} \\
+(c) $\tilde{\sigma}(\tau)$, $\sigma = 1.05\sigma_{c}$ & (d) $\tilde{k}(\tau)$, $\sigma = 1.05\sigma_{c}$\\
 \end{tabular}
-\caption{For $\omega=-1$, comparison of the means and standard deviations between the initial averaged spectral density and those generated by the NLSE and VDE.
-This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).
-The vertical line denotes the spectral width corresponding to the BFI condition.}
-\label{fig:meanstdomn1}
+\caption{For $\omega=0$ and $\sigma=1.05 \sigma_{c}$, in (a) the plot of the spectral distribution function $S(k,\tau_{f},1.05\sigma_{c})$, in (b) the plot of the mean $m(\tau)$, in (c) the plot of the standard deviation $\tilde{\sigma}(\tau)$, and in (d) the plot of the kurtosis $\tilde{k}(\tau)$.
+In the graphs, the NLSE is (- -), the VDE (-), and the initial distribution is ($\cdots$).}
+\label{fig:stbleom0}
 \end{figure}
@@
 \centering
-\begin{tabular}{c}
-\includegraphics[width=.65\textwidth]{omega_n1_ep0pt1_dt5en4_width12} \\
-(a) $\epsilon=.1$\\
-\includegraphics[width=.65\textwidth]{omega_n1_ep0pt04_width12}\\
-(b) $\epsilon=.04$ 
+\begin{tabular}{cc}
+\includegraphics[width=.48\textwidth]{pdf_w_0_ep_pt05_Nens_512_sig0_2sqrt2} & \includegraphics[width=.45\textwidth]{mean_w_0_ep_pt05_Nens_512_sig0_2sqrt2} \\
+(a) $S(k,\tau_{f},2\sqrt{2}\sigma_{c})$ & (b) $m(\tau)$, $\sigma = 2\sqrt{2}\sigma_{c}$\\
+\includegraphics[width=.48\textwidth]{std_w_0_ep_pt05_Nens_512_sig0_2sqrt2} & \includegraphics[width=.48\textwidth]{kts_w_0_ep_pt05_Nens_512_sig0_2sqrt2} \\
+(c) $\tilde{\sigma}(\tau)$, $\sigma = 2\sqrt{2}\sigma_{c}$ & (d) $\tilde{k}(\tau)$, $\sigma = 2\sqrt{2}\sigma_{c}$
 \end{tabular}
-\caption{For $\omega=-1$, comparison of the averaged spectral densities between the initial averaged spectral density and those generated by the NLSE and VDE.
-This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).}
-\label{fig:pspecomn1}
+\caption{For $\omega=0$ and $\sigma=2\sqrt{2} \sigma_{c}$, in (a) the plot of the spectral distribution function $S(k,\tau_{f},1.05\sigma_{c})$, in (b) the plot of the mean $m(\tau)$, in (c) the plot of the standard deviation $\tilde{\sigma}(\tau)$, and in (d) the plot of the kurtosis $\tilde{k}(\tau)$.
+In the graphs, the NLSE is (- -), the VDE (-), and the initial distribution is ($\cdots$).}
+\label{fig:stbleom0wide}
 \end{figure}
 
-However, the results for the VDE are markedly different.
-As seen in Figure \ref{fig:pspecomn1}, where $\sigma=12$ so that we are just past the BFI threshold, the most peculiar feature of the VDE is the bimodality of the distribution.
-There is clearly a frequency downshift of the central initial peak, but a secondary peak at high-wave number forms as well.
-This formation is less pronounced with decreasing $\epsilon$, but is clearly there nevertheless.
-This represents a kind of resonance condition for which no ready explanation exists.
-This feature is likewise responsible for the strong skewing of the means and standard deviations seen in Figure \ref{fig:meanstdomn1}.
+\subsection*{Counter-Propagating Shear: $\omega = -.5$}
 
-\subsection*{Vorticity-Free Flow: $\omega = 0$}
+For the first case of nontrivial shear, we look at setting $\omega=-.5$ which corresponds to the shear profile propagating against the movement of the surface wave.
+This parameter choice gives us the BFI $\sigma_{c} \approx 5.5$.
+Here $\tau_{f}=1$ since the MI timescale is given by $1/\alpha_{nl}\approx .3$.
+As seen in Figure \ref{fig:stbleomnh}, the BFI condition does not appear to hold, with significant deviations away from the initial Gaussian profile happening on the predicted MI timescale.
+Moreover, by again increasing the initial width to $\sigma = 2\sqrt{2}\sigma_{c}$, the overall impact of MI is markedly reduced, with the VDE and NLSE all but producing nearly identical results as seen in Figure \ref{fig:stbleomnhwide}.
+We likewise again see the inversion in the standard deviation between the two models for the narrow initial spectral width; see Figure \ref{fig:stbleomnh} (b), while the inversion is absent in the wide initial width case; see Figure \ref{fig:stbleomnhwide}.
 
-Removing vorticity, we see that we essentially recreate the standard BFI, with BFI threshold $\sigma > 5.7$, results for the NLSE.
-Again though, the VDE produces markedly different statistical results from the NLS, though to nowhere near as drastic an extent as when $\omega=-1$.
-In particular, the bimodality and strong deformation of the profile is notably ameliorated for $\epsilon=.04$; see Figures \ref{fig:meanstdom0} and \ref{fig:pspecom0}.
-Thus in some sense, though not entirely, MI has more or less similar characteristics for both the NLSE and VDE.
+That being said, while the VDE still produces a final distribution which is less Gaussian and more favorable to far tail phenomena relative to the NLSE, we also see that the counter propagating shear lowers the overall kurtosis compared to the case of $\omega=0$.
+Thus while the overall distributions are wider in the $\omega=-.5$ case relative to the $\omega=0$ case, the drop in kurtosis induced by the counter-propagating shear would seem to still reduce the relative importance of far-tail phenomena.
+We can then make a prediction that counter-propagating shear would generally reduce the chances of events like rogue-waves forming.
 
@@
 \centering
-\begin{tabular}{c}
-\includegraphics[width=.85\textwidth]{omega0_ep0pt1}\\
-(a) $\epsilon=.1$ \\
-\includegraphics[width=.85\textwidth]{omega0_ep0pt04}\\
-(b) $\epsilon=.04$
+\begin{tabular}{cc}
+\includegraphics[width=60mm, height=39mm]{pdf_w_nhf_ep_pt05_Nens_512_sig0_1pt05} & \includegraphics[width=.48\textwidth]{mean_w_nhf_ep_pt05_Nens_512_sig0_1pt05} \\
+(a) $S(k,\tau_{f},1.05\sigma_{c})$ & (b) $m(\tau)$, $\sigma = 1.05\sigma_{c}$\\
+\includegraphics[width=.48\textwidth]{std_w_nhf_ep_pt05_Nens_512_sig0_1pt05} & \includegraphics[width=.48\textwidth]{kts_w_nhf_ep_pt05_Nens_512_sig0_1pt05} \\
+(c) $\tilde{\sigma}(\tau)$, $\sigma = 1.05\sigma_{c}$ & (d) $\tilde{k}(\tau)$, $\sigma = 1.05\sigma_{c}$\\
 \end{tabular}
-\caption{For $\omega=0$, comparison of the means and standard deviations between the initial averaged spectral density and those generated by the NLSE and VDE.
-This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).
-The vertical line denotes the spectral width corresponding to the BFI condition.}
-\label{fig:meanstdom0}
+\caption{For $\omega=-.5$ and $\sigma=1.05 \sigma_{c}$, in (a) the plot of the spectral distribution function $S(k,\tau_{f},1.05\sigma_{c})$, in (b) the plot of the mean $m(\tau)$, in (c) the plot of the standard deviation $\tilde{\sigma}(\tau)$, and in (d) the plot of the kurtosis $\tilde{k}(\tau)$.
+In the graphs, the NLSE is (- -), the VDE (-), and the initial distribution is ($\cdots$).}
+\label{fig:stbleomnh}
 \end{figure}
@@
 \centering
-\begin{tabular}{c}
-\includegraphics[width=.65\textwidth]{omega0_ep0pt1_width6} \\
-(a) $\epsilon=.1$\\
-\includegraphics[width=.65\textwidth]{omega0_ep0pt04_width6}\\
-(b) $\epsilon=.04$ 
+\begin{tabular}{cc}
+\includegraphics[width=60mm, height=39mm]{pdf_w_nhf_ep_pt05_Nens_512_sig0_2sqrt2} & \includegraphics[width=.45\textwidth]{mean_w_nhf_ep_pt05_Nens_512_sig0_2sqrt2} \\
+(a) $S(k,\tau_{f},2\sqrt{2}\sigma_{c})$ & (b) $m(\tau)$, $\sigma = 2\sqrt{2}\sigma_{c}$\\
+\includegraphics[width=.48\textwidth]{std_w_nhf_ep_pt05_Nens_512_sig0_2sqrt2} & \includegraphics[width=.48\textwidth]{kts_w_nhf_ep_pt05_Nens_512_sig0_2sqrt2} \\
+(c) $\tilde{\sigma}(\tau)$, $\sigma = 2\sqrt{2}\sigma_{c}$ & (d) $\tilde{k}(\tau)$, $\sigma = 2\sqrt{2}\sigma_{c}$
 \end{tabular}
-\caption{For $\omega=0$, comparison of the averaged spectral densities between the initial averaged spectral density and those generated by the NLSE and VDE.
-This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).}
-\label{fig:pspecom0}
+\caption{For $\omega=-.5$ and $\sigma=2\sqrt{2} \sigma_{c}$, in (a) the plot of the spectral distribution function $S(k,\tau_{f},1.05\sigma_{c})$, in (b) the plot of the mean $m(\tau)$, in (c) the plot of the standard deviation $\tilde{\sigma}(\tau)$, and in (d) the plot of the kurtosis $\tilde{k}(\tau)$.
+In the graphs, the NLSE is (- -), the VDE (-), and the initial distribution is ($\cdots$).}
+\label{fig:stbleomnhwide}
 \end{figure}
 
-\subsection*{Strong-Co Shear: $\omega = 1$}
+\subsection*{Co-Propagating Shear: $\omega = 1.12$}
 
-Finally, by letting $\omega=1$, we further reduce the BFI threshold to $\sigma > 2.7$, opening up a yet wider range of stable initial conditions.
-Likewise, the VDE dynamics are far more in line with those seen from the NLSE, even in the case that $\epsilon=.1$; see Figures \ref{fig:meanstdom1} and \ref{fig:pspecom1}.
-Therefore, we see how vorticity propagating in the correct direction can vastly reduce the impact of MI, even in a far more nonlinear model like the VDE.
+By letting $\omega=1.12$, corresponding to a shear profile propagating in the direction of the surface wave, we reduce the BFI threshold to $\sigma_{c} \approx 1.01$.
+This case becomes the most interesting due to the results in \cite{curtis8} which show that for $k_{0}=1$ that $\omega=1.12$ corresponds to the shear value which most suppresses Stokes drift, i.e.
+the mean flow of Lagrangian tracers in the fluid.
+For this case, the MI timescale is given by $1/\alpha_{nl}\approx 11.8$, so $\tau_{f}\approx 18$.
+This far longer time scale is seen in Figure \ref{fig:stbleom1}, where the saturation of the statistics happens just before $\tau_{f}$.
+For the narrow initial spectral width $\sigma = 1.05\sigma_{c}$, we see some of the most pronounced deviations between the VDE and NLSE.
+As before, there is the inversion in the deviation, see Figure \ref{fig:stbleom1} (c), with the VDE tending to sharpen around the mean.
+In this case though, the downshifting in the peak and corresponding one sided skewing of the distribution, see Figure \ref{fig:stbleom1} (a) with its concomitant rise in the kurtosis, \ref{fig:stbleom1} (d), is most dramatic for the $\omega=1.12$ case.
+Thus the VDE markedly enhances far-tail phenomena and provides strikingly different predictions than its counterpart model the NLSE.
+Thus we see that the the co-propagating shear likely enhances the occurrence of freak wave or other far-tail phenomena for relatively narrow spectral profiles
 
+Of course, as before, when $\sigma=2\sqrt{2}\sigma_{c}$, the added initial spectral width appears to mitigate the impacts of any instabilities or nonlinearity; see Figure \ref{fig:stbleom1wide}.
+An interesting point though is that while relatively stable, the widths of the profile for the $\omega=1.12$ case are not as wide as in the previous cases.
+This shows that the changed physics coming with the different shear profiles do change the spectral width saturation threshold, and therefore while the BFI condition from \cite{alber} is perhaps incorrect, there clearly is a parameter dependent mechanism which controls the degree to which a given random-wave field achieves a steady statistical state.
 \begin{figure}[!ht]
 \centering
-\begin{tabular}{c}
-\includegraphics[width=.85\textwidth]{omega1_ep0pt1}\\
-(a) $\epsilon=.1$ \\
-\includegraphics[width=.85\textwidth]{omega1_ep0pt04}\\
-(b) $\epsilon=.04$
+\begin{tabular}{cc}
+\includegraphics[width=60mm, height=39mm]{pdf_w_1pt12_ep_pt05_Nens_512_sig0_1pt05} & \includegraphics[width=.48\textwidth]{mean_w_1pt12_ep_pt05_Nens_512_sig0_1pt05} \\
+(a) $S(k,\tau_{f},1.05\sigma_{c})$ & (b) $m(\tau)$, $\sigma = 1.05\sigma_{c}$\\
+\includegraphics[width=.48\textwidth]{std_w_1pt12_ep_pt05_Nens_512_sig0_1pt05} & \includegraphics[width=.48\textwidth]{kts_w_1pt12_ep_pt05_Nens_512_sig0_1pt05} \\
+(c) $\tilde{\sigma}(\tau)$, $\sigma = 1.05\sigma_{c}$ & (d) $\tilde{k}(\tau)$, $\sigma = 1.05\sigma_{c}$\\
 \end{tabular}
-\caption{For $\omega=1$, comparison of the means and standard deviations between the initial averaged spectral density and those generated by the NLSE and VDE.
-This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).
-The vertical line denotes the spectral width corresponding to the BFI condition.}
-\label{fig:meanstdom1}
+\caption{For $\omega=1.12$ and $\sigma=1.05 \sigma_{c}$, in (a) the plot of the spectral distribution function $S(k,\tau_{f},1.05\sigma_{c})$, in (b) the plot of the mean $m(\tau)$, in (c) the plot of the standard deviation $\tilde{\sigma}(\tau)$, and in (d) the plot of the kurtosis $\tilde{k}(\tau)$.
+In the graphs, the NLSE is (- -), the VDE (-), and the initial distribution is ($\cdots$).}
+\label{fig:stbleom1}
 \end{figure}
@@
 \centering
-\begin{tabular}{c}
-\includegraphics[width=.65\textwidth]{omega1_ep0pt1_width3} \\
-(a) $\epsilon=.1$\\
-\includegraphics[width=.65\textwidth]{omega1_ep0pt04_width3}\\
-(b) $\epsilon=.04$ 
+\begin{tabular}{cc}
+\includegraphics[width=.48\textwidth]{pdf_w_1pt12_ep_pt05_Nens_512_sig0_2sqrt2} & \includegraphics[width=.48\textwidth]{mean_w_1pt12_ep_pt05_Nens_512_sig0_2sqrt2} \\
+(a) $S(k,\tau_{f},2\sqrt{2}\sigma_{c})$ & (b) $m(\tau)$, $\sigma = 2\sqrt{2}\sigma_{c}$\\
+\includegraphics[width=.48\textwidth]{std_w_1pt12_ep_pt05_Nens_512_sig0_2sqrt2} & \includegraphics[width=.48\textwidth]{kts_w_1pt12_ep_pt05_Nens_512_sig0_2sqrt2} \\
+(c) $\tilde{\sigma}(\tau)$, $\sigma = 2\sqrt{2}\sigma_{c}$ & (d) $\tilde{k}(\tau)$, $\sigma = 2\sqrt{2}\sigma_{c}$
 \end{tabular}
-\caption{For $\omega=1$, comparison of the averaged spectral densities between the initial averaged spectral density and those generated by the NLSE and VDE.
-This is shown for varying strengths of the small amplitude, and thus nonlinearity, parameter $\epsilon$ with $\epsilon=.1$ in (a) and $\epsilon=.04$ in (b).}
-\label{fig:pspecom1}
+\caption{For $\omega=1.12$ and $\sigma=2\sqrt{2} \sigma_{c}$, in (a) the plot of the spectral distribution function $S(k,\tau_{f},1.05\sigma_{c})$, in (b) the plot of the mean $m(\tau)$, in (c) the plot of the standard deviation $\tilde{\sigma}(\tau)$, and in (d) the plot of the kurtosis $\tilde{k}(\tau)$.
+In the graphs, the NLSE is (- -), the VDE (-), and the initial distribution is ($\cdots$).}
+\label{fig:stbleom1wide}
 \end{figure}
 
-\section*{Conclusion}
+\subsection*{Kurtosis and the BFI}
+Comparing across the plots in Figure \ref{fig:kurtplt}, we summarize the impact that varying BFI indices, or varying initial spectral widths $\sigma$ for $1.05\sigma_{c}\leq \sigma \leq 2\sqrt{2}\sigma_{c}$, have on the kurtosis of the final spectral distributions.
+As expected from the plots above, an increasing BFI, or decreasing initial spectral width, corresponds to a stronger kurtosis and thus a stronger deviation away from normality of the final spectral distribution.
+Moreover, increasing shear enhances the kurtosis, and finally the higher-order nonlinearities of the VDE always enhance the kurtosis relative to the value predicted by the NLSE.
+Likewise, we see that the quadratic fits are clearly good, though also it is clear varying vorticity values change the coefficients of the fit.
+This conforms with the results in \cite{eeltink} which show that the quadratic relationship between the kurtosis and the BFI varies depending on different damping and wind profiles.
+Thus, varying physics has a significant impact on far-tail phenomena manifesting in real-world situations.
+\begin{figure}[!ht]
+\centering
+\begin{tabular}{c}
+\includegraphics[width=.8\textwidth]{kurtosis_fit_om_0_Nens_128}\\
+(a) $\omega = 0$ \\
+\includegraphics[width=.8\textwidth]{kurtosis_fit_om_nh_Nens_128}\\
+(b) $\omega = -.5$\\
+\includegraphics[width=.8\textwidth]{kurtosis_fit_om_1pt12_Nens_128} \\
+(c) $\omega=1.12$  
+\end{tabular}
+\caption{Fixing $\epsilon=.05$, the kurtosis as a function of the BFI for $1.05\sigma_{c}\leq \sigma\leq 2\sqrt{2}\sigma_{c}$ and the corresponding quadratic fits for $\omega=0$ (a), $\omega=-.5$ (b), and $\omega=1.12$ (c).}
+\label{fig:kurtplt}
+\end{figure}
+
+\section{Conclusion}
+
+In this work, we have studied the impact of modulational stability and other nonlinear effects on the statistical properties of deep-water-surface wave flows moving over constant-vorticity-shear currents.
+As we see, the vorticity profile has a strong influence on the stability properties and far tails of the wave train, with co-propagating shears exacerbating modulational instability and higher order nonlinear effects which produce slower decaying tales thereby increasing the likelihood of rare events.
+In contrast, counter-propagating shears reduce the impact of modulational instability and higher order nonlinear effects and in turn reduce the likelihood of rare events.
+Thus, we see vorticity effects can have significant implications for what the most likely observed deep water wavetrains and phenomena are.
+Likewise, in trying to interpret oceanographic data, vortical effects should be taken into account in modeling so as to be able to properly interpret observed data.
+
+\section*{Acknowledgements}
+The authors would like to thank the support of the NSF through DMS-1715039.
+Likewise, we appreciate the many conversations and constructive suggestions of John Carter.
+On behalf of all authors, the corresponding author states that there is no conflict of interest.
+
+\pagebreak
 \bibliography{deep_water_shear_Feb2}

### 2020-01-09 Writeup/NLS_Paths_Master_File_July30.tex
All the updates

\documentclass{JFM_Style/jfm}
\usepackage{epstopdf, epsfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}

\newcommand{\ba}{\begin{array}}
\newcommand{\ea}{\end{array}}

\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\newcommand{\bc}{\begin{center}}
\newcommand{\ec}{\end{center}}

\newcommand{\ds}{\displaystyle}

\newcommand{\bt}{\begin{tabular}}
\newcommand{\et}{\end{tabular}}

\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}

\newcommand{\bd}{\begin{description}}
\newcommand{\ed}{\end{description}}

\newcommand{\bp}{\begin{pmatrix}}
\newcommand{\ep}{\end{pmatrix}}

\newcommand{\pd}{\partial}
\newcommand{\sech}{\mbox{sech}}

\newcommand{\cf}{{\it cf.}~}

\newcommand{\ltwo}{L_{2}(\mathbb{R}^{2})}
\newcommand{\smooth}{C^{\infty}_{0}(\mathbb{R}^{2})}

\newcommand{\br}{{\bf r}}
\newcommand{\bk}{{\bf k}}
\newcommand{\bv}{{\bf v}}

\newcommand{\gnorm}[1]{\left|\left| #1\right|\right|}
\newcommand{\ipro}[2]{\left<#1,#2 \right>}

\title{Particle Paths in Nonlinear Schr\"{o}dinger Models in the Presence of Linear Shear Currents}
\shortauthor{C.W. Curtis and J.D. Carter and H. Kalisch}

\date{}

\author{C.W. Curtis \aff{1}
  \corresp{\email{ccurtis@mail.sdsu.edu}}, J.D. Carter\aff{2} \and H. Kalisch\aff{3}}

\affiliation{\aff{1}Department of Mathematics and Statistics, San Diego State University, San Diego, CA 92182
\aff{2}Mathematics Department, Seattle University, Seattle, WA 98122
\aff{3}Department of Mathematics, University of Bergen, Bergen, Norway
}

\begin{document}
\maketitle

\begin{abstract}
We investigate the effect of constant vorticity background shear on the properties of wavetrains in deep water.  Using the methodology of \cite{fokas2008}, we derive a higher-order nonlinear Schr\"{o}dinger equation in the presence of shear and surface tension.  We show that the presence of shear induces a strong coupling between the carrier wave and the mean surface displacement.  The effects of the background shear on the modulational instability of plane waves is also studied, where it is shown that shear can suppress instability, though not for all carrier wavelengths in the presence of surface tension.  These results expand upon the findings of \cite{thomas2012nonlinear}.

Using a modification of the Generalized Lagrangian Mean theory in  \cite{andrews} and approximate formulas for the velocity field in the fluid column, explicit, asymptotic approximations for the Lagrangian and Stokes drift velocities are obtained for plane-wave and Jacobi elliptic function solutions of the nonlinear Schr\"odinger equation.  Numerical approximations to particle trajectories for these solutions are found and the Lagrangian and Stokes drift velocities corresponding to these numerical solutions corroborate the theoretical results.

We show that background currents have significant effects on the mean transport properties of waves. In particular, certain combinations of background shear and carrier wave frequency lead to the disappearance of mean surface mass transport.  These results provide a possible explanation for the measurements reported in \cite{smith}.  Our results also provide further evidence of the viability of the modification of the Stokes drift velocity beyond the standard monochromatic approximation, such as recently proposed in \cite{breivik} in order to obtain a closer match to a range of complex ocean wave spectra.
\end{abstract}

\section{Introduction}

Currents are an ubiquitous feature in oceanic dynamics where they are a driving force in the formation and propagation of waves in the ocean.  \cite{helfrich} show that shallow water currents over bathymetric variations are a key mechanism for surface and internal wave generation.  Slowly varying currents likewise act as refractive medium for small-amplitude, linear surface and internal water waves, thereby strongly influencing propagation and dispersion; see \cite{mcwilliams,buhler,young}.  Concomitant with this, currents also strongly influence the mean transport properties of waves.  This issue has been extensively studied in \cite{craik1,craik2,craik3,craik4,phillips1,phillips2} where the question of how shear currents excite instabilities and drive mean wave transport in the linear limit has been thoroughly addressed.  And in the absence of depth varying currents, \cite{ardhuin} show that irrotational nonlinear Miche waves, see \cite{miche}, can induce strong differences from linear approximations to mean properties of the flow.  However, what is less well understood is the interplay of depth-varying currents with nonlinear surface waves, and how nonlinearity influences the mean transport properties of waves when currents are present.

By restricting to the case of a constant-vorticity-shear current, given its relative simplicity, detailed explorations of the existence, shape, stability, and pointwise properties of nonlinear waves in the presence of a current has been studied in a wide number of places such as \cite{freeman,brevik,simmen,pullin2,pullin3,dasilva,baumstein,choi,wahlen,wahlen2,constantin,thomas2012nonlinear,vasan2014pressure} and \cite{nachbin} among others. In particular, in \cite{thomas2012nonlinear} the impact of vorticity on the modulational instabilities (MIs) of wave trains in deep water was examined.  The MI, which is also known as the Benjamin--Feir instability, is a key factor in the understanding of a number of wave phenomena, and is particularly important in the study of issues such as wave breaking and freak wave formation.  In \cite{thomas2012nonlinear} it is shown that constant vorticity strongly modifies the onset and bandwidth of MIs. Further, it is shown that some currents can even completely suppress MIs, which is a striking result and speaks to the importance of better understanding the role of shear currents in deep water flows.

But there is far less known about how currents impact the transport properties of nonlinear surface waves.  A key measure of the strength of the material transport of surface waves is the Lagrangian drift velocity (LDV), which in the absence of background currents is the mean speed at which fluid parcels travel in a flow, see \cite{longuet}.  An understanding of the LDV, and the associated mean quantity known as the Stokes drift velocity (SDV), is critical in the study of mass and energy transport in oceanic environments, and it has been a central point of investigation in a number of studies such as \cite{mcwilliams,webb,breivik}.  Regarding the impact of constant vorticity currents on the transport properties of surface waves in deep water, the experimental work of \cite{monismith} and the field measurements of \cite{smith} seem to suggest wavetrains on deep water excite Eulerian counter currents which in effect cancel the mean flow induced by the SDV.  These counter currents cancel surface drift in the open ocean such as explained by \cite{smith} and they can even cancel the drift at both the surface and throughout the bulk of fluids in laboratory settings such as found by \cite{monismith}.  To date, there seems to be no theoretical framework explaining the mechanisms behind the development of these currents.  A different but related dilemma emerges in attempts to fit classic, current free, derivations for the SDV in deep water to oceanographic data.  Recently, \cite{breivik} put forward phenomenological modifications to the SDV profile which amount to introducing background currents. They show that the use of these modified SDV profiles is superior to the standard approach of using monochromatic SDV profiles.  However, no physical mechanism is provided which explains the origins of the modifications used in \cite{breivik}.

In order then to better understand the interplay of constant vorticity on deep-water nonlinear-surface flows, we study how a constant vorticity shear profile influences the motion and mean properties of particle paths both at and beneath waves in infinitely deep water.  Our work complements and expands on the shallow and finite depth results found in \cite{wahlen2,constantin,borluk} and \cite{nachbin}.  To do this, we first derive a higher-order nonlinear Schr\"{o}dinger (NLS) model, which we call the Vor-Dysthe (VD) equation, which describes the long temporal and spatial coupling between the nonlinear carrier wave and the mean fluid depth, in effect extending the now classic results of \cite{dysthe} to include constant vorticity and extending the results of \cite{thomas2012nonlinear} to the next asymptotic order.  This is done via the methodology of Ablowitz--Fokas--Musslimani (AFM), see \cite{afm,ashton}, which is a particular case of the more general approach of the Unified Transform Method, see \cite{fokas2008}.  Using this derivation, we determine not only the appropriate form of the NLS equation, but also the mean-surface height and tangential-surface velocity. We show that the mean-surface height is markedly increased due to the background shear current.  We characterize the impact of constant vorticity and surface tension on MIs, in which we show that surface tension in general prevents there being any vorticity value which completely supresses instability.  This expands on the zero-surface tension results in \cite{thomas2012nonlinear}.  %We note that our derivation method is quite distinct from that in \cite{thomas2012nonlinear} and goes to a higher order.  This provides yet another example of the utility of the AFM formulation.  We then derive dynamical systems which describe the particle paths.

Using these derivations, we provide a description of the impact of constant vorticity shear currents on the LDV and SDV via the techniques of the Generalized Lagrangian Mean (GLM) theory as found in \cite{andrews,buhler}. This approach provides an unambiguous way of computing mean velocities under the influence of depth varying currents, whereby we are then able to explicitly derive formulas for the LDV and SDV for a variety of solutions to the NLS equation. Our method enables the quantification of the combined effect of background shear and wave modulation, and to determine which combinations of background shear and carrier wave frequency produce particularly strong or weak mean flows.  Further, we are able to show which background shear currents quench mean surface flows, thereby providing a potential theoretical explanation for the results presented in \cite{smith}. We also show how one can derive from our model the phenomenological modifications used in \cite{breivik}.

As stated above, results on the mean transport properties of far more general shear flows have appeared in \cite{craik2, craik3, phillips1, phillips2}, which also explore the existence of transverse instabilities to said shear profiles.  This was done even in the presence of free surfaces; see \cite{phillips1,phillips2}.  However, we note that throughout all of these works, aside from just examing the linear limit of the free surface, thereby disallowing for the study of the slow modulations described by the NLS and VD equations, the computations of mean velocities were done in such a way so that fluctuations around the mean in the horizontal and vertical directions were independent.  While this is certainly appropriate in the bullk of the fluid,  building off a comment in Section 4.2 of \cite{andrews}, and paralleling work in \cite{ardhuin}, one must study fluctuations at the surface taking surface constraints into account.  Thus in this paper, we clarify the role of the free surface in a more explicit manner within the GLM methodology.

Computations show that our treatement of the surface leads to different mean velocity results than one would get through direct use of the formulas found in say \cite{craik2}, though we note in that paper rigid lids were assumed.  Our predictions of the mean surface velocities are confirmed through the numerical experiments presented in this paper, in particular with regards to the correct dertimination of those constant vorticity values which quench mean transport.  We also note that our formulas for the LDV are in terms of the slow variables underlying the NLS equation, thereby allowing for a clear understanding of the impact of nonlinear wave modulation on the mean transport properties throughout the fluid.  Thus, in this regard, our formulas are nonlinear, though not necessarily of larger amplitude than those found in previous works.

As a way of confirming our theoretical findings, we discretize the dynamical system describing particle trajectories in the NLS approximation and compute approximations of such trajectories. Our numerical results on the impact of background shear currents on particle paths and the LDV support the results predicted by our theory. We generate the numerical approximations by using using numerical solutions to the VD equation with initial conditions corresponding to the Jacobi elliptic function solutions for both the focusing and defocusing NLS equations to model deep water time-evolving surface wavetrains.  Complementing these results, we also examine using initial conditions corresponding to the defocusing plane-wave solutions of the NLS equation.  Our results show how transport properties are most enhanced for shear currents which at the surface are directed against the carrier wave.  This can occur near what appear to be resonances between nonlinearity and the background shear profile.  Likewise, in the defocusing case we numerically confirm the theoretical predictions made for those balances between vorticity and carrier wave-number which one expects to quench the mean surface drift, thereby providing a possible explanatory mechanism for the results in \cite{smith,breivik,monismith}.  In total, our theoretical and numerical results show that the transport properties of surface waves can be greatly affected by constant vorticity background shear profiles, motivating further study of the impact of rapidly depth-varying currents on transport properties on nonlinear wavetrains at the free surface.

The outline of the paper is as follows.  The derivations of the VD and NLS equations and the dynamical system describing particle paths is given in Section 2.  The explanations and derivations of the formulas necessary to compute the Stokes drift velocity are given in Section 3.  The numerical results on the particle paths and Stokes drift velocity are in Section 4.  Finally, Section 5 contains our conclusions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Derivations}
\subsection{Derivation of NLS with Constant Vorticity in Infinite Depth}
We examine the unsteady nonlinear wave propagation over a constant shear current.  To do this, we assume the fluid velocity has the form
\[
{\bf u} = u({\bf x},t)\hat{{\bf x}} + w({\bf x},t)\hat{{\bf z}} = \omega z\hat{{\bf x}} + \nabla \phi,
\]
where $\phi$ is a harmonic function.  We restrict fluid motion to the $(x,z)$-plane, thereby ignoring transverse variations in the $y$ dimension.   Following standard arguments, e.g.~see \cite{ashton}, the dynamics of the fluid can be determined by solving the free boundary value problem
\begin{align}
\Delta \phi = 0, & & -\infty< z < \eta(x,t), \nonumber \\
\eta_{t} +\left(\omega\eta+\phi_{x}\right)\eta_{x} - \phi_{z} = 0, & & z=\eta(x,t), \label{kine}\\
\phi_{t} + \omega\pd_{x}^{-1}\eta_{t} + \frac{1}{2} \left|{\bf
    u}\right|^{2} + g\eta -
\frac{\sigma}{\rho}\pd_{x}\frac{\eta_{x}}{\sqrt{1+\eta_{x}^{2}}}= 0, &
& z=\eta(x,t), \label{bernou}\\
\lim_{z\rightarrow -\infty} \phi_{z} = 0, \nonumber
\end{align}
where $\eta$ represents the free surface displacement, and $g$, $\rho$, and $\sigma$ represent the acceleration due to gravity, the fluid density, and the coefficient of surface tension respectively.

By choosing a characteristic wave height $a$ and wave length $L$, all quantities can be non-dimensionalized via
\begin{align*}
\tilde{x} = \frac{x}{L}, ~\tilde{z} = \frac{z}{L}, ~ \tilde{t} = \sqrt{\frac{g}{L}}~t, ~\omega = \sqrt{\frac{g}{L}}~\tilde{\omega}, \\
\eta = a \tilde{\eta}, ~ \phi  = a\sqrt{gL}~\tilde{\phi} , ~ \epsilon = \frac{a}{L}.
\end{align*}
By following the AFM approach described in \cite{afm,ashton} and dropping the tildes, the kinematic boundary condition, Equation \eqref{kine}, can be written in terms of surface variables alone via the integro-differential equation
\begin{equation}
\int_{\mathbb{R}}dx~e^{-ikx}e^{\epsilon|k|\eta}\left(\eta_{t} +
  \epsilon \omega \eta \eta_{x} + i \mbox{sgn}(k)Q \right) = 0, ~
k\neq 0,
\label{kine2}
\end{equation}
where $Q=q_{x}$, $q(x,t) = \phi(x,\eta(x,t),t)$.  By integrating over $\mathbb{R}$, we are assuming that both $\eta$ and $Q$ decay to zero sufficiently rapidly in the far field.  We can also readily derive a nearly identical expression on domains periodic in the horizontal variable $x$, in which case Equation \eqref{kine2} becomes
\begin{equation}
\int_{-L_{p}/2}^{L_{p}/2}dx~e^{-ikx}e^{\epsilon|k|\eta}\left(\eta_{t} +
  \epsilon \omega \eta \eta_{x} + i \mbox{sgn}(k)Q \right) = 0, ~ k = \frac{2\pi m}{L_{p}}, ~ m\in \mathbb{Z}\backslash{0},
\label{integro1per}
\end{equation}
where $L_p$ is the spatial period.  Throughout the remainder of this section we only present results over the real line, $\mathbb{R}$, since identical results can be derived for the periodic case.
Lastly, we note that if in Equation \eqref{kine2} we approach $k=0$ from both the left and the right, we get the equations
\[
\int_{\mathbb{R}}dx \left(\eta_{t} + i Q \right) = 0, ~ \int_{\mathbb{R}}dx \left(\eta_{t} - i Q \right) = 0.
\]
Thus we get the identities
\begin{align}
\p_{t}\int_{\mathbb{R}} dx ~\eta(x,t) & = 0, \label{zavsurf}\\
\int_{\mathbb{R}} dx ~Q(x,t) & = 0. \label{zavpot}
\end{align}

Taylor expanding Equation (\ref{kine2}) up to $\mathcal{O}(\epsilon^3)$ for $k\neq 0$ gives
\begin{multline}
\int_{\mathbb{R}}dx~e^{-ikx}\left(1 + \epsilon |k|\eta + \frac{\epsilon^{2}|k|^{2}\eta^{2}}{2} + \frac{\epsilon^{3}|k|^{3}\eta^{3}}{6}\right)\left(\eta_{t} + i \mbox{sgn}(k)Q \right) \\
+ \epsilon \omega \int_{\mathbb{R}}dx~e^{-ikx}\left(1 + \epsilon |k|\eta + \frac{\epsilon^{2}|k|^{2}\eta^{2}}{2}\right) \eta \eta_{x} = 0.
\label{integro1}
\end{multline}
Transforming into surface variables and Taylor expanding Equation \eqref{bernou}, Bernoulli's equation, up to $\mathcal{O}(\epsilon^3)$ gives
\begin{multline}
Q_{t} + \omega \eta_{t} + \eta_{x} - \tilde{\sigma}\eta_{xxx} + \frac{\epsilon}{2}\pd_{x}\left(-\eta_{t}^{2} + (Q+\omega \eta)^{2} \right)\\
+ \epsilon^{2}\pd_{x}\left( \frac{3}{2}\tilde{\sigma}\eta_{x}^{2}\eta_{xx} - \eta_{t}\eta_{x} \left(Q  + \omega \eta \right) \right) - \frac{\epsilon^{3}}{2}\eta_{x}^{2}\left((\omega\eta+Q)^{2}-\eta_{t}^{2}\right) = 0,
\label{berexp}
\end{multline}
where the reciprocal of the Bond number, $\tilde{\sigma}$, is given by
\[
\tilde{\sigma} = \frac{\sigma}{\rho g L^2}.
\]
We note that we have tacitly assumed $\omega = \mathcal{O}(1)$.  In physical terms, this implies that $\omega$ is comparable to the natural time scale of this problem, $\sqrt{L/g}$.  If we were to assume $\omega$ were of larger magnitude, the problem would no longer be weakly nonlinear and would be much less amenable to asymptotic analysis.  Therefore, throughout the remainder of the paper, we assume that the vorticity is not too large.

The nonlocality in the full equations comes from the Hilbert transform, $\mathcal{H}$, defined by
\[
\mathcal{H}f = \frac{1}{2\pi}\int_{\mathbb{R}}dk~ e^{ikx}  i\mbox{sgn}(k) \hat{f}(k),
\]
where $\hat{f}(k)$ is the Fourier transform of $f(x)$ defined by
\[
\hat{f}(k) = \int_{\mathbb{R}}dx~ e^{-ikx}f(x).
\]
In other words, $\mathcal{H}$ is the operator with symbol $i\mbox{sgn}(k)$.  Given that
\[
\mbox{sgn}(k_{0} + \epsilon k) = \mbox{sgn}(k_{0})\mbox{sgn}\left( 1 + \frac{\epsilon k}{k_{0}}\right),
\]
assuming $\hat{f}(k)$ is a function of rapid decay, see \cite{folland},  and assuming $k_{0}\gg \epsilon$ gives
\begin{align}
\mathcal{H}\left( f(\epsilon x) e^{ik_{0}x} \right)& =
\frac{e^{ik_{0}x}}{2\pi}\int_{\mathbb{R}}dk~ e^{ik\epsilon x}
i\mbox{sgn}(k_{0}+\epsilon k) \hat{f}(k)\nonumber \\ & \sim i\mbox{sgn}(k_{0})e^{ik_{0}x}f(\epsilon x).\nonumber
\end{align}
with the difference being exponentially small.  On the other hand, if $k_{0}=0$, then
\[
\mathcal{H}\left( f(\epsilon x) \right) = \frac{1}{2\pi}\int_{\mathbb{R}}dk~ e^{ik\epsilon x}  i\mbox{sgn}(k) \hat{f}(k) = (\mathcal{H}f)(\epsilon x).
\]
Thus, the Hilbert transform does not have any significant effect on the multiple scales ans\"{a}tze made below.

Equation (\ref{integro1}) readily leads to a recursive formula for the surface velocity potential, $Q$, of the form
\begin{equation}
Q = \mathcal{H}\p_{t}\eta + \epsilon R_{1} + \epsilon^{2} R_{2} + \epsilon^{3} R_{3},
\label{Qexp}
\end{equation}
where
\begin{align*}
R_{1}= & \mathcal{H}\p_{x}\left(\eta\mathcal{H}\p_{t}\eta+ \frac{\omega}{2}\eta^{2} - \frac{1}{2}\mathcal{H}\p_{t}\eta^{2} \right),\\
R_{2}= & \mathcal{H}\p_{x}\left(\eta R_{1} \right) - \mathcal{H}\p_{x}^{2}\left(\frac{1}{6}\p_{t}\eta^{3} + \frac{\omega}{3}\mathcal{H}\eta^{3} + \frac{1}{2}\mathcal{H}(\eta^{2}\mathcal{H}\p_{t}\eta)\right),\\
R_{3} = & \mathcal{H}\p_{x}\left(\eta R_{2} \right) + \frac{1}{2}\p_{x}^{2}\left(\eta^{2}R_{1}\right) + \mathcal{H}\p_{x}^{3}\left(\frac{1}{24}\mathcal{H}\p_{t}\eta^{4} - \frac{1}{6}\eta^{3}\mathcal{H}\p_{t}\eta - \frac{\omega}{8}\eta^{4}\right).
\end{align*}
Coupling this to Equation \eqref{berexp} gives a single scalar equation defined entirely in terms of the surface height, $\eta$.  Now, use the ansatz
\begin{multline}
\eta(x,t) = \epsilon \eta_{0}(\xi,\tau) + \eta_{1}(\xi,\tau)e^{i\theta} +  \eta_{1}^{\ast}(\xi,\tau)e^{-i\theta} + \epsilon\left(\eta_{2}(\xi,\tau)e^{2i\theta} +  \eta_{2}^{\ast}(\xi,\tau)e^{-2i\theta}\right) \\
+ \epsilon^{2}\left(\eta_{3}(\xi,\tau)e^{3i\theta} +  \eta_{3}^{\ast}(\xi,\tau)e^{-3i\theta}\right)  + \epsilon^{3}\left(\eta_{4}(\xi,\tau)e^{4i\theta} +  \eta_{4}^{\ast}(\xi,\tau)e^{-4i\theta}\right) + \cdots,
\label{nlssurfexp}
\end{multline}
where
\begin{equation}
\theta = k_{0}x + \Omega t, ~ \xi = \epsilon(x + c_{g}t), ~ \tau = \epsilon^{2}t.
\label{xidef}
\end{equation}
We note that using integration by parts shows that terms of the form $\eta_{m}e^{im\theta}$ satisfy
\[
\int_{\mathbb{R}} dx ~ \eta_{m}(\xi,\tau) e^{im\theta}  = \mathcal{O}(\epsilon^{j}), ~ j\geq 1,~ m>0,
\]
which is to say that these terms are vanishingly small in an asymptotic sense.  Thus, asymptotically, Equation \eqref{zavsurf} becomes
\begin{equation}
\p_{\tau}\int_{\mathbb{R}} d\xi ~ \eta_{0}(\xi,\tau) = 0.
\label{zavzmean}
\end{equation}

By expanding and matching terms in the fundamental harmonic, $e^{i\theta}$, at $\mathcal{O}(1)$ we obtain the linear dispersion relation,
\begin{equation}
\Omega_{\pm}(k_{0},\omega)  = \frac{1}{2}\left(s\omega \pm \sqrt{\omega^{2} + 4|k_{0}|(1+\tilde{\sigma}k_{0}^{2})}\right),
\label{LDR}
\end{equation}
and at $\mathcal{O}(\epsilon)$ we find that
\[
c_{g} = \frac{1+3\tilde{\sigma}k_{0}^{2}}{2s\Omega - \omega},
\]
where $s=\mbox{sgn}(k_{0})$.  We note that Equation (\ref{LDR}) establishes that $\Omega_{+}(k_{0},\omega) > 0$ and $\Omega_{-}(k_{0},\omega) < 0$ for $k_{0}\neq0$.  Throughout the remainder of the paper, $\Omega$ always denotes the positive branch, $\Omega_{+}$.  By looking at the second harmonic, $e^{2i\theta}$, we find that
\[
\eta_{2} = \ell_{0}\eta_{1}^{2} + i\epsilon\ell_{1}\eta_{1}\p_{\xi}\eta_{1}+\mathcal{O}(\epsilon^2),
\]
where
\begin{align*}
\ell_{0} = & -\frac{k_{0}\left(\omega^{2}+2\Omega^{2}-4s\omega\Omega\right)}{2(k_{0} + \omega\Omega - 2s\Omega^{2} + 4\tilde{\sigma}k_{0}^{3})},\\
\ell_{1} = & \frac{\left(2c_{g}k_{0}(\Omega-s\omega)-2s\omega\Omega + \Omega^{2} + \omega^{2}/2\right)+(1+\omega c_{g} + 12k_{0}^{2}\tilde{\sigma}-4sc_{g} \Omega)\ell_{0}}{k_{0} + \omega\Omega - 2s\Omega^{2} + 4\tilde{\sigma}k_{0}^{3}}.
\end{align*}
By then going up to $\mathcal{O}(\epsilon^{3})$, we get the following coupled system describing the slow spatial and temporal modulation due to the interaction of the leading harmonic term, $\eta_{1}$, and the mean surface height, $\eta_{0}$,
\begin{multline}
\epsilon c_{g}^{2}\mathcal{H}\p_{\xi}^{2}\eta_{0} + (1+\omega c_{g})\p_{\xi}\eta_{0} + \epsilon \omega \p_{\tau} \eta_{0} + \omega\left(\omega-2s\Omega\right)\p_{\xi}\left|\eta_{1}\right|^{2}\\
+\epsilon\left(c_{g}(\omega-2s\Omega)\mathcal{H}\p_{\xi}^{2}\left|\eta_{1}\right|^{2} + is\omega c_{g}\left(\eta_{1}^{\ast}\p_{\xi}^{2}\eta_{1}-\eta_{1}\p_{\xi}^{2}\eta_{1}^{\ast}\right) \right) = 0,
\label{zaveq}
\end{multline}

\begin{multline}
(\omega-2s\Omega)\p_{\tau}\eta_{1} + i(c_{g}^{2}s-3k_{0}\tilde{\sigma})\p_{\xi}^{2}\eta_{1} + 2i\epsilon s c_{g}\p^{2}_{\xi\tau}\eta_{1} - \epsilon\tilde{\sigma} \p^{3}_{\xi}\eta_{1} \\
+ ik_{0}\omega\left(\omega-2s\Omega\right)\eta_{0}\eta_{1} +
i \alpha_{0}\left|\eta_{1}\right|^{2}\eta_{1}
+ \epsilon\alpha_{1}\left|\eta_{1}\right|^{2}\p_{\xi}\eta_{1}
+ \epsilon\alpha_{2}\eta_{1}^{2}\p_{\xi}\eta^{\ast}_{1} + \\
\epsilon i\alpha_{3}\eta_{1}\mathcal{H}\p_{\xi}\left|\eta_{1}\right|^{2}
 + \epsilon \omega(\omega-2s\Omega-|k_{0}|c_{g})\eta_{1}\p_{\xi}\eta_{0} + \epsilon\omega(\omega-2s\Omega-2|k_{0}|c_{g})\eta_{0}\p_{\xi}\eta_{1} \\
+ \epsilon ik_{0}c_{g}(\omega-2s\Omega)\eta_{1}\mathcal{H}\p_{\xi}\eta_{0} = 0,
\label{mnls}
\end{multline}
where
\begin{align*}
\alpha_{0} = & k_{0}^{2}\left(-\frac{3}{2}k_{0}^{3}\tilde{\sigma} + 2s\Omega^{2}-s\omega^{2}\right) + \ell_{0}k_{0}\left(\omega^{2}+2\Omega^{2}-4s\omega\Omega\right),\\
\alpha_{1} = & k_{0}\left(4c_{g}|k_{0}|\Omega - 9\tilde{\sigma}k_{0}^{3}+6s\Omega^{2}-3s\omega^{2}\right) \\
& + \ell_{0} \left(2\omega^{2}+4\Omega^{2}-8s\omega\Omega+4k_{0}c_{g}(\Omega-s\omega)\right) + \ell_{1}k_{0}\left(4s\omega\Omega-2\Omega^{2}-\omega^{2}\right),\\
\alpha_{2} = & k_{0}\left(-\frac{3}{2}k_{0}^{3}\tilde{\sigma}+2s\Omega^{2}-s\omega^{2}\right)  + \ell_{0}\left(\omega^{2} + 2\Omega^{2}-4s\omega\Omega\right),\\
\alpha_{3} = & k_{0}\left(\omega-2s\Omega\right)^{2}.
\end{align*}

By solving Equation \eqref{zaveq} to leading order and using Equation \eqref{zavzmean}, we find that
\begin{equation}
\eta_{0}(\xi,\tau) = \frac{\omega(2s\Omega-\omega)}{1+\omega c_{g}}\left|\eta_{1}\right|^{2}(\xi,\tau) +\mathcal{O}(\epsilon),
\label{zavsol}
\end{equation}
Likewise, from \eqref{mnls} at this order, and using the leading order solution for the mean height, $\eta_{0}$, gives the time-dependent NLS equation
\begin{equation}
i\p_{\tau}\eta_{1} + \alpha_{nl}\left|\eta_{1}\right|^{2}\eta_{1} + \alpha_{d}\p_{\xi}^{2}\eta_{1} = 0,
\label{NLSS}
\end{equation}
where
\begin{align*}
\alpha_{d}(k_{0},\omega) = & \frac{(c^2_{g} - 3|k_{0}|\tilde{\sigma})}{2\Omega-s\omega},\\
\alpha_{nl}(k_{0},\omega) = & \frac{k_{0}\left( sk_{0}^{3}\left(8 + \tilde{\sigma}k_{0}^{2} + 2(\tilde{\sigma}k_{0}^{2})^{2}\right) + \omega \alpha_{v}\right)}{\left(2s\Omega -\omega\right)(1+c_{g}\omega)\left(4\Omega^2-s(2k_{0}(1+4\tilde{\sigma}k_{0}^{2})+2\omega\Omega)\right)},
\end{align*}
and
\begin{align*}
\alpha_{v}(k_{0},\omega)= & s(c_g k_0  - 2\Omega ) \omega^4 + k_{0}(4 k_{0}^{2} s \tilde{\sigma} + 2 \Omega c_g  -s) \omega^3 \\
& + k_{0}(16 c_g k_{0}^{3} \tilde{\sigma} - 8 \Omega k_{0}^{2} \tilde{\sigma} + 10 c_g k_{0} - 6 \Omega) \omega^2 \\
& -  k_{0}^{2}(15 \Omega c_g k_0^2 s \tilde{\sigma} - 16 k_{0}^{4} \tilde{\sigma}^2 - 24 k_{0}^{2} \tilde{\sigma} - 2 ) \omega \\
& + k_{0}^{3}(2 c_g k_{0}^{4} s \tilde{\sigma}^2 + c_g k_{0}^{2} s \tilde{\sigma} - 15 \Omega k_{0}  s \tilde{\sigma} + 8c_g s).
\end{align*}
Note, our results for the NLS equation agree with those in \cite{afm} for $\omega=0$, and they agree with the results in \cite{thomas2012nonlinear} for $\tilde{\sigma}=0$.  Further, all of our results in this section are derived with the aid of the computer algebra system SAGE.

We readily see that we can go to higher order in the mean term using Equation \eqref{zaveq} and the expansion
\[
\eta_{0}(\xi,\tau) = \frac{\omega(2s\Omega-\omega)}{1+\omega c_{g}}\left|\eta_{1}\right|^{2}(\xi,\tau) + \epsilon \eta_{01}(\xi,\tau) + \mathcal{O}(\epsilon^{2}),
\]
which gives
\[
\eta_{01}= \frac{c_{g}(2s\Omega-\omega)}{(1+\omega c_{g})^{2}}\mathcal{H}\p_{\xi}\left|\eta_{1}\right|^{2} + \frac{i\omega}{1+\omega c_{g}}\left(\frac{\alpha_{d}\omega(2s\Omega-\omega)}{1+\omega c_{g}} + s c_{g} \right) \left(\eta_{1}\p_{\xi}\eta_{1}^{\ast} - \eta_{1}^{\ast}\p_{\xi}\eta_{1} \right).
\]
Thus, Equation \eqref{mnls} can be written in terms of $\eta_{1}$ alone
\begin{multline}
(\omega-2s\Omega)\p_{\tau}\eta_{1} + i(c_{g}^{2}s-3k_{0}\tilde{\sigma})\p_{\xi}^{2}\eta_{1} + 2i\epsilon s c_{g}\p^{2}_{\xi\tau}\eta_{1} - \epsilon\tilde{\sigma} \p^{3}_{\xi}\eta_{1} \\
- i \alpha_{nl}(\omega-2s\Omega)\left|\eta_{1}\right|^{2}\eta_{1}
+ \epsilon \tilde{\alpha}_{1}\left|\eta_{1}\right|^{2}\p_{\xi}\eta_{1}
+ \epsilon\tilde{\alpha}_{2}\eta_{1}^{2}\p_{\xi}\eta^{\ast}_{1} +
\epsilon i\tilde{\alpha}_{3}\eta_{1}\mathcal{H}\p_{\xi}\left|\eta_{1}\right|^{2}= 0.
\label{dysthe}
\end{multline}
Obtaining the coefficients, $\tilde{\alpha}_{j}$, is straightforward, and thus we omit writing them down for the sake of brevity.  We call Equation \eqref{dysthe} the Vor-Dysthe (VD) equation.  By solving it, we in effect find the next order correction to the NLS equation, and thus, by combining with our previous results, we can find $\eta$ up to $\mathcal{O}(\epsilon^{2})$ on a $\mathcal{O}(1/\epsilon^{2})$ timescale.  This becomes a critical feature necessary to ensure the accuracy of our numerics in Section 4.

For the NLS equation, the case in which $\alpha_{d}$ and $\alpha_{nl}$ have the same sign is known as the `focusing' case while the case in which they have opposite signs is known as the `defocusing' case.  These two cases are qualitatively different.  In the focusing case, the trivial-phase Jacobi elliptic solutions are given by
\begin{equation}
\eta_{1}(\xi,\tau) = \kappa\beta \sqrt{\frac{2\alpha_{d}}{\alpha_{nl}}}~\mbox{cn}(\beta \xi;\kappa)e^{-i\alpha_{d}\beta^{2}(1-2\kappa^2)\tau},
\label{cnsolns}
\end{equation}
where $0\le\kappa\le1$ is the elliptic modulus and $\beta$ is a positive lengthscale parameter.  In the $\kappa \rightarrow 1$ limit, these solutions limit to the `bright' soliton solutions
\[
\eta_{1}(\xi,\tau)=\beta \sqrt{\frac{2\alpha_{d}}{\alpha_{nl}}}~\mbox{sech}(\beta \xi)e^{i\alpha_{d}\beta^{2}\tau}.
\]
In the defocusing case, the trivial-phase Jacobi elliptic solutions are given by
\begin{equation}
\eta_{1}(\xi,\tau) = \kappa\beta \sqrt{-\frac{2\alpha_{d}}{\alpha_{nl}}}~\mbox{sn}(\beta \xi;\kappa) e^{-i\alpha_{d}\beta^{2}(1+\kappa^2)\tau}.
\label{snsolns}
\end{equation}
In the $\kappa\rightarrow1$ limit, these solutions limit to the `dark'
soliton solutions
\[
\eta_{1}(\xi,\tau)=\beta \sqrt{-\frac{2\alpha_{d}}{\alpha_{nl}}}~ \mbox{tanh}(\beta \xi)  e^{-2i\alpha_{d}\beta^{2}\tau}.
\]
The elevated profile of the magnitude of the bright soliton is qualitatively different from the depressed profile of the magnitude of the dark soliton.  This distinction typifies the difference between the behavior of the focusing and defocusing cases.  Note that in order for $\eta(x,t)$ to be periodic when Jacobi elliptic functions are used, the following restriction must be enforced
\begin{equation}
k_{0} = \frac{\pi \epsilon m}{2K(\kappa)},~ m\in \mathbb{Z},
\label{jacwavnum}
\end{equation}
where $K(\kappa)$ is the complete elliptic integral of the first kind.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Modulational Instabilities and Currents in Infinite Depth}
The plane-wave, or Stokes wave, solutions of the NLS equation are given by
\begin{equation}
\eta_{1}(\xi,\tau) = A e^{i\alpha_{nl}A^{2}\tau},
\end{equation}
where $A>0$ is a real constant.  Complementing the results in \cite{thomas2012nonlinear}, we study the stability of these solutions by considering perturbed solutions of the form
\begin{equation}
\eta_{1p}(\xi,\tau) = \Big{(}A +\mu \big{(}u_{p}(\xi,\tau)+iv_{p}(\xi,\tau)\big{)} +\mathcal{O}(\mu^2)\Big{)}\mbox{e}^{i\alpha_{nl}A^{2}\tau},
\label{pert}
\end{equation}
where $\mu$ is a small real parameter and $u_{p}$ and $v_{p}$ are real-valued functions.  Substituting (\ref{pert}) into the NLS equation and linearizing gives
\[
\pd_{\tau}\bp u_{p}\\ v_{p}\ep = \bp 0 & -\alpha_{d}\pd_{\xi}^{2} \\ \alpha_{d}\pd_{\xi}^{2} + 2A^{2}\alpha_{nl} & 0 \ep \bp u_{p} \\ v_{p} \ep.
\]
Separating variables, applying a Fourier transform in $\xi$, and introducing the $\xi$ wave number $l$, establishes that NLS plane-wave solutions are unstable with respect to the modulational instabilities (MIs) if
\[
0 < l^{2} \leq 2\frac{\alpha_{nl}}{\alpha_{d}}A^{2}.
\]
This reduces to the classic requirement that MIs are suppressed if the coefficients of dispersion and nonlinearity have opposite signs (i.e.~the defocusing case).

Figure \ref{fig:miplot} shows the values of $\omega$ and $k_{0}$ for which MIs exist (white areas) or not (black areas) for $\tilde{\sigma}=10^{-3}$ in Figure \ref{fig:miplot} (a), $\tilde{\sigma}=10^{-5}$ in Figure \ref{fig:miplot} (b), $\tilde{\sigma}=10^{-5}$ in Figure \ref{fig:miplot} (c), and $\tilde{\sigma}=0$ in Figure \ref{fig:miplot} (d).   Note, the scale of the axes changes from Figures \ref{fig:miplot} (a) and (b) to (c) and (d).  Further, while we assumed that $\omega = \mathcal{O}(1)$ in our derivation of the NLS equation, to see the impact of surface tension and its connection to MI, we have had to plot on an exaggerated scale in Figure \ref{fig:miplot}.  As can be seen, in the presence of surface tension, there does not appear to be a value of $\omega$ that suppresses MIs across all wave numbers.  This is in contrast to the zero surface tension case, in which if $k_{0}\omega>0$ and the magnitude of the shear is sufficiently large, then the MI is suppressed across all wave numbers.  As can be seen by comparing Figures \ref{fig:miplot} (a), (b), and (c), larger surface tension facilitates the transitions between regions in which the MI is or is not suppressed for relatively smaller values of carrier wavenumber and shear magnitude.  This expands on the results in \cite{thomas2012nonlinear}, which did not consider the effect of surface tension.  Throughout the rest of the paper we take $\tilde{\sigma}=10^{-5}$ which corresponds to supposing a characteristic wave length $L\sim 1m$.
\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=.48\textwidth]{Figure_1a} & \includegraphics[width=.48\textwidth]{Figure_1b}\\
(a) & (b)\\
\includegraphics[width=.48\textwidth]{Figure_1c} & \includegraphics[width=.48\textwidth]{Figure_1d}\\
(c) & (d)
\end{tabular}
\caption{\small {\it MIs exist} - white, \small{\it MIs suppressed} - black, for $\tilde{\sigma}=10^{-3}$ (a), $\tilde{\sigma}=10^{-5}$ (b), $\tilde{\sigma}=10^{-5}$ (c), and $\tilde{\sigma}=0$ (d).  Note the change in scale of the axes in (c) and (d) relative to (a) and (b).  Note, while we assumed that $\omega = \mathcal{O}(1)$ in our derivation of the NLS equation, to see the impact of surface tension and its connection to MI, we have had to plot on an exaggerated scale in this figure.}
\label{fig:miplot}
\end{figure}

Per our convention of taking the fast phase, $\theta(x,t)$, to be
\[
\theta(x,t) = k_{0}x + \Omega(k_{0},\omega) t,
\]
if $k_{0}\Omega(k_{0},\omega)>0$, then the carrier wave propagates to the left, and if $k_0 \Omega(k_{0},\omega)<0$, then the carrier wave propagates to the right.  Figures \ref{fig:miplot}(a) and (c) then show that MIs are generally suppressed when the shear current at the surface is co-propagating with respect to the carrier wave with sufficient strength which is inversely related to the carrier wavenumber.  Thus, in order to suppress most MIs, we need either relatively high carrier wave numbers for relatively weak shear currents, or relatively strong shear currents for relatively low carrier wave numbers.  The regions in Figures \ref{fig:miplot}(a) and (c) in which counter-propagating sheer currents suppress MIs are more complicated in nature, though one can argue that counter-propagating currents generally exacerbate MIs, especially with increasing shear strength.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Derivation of Velocity Formulas}
In non-dimensional coordinates, the position of a given fluid particle, $(x(t),z(t))$, is defined by the dynamical system
\begin{align}
\begin{split}
\dot{x} = \omega z + \epsilon \phi_{x}(x,z,t), ~~~~ \dot{z} = \epsilon\phi_{z}(x,z,t),\\
x(0) =x_0,~~~~ z(0)=z_0.
\end{split}
\label{system}
\end{align}
Therefore, in order to track the motion of a fluid particle, both $\phi_x$ and $\phi_z$ must be known throughout the fluid domain.  In order to determine asymptotic formulas for these quantities, let
\begin{equation}
\phi_{x}(x,z,t) = \frac{1}{2\pi}\int_{\mathbb{R}}dk~ e^{ikx} A(k,t)e^{|k|z},
\label{phix}
\end{equation}
\begin{equation}
\phi_{z}(x,z,t) = \frac{1}{2\pi}\int_{\mathbb{R}}dk~ e^{ikx} B(k,t)e^{|k|z}.
\label{phiz}
\end{equation}
Expanding $e^{\epsilon \eta |k|}$ in Equations (\ref{phix}) and (\ref{phiz}) gives
\begin{align}
\left.\phi_{x}\right|_{z=\epsilon\eta} = & \tilde{A}(x,t) - \epsilon\eta\mathcal{H}\pd_{x}\tilde{A} +\mathcal{O}(\epsilon^{2}),\label{phixsurfsol}\\
\left.\phi_{z}\right|_{z=\epsilon\eta} = & \tilde{B}(x,t) - \epsilon\eta\mathcal{H}\pd_{x}\tilde{B} +\mathcal{O}(\epsilon^{2}),\label{phizsurfsol}
\end{align}
where
\[
\tilde{A}(x,t) = \frac{1}{2\pi}\int_{\mathbb{R}}dk~ e^{ikx} A(k,t), ~~~~ \tilde{B}(x,t) = \frac{1}{2\pi}\int_{\mathbb{R}}dk~ e^{ikx} B(k,t).
\]

The surface boundary conditions give
\begin{align*}
\left.\phi_{x}\right|_{z=\epsilon\eta} = & \frac{Q-\epsilon\eta_{x}(\eta_{t}+\epsilon\omega\eta\eta_{x})}{1+\epsilon^{2}\eta_{x}^{2}},\\
\left.\phi_{z}\right|_{z=\epsilon\eta} = & \frac{\eta_{t}+\epsilon \left(Q+\omega\eta\right)\eta_{x}}{1+\epsilon^{2}\eta_{x}^{2}}.
\end{align*}
Substituting the expansions in Equations \eqref{Qexp} and \eqref{nlssurfexp} into these equations and using the results obtained during the derivation of the NLS equation gives
\begin{align}
\left.\phi_{x}\right|_{z=\epsilon\eta} = & -2\epsilon k_{0}\Omega \left|\eta_{1} \right|^{2} + \left(-s\Omega\eta_{1} + i\epsilon s c_{g}\pd_{\xi}\eta_{1} \right)e^{i\theta} \nonumber \\
& - \epsilon(2s\Omega\eta_{2}+ k_{0}(s\omega-\Omega) \eta_{1}^{2})e^{2i\theta} + \mbox{c.c.} + \mathcal{O}(\epsilon^{2}),\label{phisurfsollhsx}\\
\left.\phi_{z}\right|_{z=\epsilon\eta} = & (i\Omega \eta_{1}+\epsilon c_{g}\pd_{\xi}\eta_{1})e^{i\theta} + i\epsilon(2\Omega\eta_{2} + k_{0}(\omega-s\Omega)\eta_{1}^{2})e^{2i\theta} + \mbox{c.c.} + \mathcal{O}(\epsilon^{2}). \label{phisurfsollhsz}
\end{align}
This motivates the expansions
\begin{align*}
\tilde{A}(x,t) = &  \epsilon \tilde{A}_{01}(\xi,\tau) + \left(\tilde{A}_{10}(\xi,\tau) + \epsilon \tilde{A}_{11}(\xi,\tau)\right)e^{i\theta(x,t)} + \epsilon \tilde{A}_{21}(\xi,\tau)e^{2i\theta(x,t)} + \mbox{c.c.} + \mathcal{O}(\epsilon^{2}),\\
\tilde{B}(x,t) = &  \epsilon \tilde{B}_{01}(\xi,\tau) + \left(\tilde{B}_{10}(\xi,\tau) + \epsilon \tilde{B}_{11}(\xi,\tau)\right)e^{i\theta(x,t)} + \epsilon \tilde{B}_{21}(\xi,\tau)e^{2i\theta(x,t)} + \mbox{c.c.} + \mathcal{O}(\epsilon^{2}).
\end{align*}
Inserting these expansions into Equations \eqref{phixsurfsol} and \eqref{phizsurfsol} and matching powers of $\epsilon$ with the expansions in Equations \eqref{phisurfsollhsx} and \eqref{phisurfsollhsz} gives
\begin{align*}
\tilde{A}_{01} & =  0,\\
\tilde{A}_{10} & =  -s\Omega \eta_{1},\\
\tilde{A}_{11} & =  isc_{g}\pd_{\xi}\eta_{1},\\
\tilde{A}_{21} & =  -(2s\Omega \eta_{2} + k_{0}(s\omega - 2\Omega)\eta_{1}^{2}).
\end{align*}
The expressions for the corresponding terms in $\tilde{B}$ can be found in a similar manner.

Inverting the Fourier transforms leads to expressions for $A(k,t)$ and $B(k,t)$, which when inserted back into Equations \eqref{phix} and \eqref{phiz} gives
\[
\phi_{x}(x,z,t) = \tilde{\phi}_{x}(x,z,t) + \mathcal{O}(\epsilon^{2}), ~ \phi_{z}(x,z,t) = \tilde{\phi}_{z}(x,z,t) + \mathcal{O}(\epsilon^{2}),
\]
where
\begin{equation}
\tilde{\phi}_{x}(x,z,t) = R_{1}(\xi,z,\tau)e^{i\theta} \\+ \epsilon R_{2}(\xi,z,\tau)e^{2i\theta} + \mbox{c.c.},
\label{phixexp}
\end{equation}
\begin{equation}
\tilde{\phi}_{z}(x,z,t) =  \tilde{R}_{1}(\xi,z,\tau)e^{i\theta} \\+ \epsilon \tilde{R}_{2}(\xi,z,\tau)e^{2i\theta} + \mbox{c.c.}.
\label{phizexp}
\end{equation}
Again, while the expansion procedure is straightforward, due to the length of the expressions involved, the details have been omitted.  However, we can still obtain approximations that will be useful later in this paper.  In particular, the leading order behaviors of $R_{1}$ and $\tilde{R}_{1}$, which in turn give the leading order behaviors of the velocity components, are given by
\begin{align}
R_{1}(\xi,z,\tau) \sim & -\frac{s\Omega}{2\pi}\int_{\mathbb{R}} dk ~ e^{i k \xi} e^{|k_{0}+\epsilon k|z}\hat{\eta}_{1}(k,\tau), \label{roeqx}\\
\tilde{R}_{1}(\xi,z,\tau) \sim & \frac{i\Omega}{2\pi}\int_{\mathbb{R}} dk ~ e^{i k \xi} e^{|k_{0}+\epsilon k|z}\hat{\eta}_{1}(k,\tau).
\label{roeqz}
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Lagrangian and Stokes Drift Velocity}
We now show how the above approximation schemes can be used to determine how a background shear current modifies the Lagrangian drift velocity  and Stokes drift velocity.  To do so, we make use of the Generalized Lagrangian Mean (GLM) formalism presented in \cite{andrews}.  This approach is built via a diffeomorphic mapping of the original Eulerian spatial coordinates, ${\bf x}$, to the mean-position coordinates, $\tilde{\bf x}$, of the form
\[
{\bf x} = \tilde{{\bf x}} + {\bf y}(\tilde{{\bf x}},t),
\]
so that if Lagrangian paths in the original coordinates are found from the differential equation
\[
\frac{d{\bf x}}{dt} = {\bf u}({\bf x},t),
\]
then the corresponding mean paths are found from the differential equation
\begin{equation}
\frac{d\tilde{{\bf x}}}{dt} = \bar{{\bf u}}^{L}(\tilde{{\bf x}},t),
\label{meanode}
\end{equation}
where the vector-field, $\bar{{\bf u}}^{L}$, is the pull back, relative to the mapping above, of the original vector field, ${\bf u}$; see \cite{buhler}.  The vector field, $\bar{{\bf u}}^{L}$, is called the Lagrangian drift velocity (LDV).  The equivalent mean-Eulerian representation of the mean differential equation is
\begin{equation}
\left(\pd_{t} + \overline{{\bf u}}^{L}(\tilde{{\bf x}},t)\cdot \nabla_{\tilde{{\bf x}}} \right) \left(\tilde{{\bf x}} + {\bf y}(\tilde{{\bf x}},t)\right) = {\bf u}\left(\tilde{{\bf x}} + {\bf y}(\tilde{{\bf x}},t),t\right).
\label{speeddef}
\end{equation}
To fully specify the mapping, and thereby connect it to an averaging procedure, we require for some chosen averaging operator $\bar{(\cdot)}$ that
\begin{align*}
\overline{{\bf y}(\tilde{{\bf x}},t)} = & \bf{0}, \\
\overline{\overline{{\bf u}}^{L}(\tilde{{\bf x}},t)} = & \overline{{\bf u}}^{L}(\tilde{{\bf x}},t).
\end{align*}
This then allows us to define the Lagrangian mean, $\bar{()}^{L}$, at a mean point, $\tilde{{\bf x}}$, relative to the disturbance, ${\bf y}(\tilde{{\bf x}},t)$, to be
\[
\bar{\varphi}^{L}(\tilde{{\bf x}},t) = \overline{\varphi^{\bf y}(\tilde{{\bf x}},t)},
\]
where, using the language in \cite{buhler}, the ${\bf y}$-lift, $\varphi^{\bf y}(\tilde{{\bf x}},t)$, of $\varphi$ is given by
\[
\varphi^{\bf y}(\tilde{{\bf x}},t) = \varphi(\tilde{{\bf x}}+{\bf y}(\tilde{\bf x},t),t).
\]
Note, if we define as in \cite{andrews} the mean material derivative $\bar{D}^{L}$ to be
\[
\bar{D}^{L} = \left(\pd_{t} + \overline{{\bf u}}^{L}(\tilde{{\bf x}},t)\cdot \nabla_{\tilde{{\bf x}}} \right),
\]
then upon averaging Equation \eqref{speeddef}, we get back Equation \eqref{meanode}, thereby showing the self-consistency of this approach.  To fix ideas, throughout the remainder of this paper, the average will be given by integration with respect to the fast phase variable $\theta$, i.e.
\[
\bar{f}(\xi,z,\tau) = \frac{1}{2\pi}\int_{0}^{2\pi} d\theta~ f(\xi,z,\tau,\theta).
\]
Note, we drop the tildes denoting the mean space coordinates for brevity and since from context which space we are in should always be clear.

\subsubsection*{Constant Vorticity, Psuedomomentum, and the Fundamental Equations of the LDV}
Let $C$ be an arbitrary smooth, closed, simple contour, and let $C^{{\bf y}}$ denote its ${\bf y}$-lift, with the restriction that it remain within the bulk of our fluid domain, i.e. it is not at the fluid surface $z=\epsilon \eta(x,t)$.  Then, for a constant vorticity flow, it is straightforward to show that the circulation $\Gamma$ around $C^{{\bf y}}$ is given by
\[
\Gamma = \oint_{C^{{\bf y}}} {\bf u}\cdot d{\bf x} = \omega \int_{\mbox{int}(C^{{\bf y}})}dA = \omega \int_{ \mbox{int}(C)} J dA,
\]
where $J$ is the Jacobian of the map ${\bf x} + {\bf y}({\bf x},t)$.   Following the arguments in \cite{buhler}, we then have
\[
\oint_{C} \left(\bar{{\bf u}}^{L} - {\bf p} \right)\cdot d{\bf x} = \int_{\mbox{int}(C)} \nabla \times \left(\bar{{\bf u}}^{L} - {\bf p} \right)dA = \omega \int_{ \mbox{int}(C)} \bar{J} dA,
\]
where the pseudomomentum ${\bf p}$ is given by
\[
{\bf p} = -\overline{\nabla_{{\bf y}}({\bf y}\cdot {\bf u}^{l})}, ~{\bf u}^{l}({\bf x},t) = {\bf u}^{\bf y}({\bf x},t) - \bar{{\bf u}}^{L}({\bf x},t),
\]
where $\nabla_{{\bf y}}$ uses Feynman notation, and where the curl term is understood to yield the relevant scalar magnitude since the problem is planar.  Since $C$ was arbitrary, we get the equivalent mean-constant vorticity expression
\begin{equation}
\nabla \times \left(\bar{{\bf u}}^{L} - {\bf p} \right) = \omega \bar{J}.
\label{curleq}
\end{equation}
Note, this expression is exact.  Likewise, since the density of the original flow is assumed to be a constant, say $\rho_{0}$, the Lagrangian-mean mass conservation equation; see \cite{andrews,buhler}, becomes after averaging
\begin{equation}
\bar{D}^{L}\bar{J} + \bar{J}\nabla\cdot \bar{{\bf u}}^{L} = 0.
\label{diveq}
\end{equation}
As given in \cite{buhler}, for planar flow the Jacobian, $J$, is found from
\[
J = 1 + \nabla \cdot {\bf y} + \p_{z}y_{1}\p_{x}y_{2} - \p_{x}y_{1}\p_{z}y_{2}.
\]
We call Equations \eqref{curleq} and \eqref{diveq} the Fundamental Mean Equations (FMEs) for the LDV.  Taken together, they give the divergence and curl of the vector field $\bar{{\bf u}}^{L}$ in terms of the fluctuation vector ${\bf y}$, thereby giving us all of the necessary information to compute the LDV.  We show in the following section how to derive an approximation to ${\bf y}$, which then allows us to asymptotically solve the FMEs.

\subsubsection*{Determining the Fluctuations {\bf y} and the FMEs}
In order to then determine the disturbance ${\bf y}$, we restrict to the special case in which the diffeomorphism between the original Eulerian coordinates and the mean coordinates is a near identity transformation of the form
\[
{\bf x} = \tilde{{\bf x}} + \epsilon{\bf y}(\tilde{{\bf x}},t).
\]
Along with this, we suppose that the LDV has the asymptotic form
\[
\bar{{\bf u}}^{L} = \bp \omega z \\ 0 \ep + \epsilon^{2}\bar{{\bf u}}^{L}_{2},
\]
so that Equation \eqref{speeddef} establishes that
\begin{equation}
\pd_{t}{\bf y} + \omega z \pd_{x}{\bf y} - \omega \bp y_{2}\\ 0 \ep = \bp R_{1}e^{i\theta}+R^{\ast}_{1}e^{-i\theta}\\ \tilde{R}_{1}e^{i\theta}+\tilde{R}^{\ast}_{1}e^{-i\theta} \ep + \mathcal{O}(\epsilon).
\label{vareqn}
\end{equation}
The method of characteristics gives
\begin{align}
y_{1} = & - \frac{i}{k_{0}\omega z + \Omega}\left(R_{1}e^{i\theta} -R^{\ast}_{1}e^{-i\theta} \right) - \frac{\omega}{(k_{0}\omega z + \Omega)^{2}}\left(\tilde{R}_{1}e^{i\theta} + \tilde{R}^{\ast}_{1}e^{-i\theta} \right),\label{meanfc1}\\
y_{2} = & - \frac{i}{k_{0}\omega z + \Omega}\left(\tilde{R}_{1}e^{i\theta} -\tilde{R}^{\ast}_{1}e^{-i\theta} \right) \label{meanfc2}.
\end{align}
Clearly this result is not valid when $z\sim z_{c} \equiv c_{p}(k,\omega)/\omega$, where the phase speed, $c_{p}$, is given by
\[
c_{p}(k_{0},\omega) = -\frac{\Omega(k_{0},\omega)}{k_{0}}.
\]
Again, due to our choice of signs in $\theta(x,t)$, we use the opposite sign on the phase speed so that a positive phase speed $c_{p}$ corresponds to a rightward propagating phase.   We see that $z_{c}$ typifies a kind of stagnation depth in which the carrier wave and the shear profile cancel one another out.  Choosing $k_{0}>0$, we see that as $\omega \gg 1$, $z_{c}\lessapprox -1$, which is, asymptotically speaking, well removed from the surface $z=\epsilon \eta(x,t)$.  However in the case that $\omega \ll -1$, we see that, while $z_{c}$ is always positive, $z_{c}\sim 0$.  Thus, for stronger negative shear values, the expansions above break down around the surface.    Throughout the remainder of the paper, when we look at the case in which $k_{0}\omega <0$, we choose other parameters so as to prevent the surface from overlapping with a region in which the GLM expansions above break down. We further explore what happens near $z_{c}$ in the case $k_{0}\omega >0$ at the end of this section.

Away from the critical depth, $z_{c}$, using the leading order expansions
\[
R_{1} \sim -s\Omega\eta_{1}e^{|k_{0}|z}, ~ \tilde{R}_{1} \sim i\Omega\eta_{1}e^{|k_{0}|z},
\]
gives
\begin{align*}
y_{1} \sim & \frac{-i}{k_{0}\omega z + \Omega}\left(\frac{\omega\Omega}{k_{0}\omega z + \Omega} - s\Omega \right)(\eta_{1}e^{i\theta} - \eta_{1}^{\ast}e^{-i\theta})e^{|k_{0}|z},\\
y_{2} \sim & \frac{\Omega}{k_{0}\omega z + \Omega}(\eta_{1}e^{i\theta} + \eta^{\ast}_{1}e^{-i\theta})e^{|k_{0}|z}.
\end{align*}
We then find the mean Jacobian, $\bar{J}$, to be
\[
\bar{J}(\xi,\tau,z) = 1 -2k\Omega^{2}|\eta_{1}(\xi,\tau)|^{2}\epsilon^{2}\p_{z}\left(\frac{e^{2|k_{0}|z}}{k_{0}\omega z + \Omega}\left(\frac{-s}{k_{0}\omega z + \Omega} + \frac{\omega}{(k_{0}\omega z + \Omega)^{2}}\right)\right) + \mathcal{O}(\epsilon^{3}).
\]
We modify our assumption for the expansion of the LDV so that the horizontal and vertical speeds do not appear at the same order, i.e.
\[
\bar{{\bf u}}^{L}(\xi,\tau,z) = \bp \omega z \\ 0 \ep + \epsilon^{2} \bp \bar{u}^{L}_{2}(\xi,\tau,z) \\ \epsilon \bar{w}^{L}_{3}(\xi,\tau,z)\ep,
\]
whereby, using the expansion for $\bar{J}$ above, the divergence equation of the FMEs, Equation \eqref{diveq}, gives
\[
\p_{\xi}\bar{u}^{L}_{2} + \p_{z}\bar{w}^{L}_{3} = 2k(c_{g}+\omega z)\Omega^{2}\p_{\xi z}^{2}\left(\frac{e^{2|k_{0}|z}|\eta_{1}|^{2}}{k_{0}\omega z + \Omega}\left(\frac{-s}{k_{0}\omega z + \Omega} + \frac{\omega}{(k_{0}\omega z + \Omega)^{2}}\right)\right) .
\]
Note, without introducing the asymmetry in scaling between the horizontal and vertical mean velocities, we do not get a meaningful leading order equation.  We can likewise readily find that
\[
{\bf u}^{l} = \epsilon \bp \omega y_{2} + \tilde{\phi}_{x} \\ \tilde{\phi}_{z} \ep + \mathcal{O}(\epsilon^{2}),
\]
and thus the curl equation of the FMEs, Equation \eqref{curleq}, gives us to the relevant asymptotic order
\[
\p_{z}\bar{u}^{L}_{2} =  -2k_{0}\Omega^{2} |\eta_{1}|^{2}\p_{z}\left(\left(\frac{2}{k_{0}\omega z + \Omega} - \frac{3s\omega}{(k_{0}\omega z + \Omega)^{2}}+\frac{2\omega^{2}}{(k_{0}\omega z + \Omega)^{3}}\right)e^{2|k_{0}|z}\right).
\]
We thus see that we can readily solve for $\bar{u}^{L}_{2}$, so that
\[
\bar{u}^{L}_{2}(\xi,\tau,z) = \tilde{C}(\xi,\tau)  -2k_{0}\Omega^{2} |\eta_{1}|^{2} \left(\frac{2}{k_{0}\omega z + \Omega} - \frac{3s\omega}{(k_{0}\omega z + \Omega)^{2}}+\frac{2\omega^{2}}{(k_{0}\omega z + \Omega)^{3}}\right)e^{2|k_{0}|z}.
\]
Then, from the divergence equation, we can find $\bar{w}^{L}_{3}$.  However, the solution for $\bar{u}^{L}_{2}$ necessarily introduces an integration constant in the form $\tilde{C}(\xi,\tau)$.  Thus, to fully determine the LDV at this asymptotic order, we need boundary conditions.  These are found via a lifting of the free surface $z=\epsilon \eta(x,t)$.

\subsubsection*{Lifting the Free Surface}
The equation defining the surface, $z=\epsilon\eta(x,t)$ is a constraint, so expanding on the comment in Section 4.2 of \cite{andrews}, and paralleling the approach in \cite{ardhuin}, to define the lift of the surface, we introduce the surface fluctuation $y_{s}(\tilde{x},t)$ so that we define the mean surface position $\tilde{z}$ to be
%and thus while we could certainly lift the quantity $\varphi_{s}(x,z,t)=z-\epsilon\eta(x,t)$, it is not clear from the GLM formalism how to lift the level sets of a function, since, at least in some non-trivial neighborhood of a root, the Implicit-Function Theorem essentially entrains one variable in terms of the others, thereby removing a degree of freedom at the surface.  Thus,
\[
\tilde{z} = \epsilon \overline{\eta(\tilde{x}+\epsilon y_{s}(\tilde{x},t),t)}\equiv \epsilon \bar{\eta}^{L}(\tilde{x},t).
\]
This likewise defines the associated surface Lagrangian mean velocity
\[
\bar{{\bf u}}^{L}_{s}(\tilde{x},t) = \overline{{\bf u}(\tilde{x}+\epsilon y_{s}(\tilde{x},t),\epsilon\eta(\tilde{x}+\epsilon y_{s}(\tilde{x},t),t),t)}.
\]
While we have only one fluctuation term, $y_{s}$, but two components of the surface Lagrangian mean velocity, it is not problematic to define
\[
\left(\p_{t} +\bar{u}^{L}_{s}\p_{\tilde x}\right)(\tilde{x}+\epsilon y_{s}(\tilde{x},t) )= u(\tilde{x}+\epsilon y_{s}(\tilde{x},t),\epsilon\eta(\tilde{x}+\epsilon y_{s}(\tilde{x},t),t),t),
\]
since the kinematic condition derived from $z=\epsilon \eta(x,t)$ naturally gives us the other mean velocity component via
\[
\bar{w}^{L}_{s} = \epsilon \overline{(u\eta_{\tilde{x}})^{y_{s}}} + \epsilon \p_{t}\bar{\eta}^{L}.
\]

Using the formula for the velocity at the surface, and noting that we readily see that $\bar{u}^{L}_{s} = \mathcal{O}(\epsilon^{2})$, gives us the evolution equation for the surface fluctuations $y_{s}$ as
\[
\p_{t}y_{s} = (\omega -s\Omega) \left(\eta_{1}e^{i\theta} + \eta_{1}^{\ast}e^{-i\theta}\right) + \mathcal{O}(\epsilon),
\]
so that
\[
y_{s} \sim -\frac{i(\omega-s\Omega)}{\Omega}  \left(\eta_{1}e^{i\theta} - \eta_{1}^{\ast}e^{-i\theta}\right).
\]
This then gives the mean Lagrangian surface as
\[
\bar{\eta}^{L}(\xi,\tau) = \epsilon \left(\eta_{0}(\xi,\tau) - \frac{2k_{0}(\omega-s\Omega)}{\Omega}|\eta_{1}(\xi,\tau)|^{2}\right) + \mathcal{O}(\epsilon^{2}),
\]
and the mean Lagrangian horizontal surface speed is found to be
\[
\bar{u}^{L}_{s}(\xi,\tau) = \omega\eta_{0}(\xi,\tau)-2k_{0}\Omega\left( 1 + \left(1-\frac{s\omega}{\Omega}\right)^{2}\right)|\eta_{1}(\xi,\tau)|^{2}.
\]
We can condense this formula into the form
\[
\bar{u}^{L}_{s}(x,t) = u^{L}_{p}(k_{0},\omega)|\eta_{1}(\xi,\tau)|^{2},
\]
where the scaling factor $u^{L}_{p}$ is given by
\[
u^{L}_{p}(k_{0},\omega) =  \frac{ \omega^{2}(2s\Omega - \omega)}{1+\omega c_{g}} - 2k_{0}\Omega\left(1 + \left(1 - \frac{s\omega}{\Omega} \right)^{2}\right).
\]
The LDV determines the mean horizontal velocity of a particle at or near the surface, and this is then controlled by the magnitude of the solution to the NLS equation we study and the magnitude of $u^{L}_{p}$.  Of particular interest given the puzzling results on the existence of drift quenching Eulerian counterflows, see \cite{monismith,smith}, we can also then determine vorticity values $\omega$ for a given wavenumber $k_{0}$ such that $u^{L}_{p}(k_{0},\omega) =0$, which corresponds to the presence of an Eulerian counterflow which counters the effect of the SDV.  We plot this zero level set for $0\leq k_{0}\leq 20$ in Figure \ref{fig:zerodriftk0}.  As can be seen, shear strengths which contribute to the suppression of particle drift are always counter-propagating relative to a positively elevated surface.
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figure_2}
\caption{\small Plot of the zero set $u^{L}_{p}(k_{0},\omega)=0$.  The dot denotes the point $(k_{0},\omega)=(1,1.6818)$.}
\label{fig:zerodriftk0}
\end{figure}

At this point, in order to properly couple our mean surface formulation to the bulk GLM formulation described above, we then require
\[
\left.\bar{{\bf u}}^{L}\right|_{\tilde{z}=\epsilon \bar{\eta}^{L}} = \bar{{\bf u}}^{L}_{s}(\tilde{x},t).
\]
Thus, we then find the integration constant $\tilde{C}(\xi,\tau)$ in the formula for the horizontal LDV $\bar{u}^{L}_{2}$ to be
\[
\tilde{C}(\xi,\tau) = 2k_{0}\omega\left(-s + \frac{\omega}{\Omega} \right)|\eta_{1}(\xi,\tau)|^{2} + \omega \eta_{0}(\xi,\tau).
\]
Thus, we can find $\bar{u}_{2}^{L}$ as
\begin{multline*}
\bar{u}^{L}_{2}(\xi,\tau,z) = 2k_{0}\omega\left(-s + \frac{\omega}{\Omega} \right)|\eta_{1}(\xi,\tau)|^{2} + \omega \eta_{0}(\xi,\tau) \\
\frac{2\Omega^{2} }{c_{p}-\omega z}\left( 2 + \frac{3\omega}{|k_{0}|(c_{p}-\omega z)}+\frac{2\omega^{2}}{k_{0}^{2}(c_{p}-\omega z)^{2}}\right)|\eta_{1}(\xi,\tau)|^{2} e^{2|k_{0}|z}.
\end{multline*}
We note that if we take the vorticity $\omega=0$, we then get for a left-traveling wave
\[
\bar{u}^{L}_{2}(\xi,\tau,z) = -4k_{0}\Omega|\eta_{1}(\xi,\tau)|^{2}e^{2|k_{0}|z}
\]
Thus, at zero vorticity, we recreate the classic result for the Stokes drift in an irrotational fluid; see \cite{longuet}.  Likewise, even in the case of zero vorticity, we see the impact of the slowly varying envelope introduced by the NLS equation in the need to solve the, in part, slow-variable dependent FMEs, which allows for a more complex relationship between $\bar{{\bf u}}^{L}_{2}$ and the pseudomomentum; see \cite{buhler} for further discussion of this issue.  We likewise note that it is in \cite{buhler} that one sees a clear program for determining $\bar{{\bf u}}^{L}$ via the vorticity and divergence equations we used above in the FMEs.  The contribution presented herein was in clearly connecting the FMEs to the lifted surface equations, thereby allowing for the connection of the bulk mean velocity to an unambiguously averaged surface result and a continuously defined LDV.

Following \cite{andrews}, we define the Stokes drift of a quantity, say $\bar{\varphi}^{S}$ to be
\[
\bar{\varphi}^{S} = \bar{\varphi}^{L} - \bar{\varphi}.
\]
Thus, if we average the fluid velocity in the bulk of the fluid, Equations \eqref{phixexp} and \eqref{phizexp} show that the bulk Eulerian mean horizontal velocity is given by $\bar{u}(\xi,\tau,z)$ where
\[
\bar{u}(\xi,\tau,z) = \omega z + \mathcal{O}(\epsilon^{3}).
\]
Thus, in the bulk, the horiontal SVD, say $\bar{u}^{S}_{2}(\xi,\tau,z) = \bar{u}^{L}_{2}(\xi,\tau,z)$.  We emphasize that $z$ is a mean coordinate, and thus we do not evaluate the LVD or SVD at the free surface $z=\epsilon \eta(x,t)$, but instead at the lifted surface $\tilde{z}=\epsilon \bar{\eta}^{L}$.  Near the lifted free surface, by removing the Eulerian mean term $\omega \eta_{0}$, we have that the surface SDV in the horizontal direction, say $\bar{u}^{S}_{s}$ is
\begin{equation}
\bar{u}^{S}_{s}(\xi,\tau) = -2\epsilon^{2}k_{0}\Omega\left(1 + \left(1-\frac{s\omega}{\Omega} \right)^{2}\right)|\eta_{1}(\xi,\tau)|^{2}.
\label{surfSDV}
\end{equation}
Note, this multi-part definition for the SDV reflects the intricacies of averaging around a freely evolving surface.  This explains our privileging of the LDV over the SDV.  Our result then provides support for the phenomenological choices of how the SDV of a plane-wave varies in depth made in \cite{breivik}.  There, the classic formula for the SDV was modified by multiplying it by terms like $1/(1+\tilde{c}z)$, where $\tilde{c}$ was then chosen to improve agreements with observed data.  Our result provides a physical motivation for the appearance of such terms.  However, while we can put forth a physical mechanism, whether unresolved shear currents near the surface are truly responsible for disagreements between theory and observation is a matter for future work.

\subsubsection*{The Surface SDV for Modulated Surface Waves}
The surface SDV in Equation \eqref{surfSDV} for the Jacobi elliptic function solutions with $\beta=1$ is given by
\[
\bar{u}^{S}_{s}(\xi,\tau) = -4\epsilon^{2}\kappa^{2}k_{0}\Omega \left|\frac{\alpha_{d}}{\alpha_{nl}}\right| \left(1 + \left(1 - \frac{s\omega}{\Omega} \right)^{2} \right)\phi(\xi;\kappa),
\]
where
\[
\phi(\xi;\kappa) = \left\{\ba{rl} \mbox{cn}^2(\xi;\kappa), & \alpha_{d}/\alpha_{nl}>0, \\  \mbox{sn}^2(\xi;\kappa), & \alpha_{d}/\alpha_{nl}<0,\ea \right.
\]
where $\xi$ is defined in Equation (\ref{xidef}).  Defining the parameter
\[
\tilde{u}^{S}(k_{0},\omega) = -4 k_{0}\Omega(k_{0},\omega) \left|\frac{\alpha_{d}(k_{0},\omega)}{\alpha_{nl}(k_{0},\omega)}\right|\left(1 + \left(1 - \frac{s\omega}{\Omega(k_{0},\omega)} \right)^{2} \right),
\]
and choosing the positive branch of the dispersion relationship, Figure \ref{fig:sdriftmag_sign} shows that this term exhibits a wide variation in magnitude.  The curves along which the largest magnitudes are seen correspond to the level set $\alpha_{nl}(k_{0},\omega)=0$.  Thus, for the class of Jacobi elliptic solutions, there appears to be a kind of nonlinear resonance between the shear current and the carrier wave.  We point out that along this curve though, the assumptions we used to derive the NLS equation are no longer valid.  Therefore, more work should be done to better elucidate the affiliated dynamics associated with parameter choices defining said curve.  Throughout the remainder of this paper, we choose parameter values that do not place us too close to the zero set of $\alpha_{nl}(k_{0},\omega)=0$.
\begin{figure}
\centering
\begin{tabular}{c}
\includegraphics[width=.55\textwidth]{Figure_3}
\end{tabular}
\caption{\small Log plot of the surface SDV parameter $\tilde{u}^{S}(k_{0},\omega)$.  The curves along which $\tilde{u}^{S}(k_{0},\omega)=0$ mirror the transitions between MI stable and unstable regions seen in Figure \ref{fig:miplot}.  The surface SDV is most augmented near the curves corresponding to $\alpha_{nl}(k_{0},\omega)=0$.  }
\label{fig:sdriftmag_sign}
\end{figure}

To get a more detailed understanding of how the surface SDV depends on the shear current strength, we choose $k_{0}=1$ and use the positive branch of the dispersion relationship.  Figure \ref{fig:stksdrfcomp} contains plots of $\tilde{u}^{S}(1,\omega)$ for $-1\leq \omega\leq 4$, $-1\leq \omega \leq 1.1$, and $1.2\le\omega\le4$.  We choose these particular ranges of shear values to in particular examine the behavior of $\tilde{u}^{S}$ around the resonance curve seen in Figure \ref{fig:sdriftmag_sign}.  The singularity seen in Figure \ref{fig:stksdrfcomp} (a) corresponds to the value $\omega$ such that $\alpha_{nl}(1;\omega)=0$, namely $\omega\approx1.1550$.  Figure \ref{fig:stksdrfcomp} (b) shows the focusing side and that increasing the shear strength increases the Stokes drift velocity.  Figure \ref{fig:stksdrfcomp} (c) shows that this relationship is reversed on the defocusing side.
\begin{figure}
\centering
\begin{tabular}{ccc}
\includegraphics[width=.32\textwidth]{Figure_4a} & \includegraphics[width=.32\textwidth]{Figure_4b} & \includegraphics[width=.32\textwidth]{Figure_4c}\\
(a) & (b) & (c)
\end{tabular}
\caption{\small Plots of $\tilde{u}^{S}(1,\omega)$ for $-1\leq \omega \leq 4$ (a), $-1\leq \omega \leq 1.12$ (b), and $1.17\leq \omega \leq 4$ (c).  Note, $\alpha_{nl}(1;1.1550)=0$.  This shows that the magnitude of the SDV increases with increasing shear on the focusing side of the singularity (b), and decreases with increasing shear on the defocusing side (c).}
\label{fig:stksdrfcomp}
\end{figure}

\subsection{Flow Near the Depth $z_{c}$}
Again the depth $z_{c}=c_{p}/\omega$ is the depth at which our approximations to the mean particle path fluctuations in Equations \eqref{meanfc1} and \eqref{meanfc2} are no longer valid.  Taking $k_{0}\omega >0$ so that $-\infty<z_{c}<-1$, it is illuminating to attempt to examine particle paths near this depth.  To do this, we look at the special case of a plane wave solution to the NLS equation, so that we can readily write
\begin{align*}
\dot{x} = & \omega z - 2\epsilon s\Omega A e^{|k_{0}|z}\cos(\tilde{\theta}) + \mathcal{O}(\epsilon^{2})\\
\dot{z} = & -2\epsilon\Omega A e^{|k_{0}|z}\sin(\tilde{\theta}) + \mathcal{O}(\epsilon^{2}),
\end{align*}
where $\tilde{\theta} = \theta + \epsilon^{2}\alpha_{nl}A^{2}t$.  At this order, we can rewrite this system as
\begin{align*}
\dot{\tilde{\theta}} = & \Omega+k_{0}\omega z - 2\epsilon |k_{0}|\Omega A e^{|k_{0}|z}\cos(\tilde{\theta}) + \mathcal{O}(\epsilon^{2}),\\
\dot{z} = & -2\epsilon\Omega A e^{|k_{0}|z}\sin(\tilde{\theta}) + \mathcal{O}(\epsilon^{2}),
\end{align*}
which to leading order is seen to have the Hamiltonian $H(\tilde{\theta},z)$ where
\[
H(\tilde{\theta},z) = \Omega z + \frac{k_{0}\omega z^{2}}{2} - 2\epsilon\Omega A e^{|k_{0}|z}\cos(\tilde{\theta}).
\]
This has critical points at
\[
\tilde{\theta}=n \pi, ~ z = z_{c} - \epsilon\frac{2c_{p} e^{|k_{0}|z_{c}}}{\omega} (-1)^{n} + \mathcal{O}(\epsilon^{2}),
\]
with even $n$ corresponding to centers and odd $n$ to saddles.  This result then is a good indication for the presence of `cats-eye' patterns around the depth $z_{c}$, a result echoing the works of \cite{simmen,dasilva,wahlen2} and \cite{nachbin} among many other authors.  This presents an interesting matching problem with regards to finding uniform approximations to the LDV.  This however is a question beyond the scope of the present paper.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Results}
In order to determine particle paths, say $(x(t),z(t))$, in general, the system in Equation (\ref{system}) must be solved.  However, we focus on determining particle paths at the surface. Thus, letting $\tilde{\eta}$ denote the truncation of Equation \eqref{nlssurfexp} up to and including $\mathcal{O}(\epsilon)$ terms, we approximate the vertical component of the path via the restriction
\[
z(t) = \epsilon \tilde{\eta}(x(t),t).
\]
We then can readily find the scalar equation for the horizontal coordinate $x(t)$ to be
\[
\dot{x} = \epsilon\omega\tilde{\eta}(x,t) + \epsilon \tilde{\phi}_{x}(x,\epsilon\tilde{\eta}(x,t),t)
\]
where the potential $\tilde{\phi}_{x}$ is found using Equation \eqref{phixexp}.  However, since the NLS solutions are in terms of the coordinates $\xi$ and $\tau$, we rewrite the above equation in these coordinates so that we have
\begin{equation}
\frac{d \xi}{d\tau} =  \frac{c_{g}}{\epsilon} + \omega\tilde{\eta}\left(\xi,\tau\right) + \tilde{\phi}_{x}(\xi,\epsilon\tilde{\eta}),
\label{pathode}
\end{equation}
\[
\xi(0)=\xi_0.
\]
We use a fourth-order Runge-Kutta (RK4) scheme with a time step of $\delta \tau = 5\times10^{-4}$ to solve this initial-value problem.  To solve for the evolution of $\eta_{1}(\xi,\tau)$, we use a pseudo-spectral scheme with $1024$ modes to discretize the VD Equation in space thereby allowing it to be coupled to the RK4 scheme used in Equation \eqref{pathode}. We can then determine $\eta_{1}(\xi,\tau)$ up to and including $\mathcal{O}(\epsilon)$ effects on a $\mathcal{O}(1/\epsilon^{2})$ timescale.  We use the Jacobi elliptic or plane wave solutions to the NLS equation as an initial condition to the VD equation.  By using the VD equation, we see that on the $\mathcal{O}(1)$ in $\tau$, or $\mathcal{O}(1/\epsilon^{2})$ in $t$, time scale of the NLS equation, we anticipate that the error from truncation in Equation \eqref{pathode} should be at worst $\mathcal{O}(\epsilon^{2})$ in $\xi$ and thus at worst $\mathcal{O}(\epsilon)$ in $x(t)$.  Per our restriction on $z(t)$, this implies that at worst we should anticipate errors of $\mathcal{O}(\epsilon^{2})$ on the longer time scales corresponding to the nonlinear effects.

For the Jacobi elliptic function solutions, we chose the initial tracer position to be
\[
\xi(0) = \frac{K(\kappa)}{128}, ~ z(0) = \epsilon \tilde{\eta}(\xi(0),0),
\]
where $K(\kappa)$ is the complete elliptic integral of the first kind and $\tilde{\eta}$ is defined by Equation \eqref{nlssurfexp}.  For the plane-wave solutions, we chose $\xi_{0}=10/128$ so the initial positions are comparable across the different classes of solutions.  We take the positive branch of the dispersion relationship, which implies that $k_{0}>0$ corresponds to a left-moving-carrier wave.  We choose $k_{0}=1$ for the plane-wave solutions.  In the case of the Jacobi elliptic solutions to the NLS equation, we choose the integer $m$ to be
\[
m = \left\lfloor \frac{2K(\kappa)}{\pi \epsilon}\right\rfloor,
\]
so that, using Equation \eqref{jacwavnum}, $k_{0}\approx 1$.  In practice, $k_{0}\approx0.98$ or $0.99$.  The results for the tracer position in physical coordinates, $(x(t),z(t))$, are found using the transformations
\[
x(t) = \frac{\xi(\tau)}{\epsilon}-\frac{c_{g}\tau}{\epsilon^{2}}, ~~~~ t = \frac{\tau}{\epsilon^{2}}.
\]
Fourier transforms and their inverses are computed numerically using the fast Fourier transform.

Finally, the expansions in \eqref{phixexp} and \eqref{phizexp}, and the associated approximations in \eqref{roeqx} and \eqref{roeqz}, show that the strength of the velocity field is largely determined by the magnitude of the solution to the NLS equation which is the leading order approximation to $\eta_{1}$.  In the case of the Jacobi elliptic function solutions, controlling for the elliptic modulus $\kappa$, the term $\sqrt{2|\alpha_{d}/\alpha_{nl}|}$ is the most significant contribution to the magnitude of $\eta_{1}$ since we have fixed $\beta=1$; see Equations \eqref{cnsolns} and \eqref{snsolns}.  Figure \ref{fig:ampcomps} contains plots of $\sqrt{2|\alpha_{d}/\alpha_{nl}|}$ versus $\omega$ for $k_0=1$.  Note, we take $\omega = \mathcal{O}(1)$ to be consistent with our derivation of the VD equation.
\begin{figure}
\centering
\begin{tabular}{ccc}
\includegraphics[width=.32\textwidth]{Figure_5a} & \includegraphics[width=.32\textwidth]{Figure_5b} & \includegraphics[width=.32\textwidth]{Figure_5c}\\
(a) & (b) & (c)
\end{tabular}
\caption{\small Strength of the surface velocity as measured via plots of $\sqrt{2|\alpha_{d}/\alpha_{nl}|}$ for $k_{0}=1$ for $-1\leq \omega \leq 4$ (a), $-1\leq \omega \leq 1.12$ (b), and $1.17\leq \omega \leq 4$ (c).  Increasing shear increases the velocity on the focusing side, and decreases it on the defocusing side.}
\label{fig:ampcomps}
\end{figure}
In the focusing case, increasing the magnitude of the Jacobi elliptic function solutions corresponds to increasing the shear strength.  In the defocusing case, decreasing the magnitude of the Jacobi elliptic function solutions corresponds to increasing the shear strength.  These two statements are justified by the similarities of Figures \ref{fig:ampcomps} (b) and (c) and \ref{fig:stksdrfcomp} (b) and (c).  Therefore, relatively large amplitude solutions inducing strong fluid particle drift is to be expected near the resonance curve.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Focusing Case}

Using the solution given in equation \eqref{cnsolns} with $k_0\approx1$ ensures that $\omega = 0$ and $\omega=\pm 1$ are in the focusing regime.   Figure \ref{fig:foc_kap_pt5} shows the paths of particles as well as there mean paths for three values of $\omega$ found by setting $\epsilon=0.1$, $\kappa=0.5$, and solving the dynamical system for the particle and mean paths up to time $t=1/\epsilon^{2}$.  The carrier profile is propagating to the left in this situation.  Therefore, if $\omega=1.12$, then the shear current is counter-propagating at the surface with respect to the carrier wave while if $\omega=-1$, the current is co-propagating.  The counter-propagating current significantly enhances the leftward horizontal motion of a tracer along the surface as seen by comparing Figures \ref{fig:foc_kap_pt5}(b) and (c) with (a).  This can be explained by comparing the values of the SDV and the LDV parameters
\[
\begin{array}{lcl}
\tilde{u}^{S}(1,0) = -0.5222, & & \tilde{u}^{L}(1,0) = -0.5222,\\
\tilde{u}^{S}(1,1.12) = -10.7373,& & \tilde{u}^{L}(1,1.12) = -5.2195,\\
\tilde{u}^{S}(1,-1) = -0.2836, & & \tilde{u}^{L}(1,-1) = -0.1636,
\end{array}
\]
where the parameter $\tilde{u}^{L}$ is defined to be
\[
\tilde{u}^{L}(k_{0},\omega) = 2\left|\frac{\alpha_{d}(k_{0},\omega)}{\alpha_{nl}(k_{0},\omega)}\right|u^{L}_{p}(k_{0},\omega).
\]
Note that for the Jacobi elliptic solutions and controlling for $\kappa$ and $\beta$, this will be the most significant contribution to the magnitude of the LDV.  This demonstrates why the counter-propagating current so enhances the leftward drift since the LDV parameter is an order of magnitude larger than in the other cases.    Further, it demonstrates why the co-propagating current, i.e. $\omega=-1$, reduces the horizontal displacement of the surface particle.  This is somewhat surprising as one might intuitively imagine that the co-propagating shear would enhance the particle drift, especially in comparison to the zero vorticity case.  However, we have effectively shown that nonlinearity makes the surface/current interaction a more complicated one than one might at first suspect.
\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=.48\textwidth]{Figure_6a} & \includegraphics[width=.48\textwidth]{Figure_6b} \\
(a) & (b) \\
\includegraphics[width=.48\textwidth]{Figure_6c} & \\
(c) &
\end{tabular}
\caption{\small {\bf Focusing Case} - Plots of particle paths correspond to the solution given in Equation (\ref{cnsolns}) with $\epsilon=0.1$, $\beta=1$, $k_{0}\approx 1$, $\kappa=0.5$, with $\omega=0$ (a), $\omega=1.12$ (b), and $\omega=-1$ (c). The grey dots indicate the starting positions of the tracers while the black dots indicate their final positions.}
\label{fig:foc_kap_pt5}
\end{figure}

Choosing $\kappa=0.99$ shows that similar results hold closer to the solitary wave solution; see Figure \ref{fig:foc_kap_pt99}, though the larger elliptic modulus results in the larger amplitudes and leftward drifts.  In particular, near the solitary wave limit, positive, counter-propagating shear can greatly enhance both the impact of nonlinearity and the transport properties of the waves.
\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=.48\textwidth]{Figure_7a} & \includegraphics[width=.48\textwidth]{Figure_7b} \\
(a) & (b) \\
\includegraphics[width=.48\textwidth]{Figure_7c} & \\
(c) &
\end{tabular}
\caption{\small {\bf Focusing Case} - Plots of particle paths correspond to the solution given in Equation (\ref{cnsolns}) with $\epsilon=0.1$, $\beta=1$, $k_{0}\approx 1$, $\kappa=0.99$, with $\omega=0$ (a), $\omega=1.12$ (b), and $\omega=-1$ (c). The grey dot indicates the starting position of the tracer while the black dot indicates the final position.}
\label{fig:foc_kap_pt99}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Defocusing Case}

For the Jacobi elliptic solutions with $k_{0}\approx 1$, the zero LDV solutions belong in the defocusing case.  Figure \ref{fig:jaczerodrift} shows the impact on surface flow particle paths when $\omega$ is chosen to zero out the LDV.
\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=.48\textwidth]{Figure_8a} & \includegraphics[width=.48\textwidth]{Figure_8b} \\
(a) $\omega=1.7027$ & (b) $\omega=1.6672$
\end{tabular}
\caption{\small {\bf Defocusing Case} - Plots of particle paths correspond to the solution given in Equation (\ref{snsolns}) with $\epsilon=0.1$, $\beta=1$, $k_{0}\approx 1$, with $\kappa=0.5$  (a), $\kappa=0.99$ (b). The choices of $\omega$ ensure that $\tilde{u}^{L}(1,\omega)=0$, see Figure \ref{fig:zerodriftk0}, thereby leading to the nearly closed, elliptical particle paths.  The grey dots indicate the starting positions of the tracers while the black dots indicate their final positions.  }
\label{fig:jaczerodrift}
\end{figure}
As can be seen, the particle paths, while not exactly closed, are spirals and are constrained in their horizontal and vertical extent by an outer elliptical perimeter.  Similarly, the shear strength can be chosen so that the LDV is zeroed out for the plane-wave solutions.  Taking $k_{0}=1$ and $A=1$, this corresponds to choosing $\omega = 1.6828$.  Figure \ref{fig:pwavezdrift} shows a nearly closed particle path.  We emphasize that these results are a confirmation of the predictions made in Figure \ref{fig:zerodriftk0}.  Thus, these numerical results validate our choice to generalize the GLM approach in so far as we see the generalization provides the correct predictions for when shear will quench surface drift.
\begin{figure}
\centering
\includegraphics[width=.48\textwidth]{Figure_9}
\caption{\small {\bf Defocusing Case} - Plot of a particle path corresponding to a plane-wave solution with $\epsilon=0.1$, $A=1$, $k_{0}=1$ , and $\omega=1.6818$, which ensures that $\tilde{u}^{L}(1,\omega)=0$, see Figure \ref{fig:zerodriftk0}, thereby leading to the nearly closed, elliptical particle paths.  The grey dots indicate the starting positions of the tracers while the black dots indicate their final positions.  }
\label{fig:pwavezdrift}
\end{figure}

In some respects then, we have also shown that by appropriately choosing the shear, we can replicate the dynamics of a Gerstner wave, see \cite{constantin} by looking instead at a plane-wave solution moving over a counter-propagating shear current.  This could point towards resolving some of the questions raised in \cite{monismith,smith}, though this is a subject of future research.

Looking beyond cases in which Eulerian counterflows quench drift, we now consider the defocusing NLS Jacobi elliptic function solutions given in Equation \eqref{snsolns}.  We choose $k_{0}\approx1$, $\epsilon=0.1$, $\kappa=0.99$ (near the dark-soliton limit).  Solving the dynamical system for the particle paths up to time $t=1/\epsilon^{2}$ generates Figure \ref{fig:defoc_kap_pt99} (a) and (b) for $\omega= 1.17$ and $\omega= 4$ respectively.  While the particle path motion seen in Figure \ref{fig:defoc_kap_pt99}(a) is rapidly oscillating, the net transport and amplitude is quite small.
\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=.48\textwidth]{Figure_10a} &\includegraphics[width=.48\textwidth]{Figure_10b} \\
(a) $\omega = 1.17$ & (b) $\omega=4$
\end{tabular}
\caption{\small {\bf Defocusing} - Near dark soliton solution for $k_{0}\approx1$, $\kappa=.99$ and $\omega=1.17$ (a) and $\omega=4$ (b).  The grey dot indicates the starting position of the tracer while the black dot indicates the final position.}
\label{fig:defoc_kap_pt99}
\end{figure}

However, if $\omega=4$ in the case of a plane wave, we do not necessarily get the same diminished response seen for the Jacobi elliptic solutions; see Figure \ref{fig:defoc_pwave}.  The highly oscillatory plane-wave profile induces a particle path which ultimately exhibits a strong rightward drift which follows the strong counter-propagating shear current.   The associated SDV and LDV at the surface for the plane wave with these parameters are given by
\begin{align*}
\bar{u}^{S}_{s} = & -8.4985\epsilon^{2} + \mathcal{O}(\epsilon^{3}),\\
\bar{u}^{L}_{s} = & 29.2719\epsilon^{2} + \mathcal{O}(\epsilon^{3}).
\end{align*}
This explains the stronger net rightward drift of the particle shown in Figure \ref{fig:defoc_pwave}.  Finally, note that this shows the SDV and the LDV can oppose one another, thereby clarifying the need to use both velocities to fully understand the drift properties associated with a surface wave.
 \begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=.5\textwidth]{Figure_11a} & \includegraphics[width=.5\textwidth]{Figure_11b}\\
(a) & (b)
\end{tabular}
\caption{\small {\bf Defocusing Case} - Plane-wave solution for $k_{0}=1$, $\omega=4$, $A=1$, with (b) providing a detail of (a).  The grey dot indicates the starting position of the tracer while the black dot indicates the final position.}
\label{fig:defoc_pwave}
\end{figure}

\section{Conclusion}

In this paper, to better understand the transport properties of nonlinear waves moving over constant-vorticity currents, we have derived an asymptotically self-consistent higher-order model for the interaction between the mean wave height and the leading order modulated carrier wave in the presence of surface tension and a constant vorticity current.  Using this, an NLS equation is derived, and the role of surface tension in determining the existence of modulational instability is fully explored.  A formula for the Stokes drift velocity at the surface is derived allowing for the identification of shear profiles which enhance or quench horizontal surface transport.  Numerical simulations corroborate the surface theory developed in this paper.

The work in this paper provides possible explanatory mechanisms for otherwise unexplained oceanic phenomena observed
in \cite{smith} and justification for phenomenological choices in oceanographic modeling put forward in \cite{breivik}.  Assessing how accurate a constant-vorticity-modulational-nonlinear wave model describes these real world problems is an important future research direction.  Likewise, while a higher-order model describing the coupled evolution of the mean and modulated carrier wave is derived, the properties of this equation are not studied directly.

\bigskip

This material is based upon work supported by the National Science Foundation under Grant No. DMS-1439786 while CWC and JDC were in residence at the Institute for Computational and Experimental Research in Mathematics in Providence, RI, during the Spring 2017 semester.  Additionally, JDC acknowledges and thanks the Norwegian Fulbright Core Scholar Program for supporting his five-month long visit to the University of Bergen to work with HK. This research was also supported by
the Research Council of Norway through grants 213474/F20 and 239033/F20.  We also express a great gratitude towards the anonymous reviewers whose comments and questions helped this become a far stronger paper.

\bibliographystyle{JFM_Style/jfm}
\bibliography{deep_water_shear_July30}

\end{document}

### 2020-01-09 Writeup/random_vorticity_revision_1_5_20.tex
All the updates

--- 2019-12-19 Writeup/random_vorticity.tex
@@
 As explained in \cite{janssen}, MI is especially important due to the relative rapidity with which it acts, being orders of magnitude faster than say four-wave resonant interactions which form the traditional backbone of our understanding of how nonlinearity drives changes in statistics of random-wave fields \cite{hasselmann,holthuijsen}.
-These results have significant impact on the understanding of deep-water-rogue wave formation among other oceanographic phenomena.
+A detailed set of theoretical and experimental results \cite{sergeeva1,sergeeva2,sergeeva3,sergeeva4} have also shown that MI could play a significant role in unidirectional deep-water-rogue wave formation among other oceanographic phenomena.
+It is important to note though that there are several possible mechanisms for rogue-wave generation; see \cite{fedele} for example which rules out MI as a mechanism altogether.
 
 However, much of the understanding around MI relies on highly idealized assumptions which cannot be expected to hold in natural settings.
-In particular, most understandings of MI rely on looking at the stability of perturbations to carrier waves at essentially a fixed wave number.
+In particular, most analytic understandings of MI rely on looking at the stability of perturbations to carrier waves at essentially a fixed wave number.
 To address this shortcoming, in now seminal papers, \cite{alber} and \cite{alber2} analytically studied families of perturbations of wave packets with narrow, but non-zero, spectral width around a central wave number.
@@
 As shown in \cite{onorato,janssen} and \cite{eeltink}, the higher-order terms associated with the Dysthe equation can have significant impacts on the statistics of the waves, so it is a nontrivial question to examine how the vorticity interacts with these higher-order nonlinearities.
+Moreover, as shown in \cite{sergeeva2}, the NLSE can rapidly lose any quantitative predictive power in experimental settings while the Dysthe equation provides far more accurate results over a relatively wide range of experimental conditions.
 
 Using our simulations, starting from normally distributed random-wave fields, we show that the BFI condition with respect to predicting stability is not especially useful, conforming to the results in \cite{janssen}.
-While we are able to show that for sufficiently wide enough initial distributions that the temporal modifications to the profile are slight, whether or not this corresponds to an actual bifurcation in stability or simply reflects that for a wide enough profile the sideband growth affiliated with the MI has little impact is not clear.
+While we are able to show that for sufficiently wide enough initial distributions that the temporal modifications to the profile are slight, which echoes the zero-vorticity results in \cite{sergeeva1,sergeeva2,sergeeva3}, whether or not this corresponds to an actual bifurcation in stability or simply reflects that for a wide enough profile the sideband growth affiliated with the MI has little impact is not clear.
 We note though that classic MI analysis does provide a useful prediction of the time scale over which the most significant dynamics in the statistics of the wave field occur.
@@
 
-The structure of the paper is an explanation of the model and discussion of the BFI in Section 2.  Section 3 presents our numerical results, while Section 4 presents a conclusion.
+The structure of the paper is an explanation of the model and discussion of the BFI in Section 2.
+Section 3 presents our numerical results, while Section 4 presents a conclusion.
+Explorations of two significant technical points are given in the Appendix.
 
@@
 In physical terms, this implies that $\omega$ is comparable to the natural time scale of this problem, $\sqrt{L_{p}/g}$.
-If we were to assume $\omega$ were of larger magnitude, the problem would no longer be weakly nonlinear and asymptotic analysis would be siginificantly more difficult.
+Letting $L_{p}\sim 1~m$, then this corresponds to the horizontal velocity of the fluid $u \sim 300 ~cm/s$ at a depth of one meter below the fluid surface.
+Thus we are modeling the impact of strong surface currents moving over deep water.
+If we were to assume $\omega$ were of larger magnitude, the problem would no longer be weakly nonlinear and asymptotic analysis would be significantly more difficult.
 Therefore, throughout the remainder of the paper, we assume that the vorticity is not too large.
+We also note that as $z\rightarrow -\infty$, the horizontal fluid velocity becomes unbounded in this model, which is physically absurd.
+While we are only interested in results near the surface in this work, in the Appendix, we explore this issue and show that a more physically realistic shear profile would produce similar results at the surface to those presented throughout the remainder of the paper.
 
@@
 \[
-\Omega(k_{0},\omega) = \frac{1}{2}\left(s\omega \pm \sqrt{\omega^{2} + 4|k_{0}|\left(1+\tilde{\sigma}k_{0}^{2}\right)} \right).
+\Omega(k_{0},\omega) = \frac{1}{2}\left(s\omega \pm \sqrt{\omega^{2} + 4|k_{0}|\left(1+\tilde{\sigma}_{s}k_{0}^{2}\right)} \right).
 \]
 We note that the form for $\phi$ follows directly from this expansion, allowing for everything to be modeled by the terms in the expansion in $\eta$.
-Here, $s = \mbox{sgn}(k_{0})$, $\omega$ is the non-dimensionalized magnitude of the vorticity of the flow, and $\tilde{\sigma}$ is the surface tension, or where the reciprocal of the Bond number, $\tilde{\sigma}$, is given by
+Here, $s = \mbox{sgn}(k_{0})$, $\omega$ is the non-dimensionalized magnitude of the vorticity of the flow, and $\tilde{\sigma}_{s}$ is the surface tension, or where the reciprocal of the Bond number, $\tilde{\sigma}$, is given by
 \[
-\tilde{\sigma} = \frac{\sigma}{\rho g L_{p}^2}.
+\tilde{\sigma}_{s} = \frac{\sigma}{\rho g L_{p}^2}.
 \]
@@
 The first model is the the nonlinear Schr\"{o}dinger equation (NLSE) derived in \cite{curtis8}, though see also \cite{thomas2012nonlinear}, given by
-\[
+\begin{equation}
 i\pd_{\tau}\eta_{1} + \alpha_{d}\pd_{\xi}^{2}\eta_{1} + \alpha_{nl}\left|\eta_{1} \right|^{2}\eta_{1} = 0, 
-\]
+\label{nlse}
+\end{equation}
 where 
 \begin{align*}
-c_{g} = & \frac{1+3\tilde{\sigma}k_{0}^{2}}{2s\Omega - \omega},\\
-\alpha_{d}(k_{0},\omega) = & \frac{(c^2_{g} - 3|k_{0}|\tilde{\sigma})}{2\Omega-s\omega},\\
-\alpha_{nl}(k_{0},\omega) = & \frac{k_{0}\left( sk_{0}^{3}\left(8 + \tilde{\sigma}k_{0}^{2} + 2(\tilde{\sigma}k_{0}^{2})^{2}\right) + \omega \alpha_{v}\right)}{\left(2s\Omega -\omega\right)(1+c_{g}\omega)\left(4\Omega^2-s(2k_{0}(1+4\tilde{\sigma}k_{0}^{2})+2\omega\Omega)\right)}.
+c_{g}(k_{0},\omega) = & \frac{1+3\tilde{\sigma}_{s}k_{0}^{2}}{2s\Omega - \omega},\\
+\alpha_{d}(k_{0},\omega) = & \frac{(c^2_{g} - 3|k_{0}|\tilde{\sigma}_{s})}{2\Omega-s\omega},\\
+\alpha_{nl}(k_{0},\omega) = & \frac{k_{0}\left( sk_{0}^{3}\left(8 + \tilde{\sigma}_{s}k_{0}^{2} + 2(\tilde{\sigma}_{s}k_{0}^{2})^{2}\right) + \omega \alpha_{v}\right)}{\left(2s\Omega -\omega\right)(1+c_{g}\omega)\left(4\Omega^2-s(2k_{0}(1+4\tilde{\sigma}_{s}k_{0}^{2})+2\omega\Omega)\right)},\\
+\alpha_{v}(k_{0},\omega) = & s(c_g k_0  - 2\Omega ) \omega^4 + k_{0}(4 k_{0}^{2} s \tilde{\sigma}_{s} + 2 \Omega c_g  -s) \omega^3.
 \end{align*}
-However, the NLSE is a relatively low-order model known to be only of limited use in real-world modeling situations.
+However, the NLSE is a relatively low-order model known to be only of limited use in real-world modeling situations; see \cite{sergeeva2} in particular.
 A more robust model, following the work in \cite{dysthe1}, is found by going to the next higher order in $\epsilon$, giving us the equation
 \begin{multline}
-(\omega-2s\Omega)\pd_{\tau}\eta_{1} + i(c_{g}^{2}s-3k_{0}\tilde{\sigma})\pd_{\xi}^{2}\eta_{1} + 2i\epsilon s c_{g}\pd^{2}_{\xi\tau}\eta_{1} - \epsilon\tilde{\sigma} \pd^{3}_{\xi}\eta_{1} + i\epsilon^{2}c_{h}\pd_{\xi}^{4}\eta_{1} \\
+(\omega-2s\Omega)\pd_{\tau}\eta_{1} + i(c_{g}^{2}s-3k_{0}\tilde{\sigma}_{s})\pd_{\xi}^{2}\eta_{1} + 2i\epsilon s c_{g}\pd^{2}_{\xi\tau}\eta_{1} - \epsilon\tilde{\sigma}_{s} \pd^{3}_{\xi}\eta_{1} + i\epsilon^{2}c_{h}\pd_{\xi}^{4}\eta_{1} \\
 - i \alpha_{nl}(\omega-2s\Omega)\left|\eta_{1}\right|^{2}\eta_{1}
@@
 \[
-c_{h} = 2sc_{g}(2sc_{g}\alpha_{d}+\tilde{\sigma})/(\omega-2s\Omega) - s\alpha_{d}^{2},
+c_{h} = 2sc_{g}(2sc_{g}\alpha_{d}+\tilde{\sigma}_{s})/(\omega-2s\Omega) - s\alpha_{d}^{2},
 \]
-was necessary to remove artificial numerical instabilities.
-Similar issues with Dysthe like equations have been noted in \cite{eeltink} among other sources.
+was necessary to remove artificial instabilities.
+Similar issues with Dysthe like equations have been noted in \cite{eeltink} among other sources; see the Appendix for a discussion of this issue.
 Moreover, the coefficients $\tilde{\alpha}_{j}$ are readily derived, but their forms are cumbersome and thus we omit explicitly writing them for the sake of readability.
 
-To generate simulations, we start both the NLSE and VDE from the same collection of $2L$ periodic-initial conditions
+The traditional definition of modulational instability (MI) begins by studying the stability of the plane-wave, or Stokes wave, solutions of the NLS equation given by
+\[
+\eta_{1}(\xi,\tau) = A e^{i\alpha_{nl}A^{2}\tau},
+\]
+where $A>0$ is a real constant.  Complementing the results in \cite{thomas2012nonlinear}, we study the stability of these solutions by considering perturbed solutions of the form
+\begin{equation}
+\eta_{1p}(\xi,\tau) = \Big{(}A +\mu \big{(}u_{p}(\xi,\tau)+iv_{p}(\xi,\tau)\big{)} +\mathcal{O}(\mu^2)\Big{)}\mbox{e}^{i\alpha_{nl}A^{2}\tau},
+\label{pert}
+\end{equation}
+where $\mu$ is a small real parameter and $u_{p}$ and $v_{p}$ are real-valued functions.  Substituting Equation \eqref{pert} into the NLSE given by Equation \eqref{nlse} and linearizing gives
+\[
+\pd_{\tau}\bp u_{p}\\ v_{p}\ep = \bp 0 & -\alpha_{d}\pd_{\xi}^{2} \\ \alpha_{d}\pd_{\xi}^{2} + 2A^{2}\alpha_{nl} & 0 \ep \bp u_{p} \\ v_{p} \ep.
+\]
+Separating variables, applying a Fourier transform in $\xi$, and introducing the $\xi$ wave number $l$, establishes that NLS plane-wave solutions are unstable with respect to MIs if
+\[
+0 < l^{2} \leq 2\frac{\alpha_{nl}}{\alpha_{d}}A^{2}.
+\]
+This reduces to the classic requirement that MIs are suppressed if the coefficients of dispersion and nonlinearity have opposite signs (i.e.~the defocusing case).
+Thus in the Results section we make choices of $\omega$ which keep us in the focusing case.
+See \cite{curtis8} for further details.
+
+As noted though in \cite{alber}, this notion of MI is only defined for perturbations around a monochromatic wave train.
+We are thus motivated to examine perturbations around more complicated wavetrains.
+To generate these simulations, we start both the NLSE and VDE from the same collection of $2L$ periodic-initial conditions
 \[
@@
 Note, this definition is identical with the spectral width criterion of \cite{alber} and the BFI condition given in \cite{thomas2012nonlinear}; see also \cite{onorato}.
-It was claimed in \cite{alber}, and in some sense verified in \cite{dysthe2}, that wave packets should be stable with respect to MI's for $\mbox{BFI}<1$, which corresponds to the initial spectral width $\sigma>\sigma_{c}$, where
+It was claimed in \cite{alber}, and in some sense verified in \cite{dysthe2}, that wave packets should be stable with respect to MIs for $\mbox{BFI}<1$, which corresponds to the initial spectral width $\sigma>\sigma_{c}$, where
 \[
@@
 The number of modes is $K_{T} = 1024$ or $K=512$.
-The surface tension is $\tilde{\sigma} = 10^{-5}$.
+Letting the length scale $L_{p}=1m$, the non-dimensional surface tension is $\tilde{\sigma}_{s} = 10^{-5}$.
+It is thus ignored throughout the remainder of the paper.
 We then compute the averaged spectral density $S(k,\tau;\sigma)$ where
@@
 Comparing the VDE to the NLSE, we see the added nonlinearity in the VDE allows for yet greater deviations away from normality through the downshifting of the spectral peak; see Figure \ref{fig:stbleom0} (a), and the asymmetric widening of the VDE profile as evidenced by the enhanced kurtosis and strong shift in the mean as seen in Figures \ref{fig:stbleom0} (d) and (b) respectively.
+Note, that the peak shifts left but the mean shifts right indicates the strong degree of skewness ,or asymmetry, in the distribution the VDE induces.
+This effect is seen to varying degrees in all subsequent plots.
 
@@
 This far longer time scale is seen in Figure \ref{fig:stbleom1}, where the saturation of the statistics happens just before $\tau_{f}$.
+Note, we also studied results on yet far longer time scales where $t_{f}=60$ and the results obtained were essentially the same as those presented.
 For the narrow initial spectral width $\sigma = 1.05\sigma_{c}$, we see some of the most pronounced deviations between the VDE and NLSE.
@@
 Thus the VDE markedly enhances far-tail phenomena and provides strikingly different predictions than its counterpart model the NLSE.
-Thus we see that the the co-propagating shear likely enhances the occurrence of freak wave or other far-tail phenomena for relatively narrow spectral profiles
+We see then that the the co-propagating shear likely enhances the occurrence of freak wave or other far-tail phenomena for relatively narrow spectral profiles.
 
@@
 Moreover, increasing shear enhances the kurtosis, and finally the higher-order nonlinearities of the VDE always enhance the kurtosis relative to the value predicted by the NLSE.
-Likewise, we see that the quadratic fits are clearly good, though also it is clear varying vorticity values change the coefficients of the fit.
+They also in the presence of stronger shear can cause a greater deviation away from the quadratic fit; see Figure \ref{fig:kurtplt} (c).
+Likewise, we see that the quadratic fits are relatively good, though also it is clear varying vorticity values change the coefficients of the fit.
 This conforms with the results in \cite{eeltink} which show that the quadratic relationship between the kurtosis and the BFI varies depending on different damping and wind profiles.
@@
 
-\section*{Acknowledgements}
+\section*{Appendix}
+\subsection*{Shear Profiles over Infinitely Deep Fluids}
+As noted in the text, the assumption that the fluid velocity is to leading order given by 
+\[
+{\bf u} \approx \omega z \hat{{\bf i}}
+\]
+is clearly unrealistic insofar as it leads to currents of infinite speed as one descends through the fluid.
+A more realistic, though also more complicated, ansatz is to suppose that to leading order we have that ${\bf u}\approx u(z)\hat{{\bf i}}$ where
+\[
+u(z) = \left\{\ba{rl} \omega z, & -h_{0} \leq z < 0 \\
+\frac{-\omega h_{0}}{h_{1}-h_{0}} (z+h_{1}), & -h_{1} \leq z < -h_{0} \\
+0, & z < -h_{1}
+\ea \right.
+\] 
+where $h_{1}>h_{0}\gg 1$, so that we are looking at a deep, continuous shear profile which is zero at or near the surface $z=\epsilon \eta(x,t)$ and at the depth $z=-h_{1}$, after which the fluid is to leading order quiescent.
+Note, we can also see this profile as satisfying to leading order two `no-slip' conditions, one near the free surface at $z=0$ and one near $z=-h_{1}$.
+
+While a full, nonlinear description of the above shear profile would require two more freely evolving interfaces, and thus is beyond the scope of this paper, we can readily find the dispersion relationship affiliated with this profile.
+This then allows us to provide some analytic argument for why we study an otherwise unphysical velocity profile in the main body of the text.
+Likewise, information from the dispersion relationship provides us with a better understanding of how depth varying shear profiles induce both surface and internal waves.
+
+Following relatively classical approaches, we introduce three fluid velocities
+\begin{align*}
+{\bf u}_{1} = \omega z \hat{{\bf i}} + \epsilon \nabla \phi_{1}, &  ~-h_{0}+\epsilon \eta_{2} < z < \epsilon \eta_{1}\\
+{\bf u}_{2} = \frac{-\omega h_{0}}{h_{1}-h_{0}} (z+h_{1}) \hat{{\bf i}} + \epsilon \nabla \phi_{2}, & ~-h_{1}+\epsilon \eta_{3} < z < -h_{0}+\epsilon \eta_{2}\\
+{\bf u}_{3} = \epsilon \nabla \phi_{3}, & ~z < -h_{1} + \epsilon \eta_{3}
+\end{align*}    
+so that after linearizing around the small disturbances, we have the following system of evolution equations describing the behavior of small disturbances
+\[
+\pd_{t}\phi_{1} + g\eta_{1} + \omega \pd_{t}\pd_{x}^{-1}\eta_{1} = 0, ~ z = 0 
+\]
+\[
+\pd_{t}\eta_{1} = \pd_{z}\phi_{1}, ~ z = 0, 
+\]
+\[
+\pd_{t}\phi_{2} - \frac{\omega h_{1}}{h_{1}-h_{0}} \pd_{t}\pd_{x}^{-1}\eta_{2} = \pd_{t}\phi_{1} , ~ z = -h_{0} 
+\]
+\begin{align*}
+\pd_{t}\eta_{2} = \pd_{z}\phi_{2}, & ~ z = -h_{0},\\
+\pd_{t}\eta_{2} = \pd_{z}\phi_{1}, & ~ z = -h_{0}, 
+\end{align*}
+and
+\[
+\pd_{t}\phi_{3} = \pd_{t}\phi_{2} - \frac{\omega h_{0}}{h_{1}-h_{0}} \pd_{t}\pd_{x}^{-1}\eta_{3}, ~ z = -h_{1} 
+\]
+\begin{align*}
+\pd_{t}\eta_{3} = \pd_{z}\phi_{3}, & ~ z = -h_{1},\\
+\pd_{t}\eta_{3} = \pd_{z}\phi_{2}, & ~ z = -h_{1}. 
+\end{align*}
+Taking each function $\phi_{j}$ to be harmonic and letting $\pd_{z}\phi_{3}\rightarrow 0 $ as $z\rightarrow -\infty$ gives us the following solutions for each of the scalar velocity potentials
+\[
+\phi_{1}(x,z,t) = \left(\alpha_{11}\cosh(kz)+\alpha_{12}\sinh(kz)\right)e^{i\theta(x,t)} + \mbox{cc}
+\]
+\[
+\phi_{2}(x,z,t) = \left(\alpha_{21}\cosh(k(z+h_{0}))+\alpha_{22}\sinh(k(z+h_{0}))\right)e^{i\theta(x,t)} + \mbox{cc}
+\]
+and
+\[
+\phi_{3}(x,z,t) = \alpha_{3}e^{|k|(z+h_{1})}e^{i\theta(x,t)} + \mbox{cc}
+\]
+where $\theta(x,t) = kx + \Omega(k,\omega)t$, and `cc' denotes the complex conjugate.
+
+This then gives us the linear system of equations 
+\begin{align*}
+\Omega \alpha_{11} = & \left(\frac{gk}{\Omega} + \omega \right)\alpha_{12} \\
+\Omega\left(\alpha_{11}\cosh(kh_{0}) -\alpha_{12}\sinh(kh_{0}) \right) = &  \Omega \alpha_{21} + \frac{\omega h_{1}}{\delta h}\alpha_{22}\\
+\Omega\left(\alpha_{21}\cosh(k\delta h) - \alpha_{22}\sinh(k \delta h) \right) = & \left(\Omega - \frac{\omega h_{0} s}{\delta h} \right)\alpha_{3}\\
+\alpha_{22} = & -\alpha_{11}\sinh(kh_{0}) + \alpha_{12}\cosh(kh_{0})  \\
+\alpha_{3} = & s_{k}\left(-\alpha_{21}\sinh(k\delta h) + \alpha_{22}\cosh(k\delta h)\right)
+\end{align*}
+where $\delta h = h_{1}-h_{0}$ and $s=\mbox{sgn}(k)$.  From this we derive the dispersion relationship 
+\[
+\frac{gk}{\Omega} + \omega -\Omega \tanh(kh_{0}) + \left(\frac{1}{\Omega}\left(\frac{gk}{\Omega} + \omega\right)\tanh(kh_{0}) - 1\right)\left(\frac{\omega h_{1}}{\delta h} + \Omega \Omega_{1}(k\delta h)\right) = 0
+\]
+where
+\[
+\Omega_{1}(k\delta h) = \frac{s\left(\Omega -\frac{\omega h_{0}}{\delta h}s\right)+\Omega \tanh(k\delta h)}{\Omega +s\left(\Omega - \frac{\omega h_{0}}{\delta h}s\right)\tanh(k\delta h)}.
+\]
+While in general we would have to find the roots of a fifth-order polynomial to determine the values of $\Omega$, we see for $h_{1}\gg h_{0}$ that $\Omega_{1}\sim s$, and thus we get that the dispersion relationship simplifies to
+\[
+\frac{gk}{\Omega} + \omega -\Omega \tanh(kh_{0}) + \left(\frac{1}{\Omega}\left(\frac{gk}{\Omega} + \omega\right)\tanh(kh_{0}) - 1\right)\left(\tilde{\omega} + s\Omega \right) \sim 0,
+\]
+where 
+\[
+\tilde{\omega} = \frac{\omega h_{1}}{\delta h}.
+\]
+Since $\tilde{\omega}$ only approaches $\omega$ at an algebraic rate, it seems appropriate to keep it included in the analysis.  Letting 
+\[
+\left|\tanh(kh_{0}) \right| = 1 - \tilde{\epsilon}, 
+\]
+and noting that $\tilde{\epsilon}$ vanishes to zero exponentially fast as we increase $h_{0}$, we see the reduced dispersion relationship factors into the form
+\[
+\left(2\Omega + s\tilde{\omega} \right)\left(g\left|k\right| + s\omega \Omega - \Omega^{2} \right) + \tilde{\epsilon}\left(\Omega^{3} - (s\tilde{\omega} + \Omega)(g|k|+\omega s\Omega) \right) \sim 0.
+\]
+Thus we have a completely regular perturbation problem for the roots, which we readily see are given by
+\[
+\Omega \sim \frac{1}{2}\left(-s\omega \pm \sqrt{\omega^{2}+4g|k|} \right), ~ -\frac{s\tilde{\omega}}{2}.
+\]
+Thus of the three roots we find, two give the dispersion relationship we find in the body of the text using ${\bf u}\sim \omega z \hat{{\bf i}}$.
+Looking at the affiliated disturbances of the relevant free surface and internal wave, we find that
+\[
+\eta_{1} \sim \alpha_{12}\frac{ik}{\Omega}e^{i\theta} + \mbox{cc}, ~ \eta_{2} \sim \alpha_{12}\frac{ik}{\Omega^{3}}\cosh(kh_{0})\left(\Omega^{2}-(1-\tilde{\epsilon})\left( g|k|+s\omega\Omega\right)\right)e^{i\theta}+ \mbox{cc}.
+\]
+Thus if we choose $\Omega$ so that $\Omega^{2}-\left( g|k|+s\omega\Omega\right)= 0$, the magnitude of the internal wave essentially vanishes, thereby localizing dynamics along the free surface near $z=0$.
+Therefore, while not necessarily physically justifiable throughout the bulk of the fluid, our simplified shear profile assumption produces results which are asymptotically consistent with a more sophisticated treatment of the shear profile.
+
+We note though that the more physically realistic shear profile shows that there is a choice of $\Omega$ which corresponds to a non-trivial internal mode near $z=-h_{0}$.
+We also note that this choice of $\Omega$ makes many of the terms in our modulation theory singular, thus showing that this case would require a markedly different treatment.
+While interesting, this issue is beyond the scope of the current paper and will be addressed in future research.
+
+\subsection*{Higher-Order-Dispersive Corrections to the VDE}
+
+To better understand why we must include the higher-order corrections to the dispersion used in the VDE, we examine how such terms are found from first principles.
+To leading order from Equation \eqref{kine} over an infinitely deep fluid we have that
+\[
+\eta_{t} = -\mathcal{H}Q,
+\]
+where $q(x,t) = \phi(x,\eta(x,t),t)$, $Q = q_{x}$, and $\mathcal{H}$ is the Hilbert transform.  Likewise, from Equation \eqref{bernou} we have to leading order that
+\[
+Q _{t}+ \omega\eta_{t} + \eta_{x}= 0, 
+\]
+so that by combining the two expressions we have the wave equation in $\eta$ alone given by 
+\[
+\eta_{tt}- \omega\mathcal{H}\eta_{t} - \mathcal{H}\eta_{x}= 0.
+\]
+Using Fourier transforms, we can readily write a solution to the affiliated initial-value problem in the form
+\[
+\eta(x,t) = \frac{1}{2\pi}\int_{\mathbb{R}}\hat{\eta}_{0}(k) e^{ikx+i\Omega_{+}(k)t} dk + \frac{1}{2\pi}\int_{\mathbb{R}}\hat{\eta}_{0}(k) e^{ikx+i\Omega_{-}(k)t} dk
+\]
+where
+\[
+\Omega_{\pm}(k) = \frac{1}{2}\left(\omega s \pm \sqrt{\omega^{2}+4|k|} \right).
+\]
+If we assume, as per usual when deriving NLS-type models, that 
+\[
+\hat{\eta}_{0}(k) = \frac{1}{\epsilon}\hat{A}\left(\frac{k-k_{0}}{\epsilon} \right), 
+\]
+representing the assumption that the initial conditions are narrowly banded around a carrier wave number $k_{0}$, then we see that, taking only the plus branch of $\Omega_{\pm}$, that  
+\begin{align*}
+\eta(x,t) = & \frac{e^{i\theta(x,t,k_{0})}}{2\pi}\int_{\mathbb{R}}\hat{A}(\tilde{k})e^{i\tilde{k}\xi} e^{i\tau \tilde{\Omega}(k_{0},\tilde{k})}d\tilde{k}\\
+\sim & \frac{e^{i\theta(x,t,k_{0})}}{2\pi}\int_{\mathbb{R}}\hat{A}(\tilde{k})e^{i\tilde{k}\xi}e^{i\tau\left(\frac{\Omega''(k_{0})}{2}\tilde{k}^{2} + \frac{\epsilon}{6}\Omega'''(k_{0})\tilde{k}^{3} + \frac{\epsilon^{2}}{24}\Omega''''(k_{0})\tilde{k}^{4}\right)}d\tilde{k},
+\end{align*}
+where
+\[
+\theta(x,t,k_{0}) = k_{0}x + \Omega(k_{0})t, ~c_{g}(k_{0}) = \Omega'(k_{0}), ~ \xi = \epsilon(x+c_{g}t), ~ \tau = \epsilon^{2}t, 
+\]
+and where
+\[
+\tilde{\Omega}(k_{0},\tilde{k}) = \frac{1}{\epsilon^{2}}\left(\Omega(k_{0}+\epsilon \tilde{k})-\Omega(k_{0})-\Omega'(k_{0})\epsilon\tilde{k}\right) 
+\]
+and where we have expanded the dispersion relationship up to $\mathcal{O}(\epsilon^{3})$ as is done in the main body of the text.
+This expansion corresponds to the affiliated linear-evolution equation for the slowly-evolving envelope $A(\xi,\tau)$
+\[
+i\pd_{\tau}A = \left(\frac{\Omega''(k_{0})}{2}\pd_{\xi}^{2}-\frac{i\epsilon}{6}\Omega'''(k_{0})\pd_{\xi}^{3} - \frac{\epsilon^{2}}{24}\Omega''''(k_{0})\pd_{\xi}^{4}\right)A. 
+\]
+
+As to the issue of why to include the $\epsilon^{2}$ term, we note that the affiliated group velocity of the slowly-evolving envelope is given by 
+\begin{align*}
+\frac{d}{d\tilde{k}}\tilde{\Omega} = & \frac{1}{\epsilon}\left(\Omega'(k_{0}+\epsilon \tilde{k}) - \Omega'(k_{0})\right)\\
+\sim & \Omega''(k_{0})\tilde{k}\left(1 - \frac{3\epsilon c_{g}^{2}}{s}\tilde{k} + 10 \epsilon^{2}c_{g}^{4}\tilde{k}^{2}\right).
+\end{align*}
+We see that the full dispersion relationship has exactly one root at $\tilde{k}=0$, so that this root is the only point of stationary phase in the corresponding integral representation of the solution $\eta(x,t)$.
+However, using our Taylor series expansion, were we to ignore the $\epsilon^{2}$ term, we would introduce a new stationary-phase point at $\tilde{k}_{\ast}$ where
+\[
+\tilde{k}_{\ast} = \frac{s}{3\epsilon c_{g}^{2}}.
+\]
+The presence of this stationary-phase point allows for the accumulation of energy at high-frequencies, which without a corresponding band-limiting requirement introduces non-physical effects into the model.
+By including the $\epsilon^{2}$ term though, we remove this spurious high-frequency stationary-phase point, thereby avoiding the inclusion of an otherwise unnecessary bandwidth condition on the slowly-evolving envelope.
+
+\section*{Acknowledgments}
 The authors would like to thank the support of the NSF through DMS-1715039.

### 2020-01-09 Writeup/Reviewer_Replies/reviewer_replies.tex
This will work?

 \documentclass[a4paper,11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{epstopdf}
\epstopdfsetup{update}

\newcommand{\ba}{\begin{array}}
\newcommand{\ea}{\end{array}}

\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\newcommand{\bc}{\begin{center}}
\newcommand{\ec}{\end{center}}

\newcommand{\ds}{\displaystyle}

\newcommand{\bt}{\begin{tabular}}
\newcommand{\et}{\end{tabular}}

\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}

\newcommand{\bd}{\begin{description}}
\newcommand{\ed}{\end{description}}

\newcommand{\bp}{\begin{pmatrix}}
\newcommand{\ep}{\end{pmatrix}}

\newcommand{\pd}{\partial}
\newcommand{\sech}{\mbox{sech}}

\newcommand{\cf}{{\it cf.}~}

\newcommand{\ltwo}{L_{2}(\mathbb{R}^{2})}
\newcommand{\smooth}{C^{\infty}_{0}(\mathbb{R}^{2})}

\newcommand{\br}{{\bf r}}
\newcommand{\bk}{{\bf k}}
\newcommand{\bv}{{\bf v}}

\newcommand{\gnorm}[1]{\left|\left| #1\right|\right|}
\newcommand{\ipro}[2]{\left<#1,#2 \right>}

\begin{document}
\noindent To the Editor,\\
\\
\indent We wish to thank the reviewers for their helpful comments, insights, and suggestions.  The paper is stronger for their efforts.  See below for a list of replies to their comments.  \\
\\
\noindent Sincerely,\\
\\
Christopher W. Curtis \\
Mackensie Murpy
\subsection*{Replies to the First Reviewer}
\begin{itemize}
\item {\it The authors chose the flow with constant vorticity for simplicity. For the finite depth this assumption looks physically correct, but for deep water it leads to the infinity values of velocities, and the problem becomes only an academic one.}

This comment is an accurate description of our work as it stands.  To address this issue, we have added in the Appendix an examination of the dispersion relationship of a more realistic shear profile which we show generates results very close to those used in the paper.  Thus, while the model we used in the body of the paper for the nonlinear modulation results is not necessarily realistic in all regions of the fluid, the work in the Appendix shows that near the surface it should be sufficient to generate meaningful results.

\item {\it The surface tension is chosen fixed in simulations, and its role is not evident in the obtained results. This item should either be discussed in the text or the surface tension effect should be omitted.}

We have removed all mention of the surface tension in the Results and the Appendix.

\item {\it    By the way, the sigma is applied in the text as the surface tension and the standard deviation. It is better to change one definition.
}

The $\tilde{\sigma}$ was replaced by $\tilde{\sigma}_{s}$ where appropriate.

\item {\it  I would suggest extending a little the theoretical part of the paper to remind the readers of the modulational instability existence in the NLS and vor-Dysthe equations. It is important because the coefficient sings can be either positive or negative in the general case and their values can tend to infinity.}

A discussion on this point has been added in Section 2 just after the VDE has been introduced and before the discussion of the BFI.

\item {\it The case with zero vorticity was studied also early in many papers, see below some of them. The comparison between the results obtained in the reviewed paper with the above-mentioned papers should be darwn discussed.}

References to the four papers were included in the Introduction, and discussion of the results was added throughout the Introduction as well.  Direct comparison of our results to those in said papers is difficult since in this work a space like NLS model is used whereas the referenced papers use a time like one.  Nevertheless, the general similarities in results are commented on in the current revision.   
\end{itemize}

\subsection*{Replies to the Second Reviewer}
\begin{itemize}
\item {\it In the first place, the paper deals with singular phenomena such as rogue waves in the ocean.
In that case the assumption of one-dimensional propagation should be a very big limitation.

In the framework of one-dimensional propagation, as is well known, modulational instability increases
and the probability of occurrence of rogue waves increases as the nonlinearity, i.e. BFI, increases.

However, it is also known that if the directional distribution of the spectrum becomes wide to some extent,
even if the nonlinearity of the system is increased, the occurrence probability of rogue waves and
the value of kurtosis do not change sensitively. I myself have confirmed this fact by numerical calculation based on HOSM.
In recent years, the idea that modulational instability is the key mechanism for the generation of rogue waves seems to have
been questioned considerably.
In any case, it seems somewhat dangerous to discuss the probability of occurrence of rogue waves in a framework
based on one-dimensional propagation.}

So while the author understands the reviewers point, this issue is somewhat complicated by the large amount of existing literature which attempts to link MI to rogue-wave formation.  In particular, see the references the other reviewer recommended that the author discuss in this revision.  To thread this proverbial needle then, the author has in the Introduction softened language around rogue waves while adding the references the other reviewer suggested.  To better address the current reviewer's point though, a citation of the 2016 Scientific Reports paper by Fedele et. al. is provided which presents a counterpoint to the argument that MI is responsible for rogue wave formation.  Hopefully this strikes a more appropriate balance between the various points of view on this complicated issue.

\item {\it Is it necessary to consider surface tension when talking about rogue waves?
I wonder how short wavelength scale the authors have in mind.
}

The authors have removed all mention of the surface tension in the Results section of the paper.  A brief explanation for this decision was provided at the beginnning of the Results section.

\item {\it In page 5, an artificial additional term that does not appear in systematic perturbation analysis seems to be  added to the Dysthe model.
I want some more comments about it.
I do not understand why the numerical instability can be suppressed by the effect of higher-order dispersion rather than dissipation.}

As to this point, the authors have included results in the Appendix clearly explaining the need for the higher-order term in the dispersion relation used in the VDE.

\item {\it I cannot understand the principle of numbering of equations.
Only three specific equations are numbered throughout the paper. In what sense do these equations need special treatment?
Why are the other equations not numbered?}

The authors follow the convention of only numbering those equations which are directly referred to in the body of the text.  The authors generally believe this vastly increases the readability of a paper.  Some inconsistencies in the prior draft though were corrected.

\item {\it How realistic is the value of $\omega=-0.5$ or $\omega=1.12$ which are employed in the work?
What kind of situations do these setting correspond to when returning to the actual physical quantities with dimensions?}

A brief discussion of this point has been added to Section 2 when discussing the non-dimensionalization of the system.  In the end, a choice of $\omega = \mathcal{O}(1)$ corresponds to a wave with a meter long wavelength moving over a current whose horizontal velocity is $300 ~cm/s$ at a depth of one meter below the surface.  While fast, this is not unheard of in real oceanic circumstances, and that rare currents might themselves enhance the likelihood rare events is an interesting issue to explore.

\item {\it In Fig.1(b) through Fig.6 (b), if the spectrum is downshifted, I think $m(\tau)$ should be negative instead of positive.}

This is a reasonable statement, but in the case of the results in the paper, that the peak shifts left but the mean shifts right is in effect a sign of the degree of skewness that the VDE induces in the initial distribution.  A brief explanation of this has been added to Results section in the first case of $\omega=0$.  We do not add a plot of the skewness since this would, in the authors' opinion, clutter the paper.

\item {\it Does Figure~7 show the relationship between BFI at $t=0$ and the kurtosis at the final time?
Looking at the time evolution of kurtosis shown in Fig.~5, for example, the wave field seems to be still evolving
and the kurtosis does not seem to have reached any steady value.
If the value of kurtosis is still fluctuating largely even at the final time,
I wonder if it is reasonable to plot the value as a function of the initial BFI or other parameters.
If it is non-stationary until the final time,
how about plotting the values of kurtosis $\tilde{k}(\tau)$ and ${\rm BFI}(\tau)$ at each time in the temporal evolution
to investigate there relationship?}

Yes, the BFI in Fig. 7 is found at $t=0$.  This is done so as to keep with the original Janssen paper which uses this as a predictive mechanism corresponding to one measurement of a given wavefield.  While we appreciate the reviewer's point, we argue that the fluctuations at the final time in the kurtosis in Fig. 5 are not significant.  Moreover, that particular plot represents only one point used in the fit in Fig. 7, so longer time simulations should not produce significant variations in the stated results.  Further, we feel another plot of the type suggested by the reviewer would clutter the paper and deviate away from the point of trying to use the BFI measured at one time as a longer time predictive tool, which it seems to the authors is the real point in the end of studying it at all.  That being all said, we also reran this particular case up to $t_{f}=60$, which is about a factor of 3 times longer than for the figures used in the paper.  The results did not fundamentally change in any significant way.  A note on this point has been added to the relevant part of the text.

\item {\it Janssen's result that kurtosis is proportional to the square of BFI is derived under the assumptions of one-dimensional
propagation and narrow band, and the proportionality coefficient can be expressed by the coefficients of NLSE.
If using Dysthe equation as a model and taking in to account of higher-order nonlinearity, I suppose that this quadratic relationship should be altered.
If the authors intend to emphasize the importance of the difference between NLSE and Dysthe, in the discussion on Fig.~7 for example,
I expect the author to argue about the deviation from the quadratic relation which the results of Dysthe equation show in relation
to the effect of higher-order nonlinearity included only in the Dysthe equation.}

The authors apologize, but they are not entirely clear what the reviewer is asking for on this point.  A brief statement that the higher-order nonlinearities of the VDE are responsible for the differences seen in Figure 7 is in the original draft.  We added a brief expansion on this original statement.  We hope this is what the reviewer was asking for, but we are happy to modify our explanation further if this is not satisfactory.  
\end{itemize}

\end{document}

### 2020-01-09 Writeup/random_vorticity_revision_1_5_20.tex
Almost there

@@
 A detailed set of theoretical and experimental results \cite{sergeeva1,sergeeva2,sergeeva3,sergeeva4} have also shown that MI could play a significant role in unidirectional deep-water-rogue wave formation among other oceanographic phenomena.
-It is important to note though that there are several possible mechanisms for rogue-wave generation; see \cite{fedele} for example which rules out MI as a mechanism altogether.
+It is important to note though that there are several possible mechanisms for rogue-wave generation; see \cite{fedele} for example which rules out MI as a mechanism altogether in certain oceanic conditions.
 
@@
     u}\right|^{2} + g\eta -
-\frac{\sigma}{\rho}\pd_{x}\frac{\eta_{x}}{\sqrt{1+\eta_{x}^{2}}}= 0, &
+\frac{\sigma_{s}}{\rho}\pd_{x}\frac{\eta_{x}}{\sqrt{1+\eta_{x}^{2}}}= 0, &
 & z=\eta(x,t), \label{bernou}\\
@@
 We note that the form for $\phi$ follows directly from this expansion, allowing for everything to be modeled by the terms in the expansion in $\eta$.
-Here, $s = \mbox{sgn}(k_{0})$, $\omega$ is the non-dimensionalized magnitude of the vorticity of the flow, and $\tilde{\sigma}_{s}$ is the surface tension, or where the reciprocal of the Bond number, $\tilde{\sigma}$, is given by
+Here, $s = \mbox{sgn}(k_{0})$, $\omega$ is the non-dimensionalized magnitude of the vorticity of the flow, and $\tilde{\sigma}_{s}$ is the non-dimensional surface tension given by
 \[
-\tilde{\sigma}_{s} = \frac{\sigma}{\rho g L_{p}^2}.
+\tilde{\sigma}_{s} = \frac{\sigma_{s}}{\rho g L_{p}^2}.
 \]
@@
 
-The traditional definition of modulational instability (MI) begins by studying the stability of the plane-wave, or Stokes wave, solutions of the NLS equation given by
+The traditional definition of modulational instability (MI) begins by studying the stability of the plane-wave, or Stokes wave, solutions of the NLSE given by
 \[
@@
 Our results likewise support this.
-However, also shown in \cite{janssen} and followed up on in several later works; see \cite{onorato,slunyaev,eeltink}, it is shown that the kurtosis of the corresponding distribution of the surface waves, say $\tilde{k}(\tau)$ is a quadratic function of the time evolving $\mbox{BFI}(\tau)$.
+However, also shown in \cite{janssen} and followed up on in several later works; see \cite{onorato,slunyaev,eeltink}, it is shown that the kurtosis of the corresponding quasi-steady distribution of the surface waves, say $\tilde{k}(\tau_{f})$, where we take $\tau_{f}$ to be the timescale over which MI acts, is a quadratic function of the the $\mbox{BFI}$.
 We likewise aim to establish a similar relationship, thereby maintaining the utility of the BFI insofar as it then helps predict a key feature of nonlinear-random wave fields.
@@
 The number of modes is $K_{T} = 1024$ or $K=512$.
-Letting the length scale $L_{p}=1m$, the non-dimensional surface tension is $\tilde{\sigma}_{s} = 10^{-5}$.
+Letting the length scale $L_{p}=1~m$, the non-dimensional surface tension is $\tilde{\sigma}_{s} = 10^{-5}$.
 It is thus ignored throughout the remainder of the paper.

### 2020-01-09 nls_Dysthe_comparison.m

@@ -75,12 +75,14 @@ function [nls_avg,nls_std,nls_kts,nls_bfi,dysthe_avg,dysthe_std,dysthe_kts,dysth
     plot(kvec,pdist_int,'k:',kvec,pdist_nls,'k--',kvec,pdist_dysthe,'k-','LineWidth',2)
     h = set(gca,'FontSize',30);
     set(h,'Interpreter','LaTeX')
+    xlim([-5,5])
     xlabel('$k$','Interpreter','LaTeX','FontSize',30)
     ylabel('$S(k,\tau_{f},\sigma)$','Interpreter','LaTeX','FontSize',30)
     %legend({'Initial','NLS','Dysthe'},'Interpreter','LaTeX','FontSize',30)
     
     figure(2)
     plot(samp_times,nls_avg_plot,'k--',samp_times,dysthe_avg_plot,'k-','LineWidth',2)
+    h = set(gca,'FontSize',30);
     set(h,'Interpreter','LaTeX')
     xlabel('$\tau$','Interpreter','LaTeX','FontSize',30)
     ylabel('$m(\tau)$','Interpreter','LaTeX','FontSize',30)
@@ -88,6 +90,7 @@ function [nls_avg,nls_std,nls_kts,nls_bfi,dysthe_avg,dysthe_std,dysthe_kts,dysth
         
     figure(3)
     plot(samp_times,nls_std_plot,'k--',samp_times,dysthe_std_plot,'k-','LineWidth',2)
+    h = set(gca,'FontSize',30);
     set(h,'Interpreter','LaTeX')
     xlabel('$\tau$','Interpreter','LaTeX','FontSize',30)
     ylabel('$\tilde{\sigma}(\tau)$','Interpreter','LaTeX','FontSize',30)
@@ -95,12 +98,14 @@ function [nls_avg,nls_std,nls_kts,nls_bfi,dysthe_avg,dysthe_std,dysthe_kts,dysth
     
     figure(4)
     plot(samp_times,nls_skw_plot,'k--',samp_times,dysthe_skw_plot,'k-','LineWidth',2)
+    h = set(gca,'FontSize',30);
     set(h,'Interpreter','LaTeX')
     xlabel('$\tau$','Interpreter','LaTeX','FontSize',30)
     ylabel('$\tilde{s}(\tau)$','Interpreter','LaTeX','FontSize',30)
         
     figure(5)
[…]

### 2020-01-09 plot_maker.m

@@ -6,7 +6,7 @@ dk = pi/Llx;
 ep = .05;
 dt = 2e-3;
 om = 1.12;
-sig = 1e-5;
+sig = 0;
 Nens = 512;
 k0 = 1;
 [Om,cg,ad,anl] = param_maker(k0,om,sig);
